### 2017-04-03 paper.tex
Initialized paper with structure, added one picture, added a .gitignore.

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
 \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
 \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
 \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Enabling Ambient Backscatter Using Low-cost Software-defined-radio}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Maximilian Stiefel}
\IEEEauthorblockA{Master Programme Embdedded Systems\\
Uppsala University\\
maximilian.stiefel.8233@student.uu.se}
\and
\IEEEauthorblockN{Elmar van Rijnswou}
\IEEEauthorblockA{Master Programme Embdedded Systems\\
Uppsala University\\
elmar.vanrijnswou.9818@student.uu.se}
\and
\IEEEauthorblockN{Carlos Pérez-Penichet}
\IEEEauthorblockA{Communication Research Group\\
Uppsala University\\
carlos.penichet@it.uu.se}
\and
\IEEEauthorblockN{Ambuj Varshney}
\IEEEauthorblockA{Communication Research Group\\
Uppsala University\\
ambuj.varshney@it.uu.se}
\and
\IEEEauthorblockN{Christian Rohner}
\IEEEauthorblockA{Communication Research Group\\
Uppsala University\\
christian.rohner@it.uu.se}
\and
\IEEEauthorblockN{Thiemo Voigt}
\IEEEauthorblockA{Communication Research Group\\
Uppsala University\\
thiemo.voigt@it.uu.se}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Backscatter communication is attractive for energy-constrained devices due to its very low power requirements. Ambient backscatter takes this aspect to the limit by leveraging existing radio frequency signals for the purpose of communication without the need for generating energy expensive carrier signal. In this paper we investigate the use of ambient television broadcast signals for communication. As opposed to state-of-the-art restricted to operations under conditions of strong signal strength, we demonstrate a low cost software defined radio as receiver 
enables operations even in conditions when ambient signals are  weak in strength. We build the system using low-cost off-the-shelf microcontroller, and RTLSDR software-defined radio receiver. We also conduct survey of signal strength of TV broadcast in a mid-sized swedish city, and observe that our system can operate in most parts of the city.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
This demo file is intended to serve as a ``starter file''
for IEEE conference papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Bachground}

\subsection{RTL2832U}

\section{Design}

\subsection{Sender}

\subsection{Receiver}
\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{fig/receiver_arch}
\caption{This figure shows the receiver architecture from an abstract point of view.}
\label{fig_receiver_arch}
\end{figure}

\section{Evaluation}

\subsection{Signal Strength Variation within the City}

\subsection{Signal Strength Variation over Time}

\subsection{Communication Performance}

\section{Discussion}

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2017-04-03 paper.tex
Wrote a lot about the receiver. Done for today.

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
 %
-% \usepackage{dblfloatfix}
+% dblfloatfix}
 % The latest version can be found at:
@@
 %
-%\usepackage{url}
+\usepackage{url}
 % url.sty was written by Donald Arseneau. It provides better support for
@@
 \subsection{RTL2832U}
+\label{sub:rtl2832}
 
@@
 \subsection{Receiver}
-\begin{figure}[!t]
+\begin{figure}[h]
 \centering
-\includegraphics[width=2.5in]{fig/receiver_arch}
+\includegraphics[width=\columnwidth]{fig/receiver_arch}
 \caption{This figure shows the receiver architecture from an abstract point of view.}
-\label{fig_receiver_arch}
+\label{fig:receiver_arch}
 \end{figure}
+The receiver was the most work intensive part of the design.
+C++ has been used to implement the receiver.
+We provide the code publicly available under \url{https://github.com/s3xm3x/backscatterBASKReceiver}.
+In figure \ref{fig:receiver_arch} one can see the architecture of the receiver from an abstarct point of view.
+A highly sophisticated bus system has been developed to exchange data between the different components of the system.
+The signal flow is from the left to the right.
+\textit{Librtlsdr} (cf.
+\cite{github:librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program.
+As described under \ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are in SW adjustable).
+Subsequently the data is mixed with sine and cosine, low-pass filtered and finally transfered in the digital world.
+This data is available as two 8 Bit values.
+It is a well-known fact, that this technique, which is commonly known as I/Q demodulation, results in a real and an imagenery part.
+With this two values phase and absolute value of the data can be determined for every sample.
+So every sample consists out of a real and and imaginary part.
+And can be seen as:
+\begin{equation}
+	I+j\,Q = abs(I,Q) \cdot e^{ang(I,Q)} 
+\end{equation} 
+So the first block, which is entitled with RTL2832U, provides the interface to the TV dongle.
+This block is represented as one object.
+It can set the frequency \ensuremath{f_{tuned}}, where the receiver is listening.
+Also it can set the sampling frequency \ensuremath{f_{samp}} and the analog gain of the tuner.
+The raw samples are pushed to the bus.
+Demodulator is the block called, which is responsible to convert the sampled signals into something, which looks like ones and zeroes.
+Thereforre the real and imgainary values first have to be deinterleaved out of the sample message.
+Now they are represented as integer values between 0 and 255 (8 bit).
+So one converts them to floats, whereas 127.5 equals 0.
+After that the values have to be downsampled.
+Usually we were sampling with 250 kHz.
+This is still much more than we need assuming an additive white gausian noise (AWGN) channel with a \ensuremath{S/N = 10\,dB} (cf.
+equation \ref{eq:awgn}).
+So we decided to reduce the sampling frequency to 25 kHz.
+\begin{equation}
+	\label{eq:awgn}
+	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
+\end{equation}     
+To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
+\begin{equation}
+	\label{eq:nyquist}
+	f_{sampling} > 2 \cdot f_{highestfregencyinsignal}
+\end{equation}
+A FIR filter naivly implemented with floats has been used to achieve this.
+10 kHz is the cut-off frequency of the first filter.
+All filters have been constructed with the internal filter design tool of GNU radio.
+After downsampling another filter is used to suppress noise.
+This filter has a cut-off frequency of 3 kHz.
+The last demodulation step is to take the signal, rectify it (cf.
+equation \ref{eq:rectify}) and decide with a software-defined Schmitt trigger whether a sample is a 1 or a 0.
+\begin{equation}
+	\label{eq:rectify}
+	abs(I,Q) = \sqrt{I^2 +  Q^2}
+\end{equation} 
+These samples are broadcasted on the bus.
+A registered listener receives the processed samples.
+This registered listener is the decoder, which is represented as an object as well.
+Thereupon the decoder decides when a frame is sent or when the channel is idle.
+On the basis of a defined threshold the decoder decides wether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
+The baudrate value and the sampling rate value lead to a certain length of one bit.
+For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combinded with a baudrate of 1 kbit/s lead to 25 samples/bit.
 
@@
 
+\bibitem{github:librtlsdr}
+Steve Markgraf et. al., \emph{librtlsdr}, \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
+
 \end{thebibliography}

### 2017-04-04 paper.tex
Finished receiver description and description of RTL2832U. Lunch now.

@@
 % no \IEEEPARstart
-This demo file is intended to serve as a ``starter file''
-for IEEE conference papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
 % You must have at least 2 lines in the paragraph with the drop letter
 % (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
-
-\hfill August 26, 2015
+\hfill April 7, 2017
 
@@
 
-\section{Bachground}
+\section{Background}
 
@@
 \label{sub:rtl2832}
-
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
+\caption{This figure is a photo of the RTLSDR used in this project.}
+\label{fig:receiver_arch}
+\end{figure}
+The Raeltek RTL2832U is a terrestrial digital video broadcast (DVB-T) chip, which is the basis of a very low-cost software defined radio: The RTL-SDR.
+One feature of this chip is, that it supports USB 2.0.
+Ham radio enthusiasts combined their efforts to enable accessing the raw I/Q data.
+With their software driver (basically the \textit{librtlsdr}, cf.
+\cite{github:librtlsdr}) DVB-T sticks, which are based on the RTL2832, can be converted into a wideband SDR.
+This sort of hardware has been in the range of a few thousands USD a few years ago.
+An RTL2832U based DVB-T receiver can be bought with antenna for less than 10 USD.
+What is a software defined radio?
+Traditionally basic radio components like mixers or filters have been realized in hardware.
+Nowadays as soon as one works in the baseband frequency the coputational resources of a personal computer are sufficient to realize these components with signal processing done in software.
+An RTL-SDR does not only consist of the RTL2832U.
+The second essential component is the tuner, which mixes down the radio frequency (RF) to a intermediate frequency (IF).
+In the case of the used hardware this tuner component was represented by a \textit{Rafael Micro R820T/2}.
+It subsumes a low-noise amplifier (LNA) and filters.
+The image frequency is rejected automatically with 65 dBc and the noise figure is with 3.5 dB quite low.
+It alows a frequency range from 42 to 1002 MHz (cf.
+\cite{rafael:r820t}).
+The I/Q demodulator, also shown in figure \ref{fig:receiver_arch}, is the most important part of the \textit{RTL2832U} for our work. The received signal can be intepreted as
+\begin{equation}
+	s_{IF}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
+\end{equation}     
+This is multiplied with a cosine (and a sine as well) from a local oscillator.
+\begin{multline}
+        s_{IF}(t) \cdot s_L(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+\end{multline}
+With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
+\begin{multline}
+        2 \cdot s_{IF}(t) \cdot s_L(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+\end{multline}
+One can see, that the interesting inphase part in this case is represented by a DC value after mixing.
+With a low-pass filter this DC value can be seperated from the undesired rest.
+Similar things are hapening with the quadrature part when mixing with a sine, which is indeed done to get the quadrature part.
 \section{Design}
@@
 \centering
-\includegraphics[width=\columnwidth]{fig/receiver_arch}
+\includegraphics[width=\columnwidth]{./fig/receiver_arch}
 \caption{This figure shows the receiver architecture from an abstract point of view.}
@@
 	\label{eq:nyquist}
-	f_{sampling} > 2 \cdot f_{highestfregencyinsignal}
+	f_{samp} \geq 2 \cdot f_{a}
 \end{equation}
+\ensuremath{f_{a}} is the highest frequency in the signal. 
 A FIR filter naivly implemented with floats has been used to achieve this.
@@
 This registered listener is the decoder, which is represented as an object as well.
-Thereupon the decoder decides when a frame is sent or when the channel is idle.
-On the basis of a defined threshold the decoder decides wether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
-The baudrate value and the sampling rate value lead to a certain length of one bit.
-For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combinded with a baudrate of 1 kbit/s lead to 25 samples/bit.
+After the decoder received the samples it decides when a frame is sent or when the channel is idle.
+On the basis of a defined threshold the decoder determines wether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
+The baudrate and the sampling rate lead to a certain length of one bit.
+For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combinded with a baudrate of 1 kbit/s leads to 25 samples/bit.
+The decoder also includes a function to correlate the received bitstream with an expected pattern.
+Hence it is able to determine the bit ro error ratio (BER).
+Furthermore the code we have written so far subsumes a simulator to simulate ideal data coming out of the demodulator to test the algorithms behind the decoder.
+Also a simulator for the RTL2832U has been written to replay recorded samples, which facialiates debugging as well.
 
@@
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
 \bibitem{github:librtlsdr}
 Steve Markgraf et. al., \emph{librtlsdr}, \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
+
+\bibitem{rafael:r820t}
+\emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
 

### 2017-04-04 paper.tex
Wrote about the antenna design. Sender part almost done.

@@
 
+% For special characters like ä,ö,...
+\usepackage[utf8x]{inputenc}
+
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 \subsection{Sender}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.6\columnwidth]{./fig/s11}
+\caption{A plot of the input reflection coefficient or \ensuremath{S_{\text{11}}} is shown in this figure.}
+\label{fig:s11}
+\end{figure}
+A backscattering antenna is the most important part of backscattering sender.
+Our antenna is optimized for usage with the television signal, which has its center frequency at 626 MHz.
+It is a ground plane antenna, which has been built up on a big PCB.
+This PCB is covered with a copper layer on one side and left blank on the other.
+A hole in the midle of the PCB serves for connecting a thicker wire to a 50 \ensuremath{\Omega} microstrip transmission line.
+The wire is the vertical antenna element and is a bit longer than \ensuremath{\lambda/4 \approx \text{239.6 mm}}.
+It is soldered to the transmission line.
+The transmission line has been calculated for 626 MHz using FR4 substrate with a thickness of 1.6 mm.
+According to microstrip line theory the length of the line changes the phase and the width of the line changes the impedance (cf.
+chapter 3 in \cite{book:pozar}).
+The design equations given in \cite{books:pozar} are implemented in various programs e.g.
+KiCAD, which has been used for solving the design equations.
+The given parameters lead to a microstrip line width of 3 mm.
+An open-stub, attached directly at the puncture where the vertical antenna element is attached to the transmission line, is used for fine tuning of the input impedance.
+The stub length has been adjusted manually (shortened) while measuring the reflection coefficient with a vector network analyzer (VNA).
+Open-stub matching is discussed in chapter 5 in \cite{book:pozar}.
+Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA.
+At some point the \ensuremath{S_{\text{11}}} had the minimum quite close to where we wanted to have it (cf.
+\ref{fig:s11}).
+In the plot one can also see, that the antenna is quite narrowband.
+For backscattering this is not necessarily a disadvantage, since everything, which is within the operating frequency of the antenna is backscattered.
+So with a narrowband antenna the effect, that signals are backscattered, which are not targeted, is reduced.
+This statement has to be qualified as the antenna also has multiple resonance frequencies in multiple bands.
+With our approach we were able to achieve a input resistance at 634 MHz of \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}.
+At 626 MHz \ensuremath{S_{\text{11}}} is still approximately -17 dB.
 
@@
 Also a simulator for the RTL2832U has been written to replay recorded samples, which facialiates debugging as well.
+Another handy utility, which has been written in Octave, is a tool (an oscilloscope) to look at the data, which is written into files by the demodulator.
+This utility has been captured in action, as can be seen in figure \ref{fig:osci}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{./fig/osci}
+\caption{A transmission of 101010... displayed with the Octave oscilloscope is shown in this figure.}
+\label{fig:osci}
+\end{figure}
 
@@
 \bibitem{github:librtlsdr}
-Steve Markgraf et. al., \emph{librtlsdr}, \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
+Steve Markgraf et. al. \emph{librtlsdr}. \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
 
@@
 \emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
+
+\bibitem{book:pozar}
+D. M. Pozar. \emph{Microwave Engineering}. Fourth Edition. John Wiley \& Sons Inc. 111 River Street, Hoboken. 2012
 

### 2017-04-05 paper.tex
Made some captions more descriptive. And changed plots, that one can see them better.

@@
 \includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
-\caption{This figure is a photo of the RTLSDR used in this project.}
+\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner with integrated LNA \textit{Rafael Micro R820T/2} (right IC).}
 \label{fig:receiver_arch}
@@
 
-\subsection{Sender}
+\subsection{Transmitter}
 \begin{figure}[h]
 \centering
-\includegraphics[width=0.6\columnwidth]{./fig/s11}
-\caption{A plot of the input reflection coefficient or \ensuremath{S_{\text{11}}} is shown in this figure.}
+\includegraphics[width=0.8\columnwidth]{./fig/s11v2}
+\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the selfmade backsacttering ground plane antenna in dB.}
 \label{fig:s11}
@@
 It is a ground plane antenna, which has been built up on a big PCB.
-This PCB is covered with a copper layer on one side and left blank on the other.
+This rectangular PCB, with dimensions of 200 mm x 300 mm, is covered with a copper layer on one side and left blank on the other.
 A hole in the midle of the PCB serves for connecting a thicker wire to a 50 \ensuremath{\Omega} microstrip transmission line.
-The wire is the vertical antenna element and is a bit longer than \ensuremath{\lambda/4 \approx \text{239.6 mm}}.
+The wire is the vertical antenna element has a length of 350 mm, which is roughly 0.73 \ensuremath{\lambda}.
 It is soldered to the transmission line.
@@
 At 626 MHz \ensuremath{S_{\text{11}}} is still approximately -17 dB.
+Besides the antenna the transmitter consists out of a \textit{MSP430}, which acts as subscriber, that sends data to the receiver.
+This transmitter is connected with one I/O pin and ground to a rf switch.
+The switch can change between matched (50 \ensuremath{\Omega}) and open.
+The software now steers this switch with a two 2 MHz rectangular timer signal for a 1 and is just leavs it open for a 0.
+Therefore the result is a shift of the television signal by 2 MHz for a 1 and no shift for a 0.
+Hence a classical binary amplitude shift keying is the implemented modulation technique.
 
@@
 \includegraphics[width=\columnwidth]{./fig/receiver_arch}
-\caption{This figure shows the receiver architecture from an abstract point of view.}
+\caption{Receiver architecture from a system point of view. Sinal flow is from the left to the right.}
 \label{fig:receiver_arch}
@@
 Another handy utility, which has been written in Octave, is a tool (an oscilloscope) to look at the data, which is written into files by the demodulator.
-This utility has been captured in action, as can be seen in figure \ref{fig:osci}.
+An example of a recorded data stream with 25 kS/s can be seen in figure \ref{fig:osci}.
 \begin{figure}[h]
 \centering
-\includegraphics[width=\columnwidth]{./fig/osci}
-\caption{A transmission of 101010... displayed with the Octave oscilloscope is shown in this figure.}
-\label{fig:osci}
+\includegraphics[width=\columnwidth]{./fig/frame_and_idle}
+\caption{Transmission of 101010.. with 1 kbit/s and idle state. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.}
+\label{fig:frame}
 \end{figure}
@@
 \subsection{Signal Strength Variation within the City}
+To observe the signal strength variation within the city the \textit{RTL-SDR} has been used as well.
 

### 2017-04-06 paper.tex
Moved bibliography to bibtex

@@
 With their software driver (basically the \textit{librtlsdr}, cf.
-\cite{github:librtlsdr}) DVB-T sticks, which are based on the RTL2832, can be converted into a wideband SDR.
+\cite{steve-m_librtlsdr}) DVB-T sticks, which are based on the RTL2832, can be converted into a wideband SDR.
 This sort of hardware has been in the range of a few thousands USD a few years ago.
@@
 It alows a frequency range from 42 to 1002 MHz (cf.
-\cite{rafael:r820t}).
+\cite{rafael_r820t_2011}).
 The I/Q demodulator, also shown in figure \ref{fig:receiver_arch}, is the most important part of the \textit{RTL2832U} for our work. The received signal can be intepreted as
@@
 According to microstrip line theory the length of the line changes the phase and the width of the line changes the impedance (cf.
-chapter 3 in \cite{book:pozar}).
-The design equations given in \cite{books:pozar} are implemented in various programs e.g.
+chapter 3 in \cite{pozar_microwave_2011}).
+The design equations given in \cite{pozar_microwave_2011} are implemented in various programs e.g.
 KiCAD, which has been used for solving the design equations.
@@
 The stub length has been adjusted manually (shortened) while measuring the reflection coefficient with a vector network analyzer (VNA).
-Open-stub matching is discussed in chapter 5 in \cite{book:pozar}.
+Open-stub matching is discussed in chapter 5 in \cite{pozar_microwave_2011}.
 Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA.
@@
 \textit{Librtlsdr} (cf.
-\cite{github:librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program.
+\cite{steve-m_librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program.
 As described under \ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are in SW adjustable).
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliography{paper}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
 %
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{github:librtlsdr}
-Steve Markgraf et. al. \emph{librtlsdr}. \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
-
-\bibitem{rafael:r820t}
-\emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
-
-\bibitem{book:pozar}
-D. M. Pozar. \emph{Microwave Engineering}. Fourth Edition. John Wiley \& Sons Inc. 111 River Street, Hoboken. 2012
-
-\end{thebibliography}
+%\begin{thebibliography}{1}
+%
+%\bibitem{github:librtlsdr}
+%Steve Markgraf et. al. \emph{librtlsdr}. \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
+%
+%\bibitem{rafael:r820t}
+%\emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
+%
+%\bibitem{book:pozar}
+%D. M. Pozar. \emph{Microwave Engineering}. Fourth Edition. John Wiley \& Sons Inc. 111 River Street, Hoboken. 2012
+%
+%\end{thebibliography}
 

### 2017-04-06 paper.tex
Some reordering in the background section

@@
 
+\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
+
 \begin{document}
@@
 \IEEEauthorblockN{Carlos Pérez-Penichet}
-\IEEEauthorblockA{Communication Research Group\\
-Uppsala University\\
+\IEEEauthorblockA{Uppsala University\\
 carlos.penichet@it.uu.se}
@@
 \IEEEauthorblockN{Ambuj Varshney}
-\IEEEauthorblockA{Communication Research Group\\
-Uppsala University\\
+\IEEEauthorblockA{Uppsala University\\
 ambuj.varshney@it.uu.se}
@@
 \IEEEauthorblockN{Christian Rohner}
-\IEEEauthorblockA{Communication Research Group\\
-Uppsala University\\
+\IEEEauthorblockA{Uppsala University\\
 christian.rohner@it.uu.se}
@@
 \IEEEauthorblockN{Thiemo Voigt}
-\IEEEauthorblockA{Communication Research Group\\
-Uppsala University\\
-thiemo.voigt@it.uu.se}}
+\IEEEauthorblockA{Uppsala University\\
+and SICS Swedish ICT\\
+thiemo@sics.se}}
 
@@
 
+In this section we examine some background aspects that are essential to our
+work. \cp{Put a simple intro to each section so that they don't look empty
+before the first subsection.}
+
+\subsection{Software Defined Radio}
+Traditionally basic radio components like mixers or filters have been realized
+in hardware. Nowadays as soon as one works in the baseband frequency the
+computational resources of a personal computer are sufficient to realize these
+components with signal processing done in software. \cp{We need to expand on
+this introduction to SDR, assuming we need it and not just assume everyone
+knows what that is.}
+
 \subsection{RTL2832U}
@@
 \end{figure}
-The Raeltek RTL2832U is a terrestrial digital video broadcast (DVB-T) chip, which is the basis of a very low-cost software defined radio: The RTL-SDR.
-One feature of this chip is, that it supports USB 2.0.
-Ham radio enthusiasts combined their efforts to enable accessing the raw I/Q data.
-With their software driver (basically the \textit{librtlsdr}, cf.
-\cite{steve-m_librtlsdr}) DVB-T sticks, which are based on the RTL2832, can be converted into a wideband SDR.
-This sort of hardware has been in the range of a few thousands USD a few years ago.
-An RTL2832U based DVB-T receiver can be bought with antenna for less than 10 USD.
-What is a software defined radio?
-Traditionally basic radio components like mixers or filters have been realized in hardware.
-Nowadays as soon as one works in the baseband frequency the coputational resources of a personal computer are sufficient to realize these components with signal processing done in software.
+The Raeltek RTL2832U is a terrestrial digital video broadcast (DVB-T) demodulator IC, which is at the core of a wide range of USB-based DVB-T receiver dongles.
+One prominent feature of this chip is that it allows to retrieve the raw I/Q samples via USB.
+This is originally intended for the chip to work as a simple DAB/FM software receiver.
+Ham radio enthusiasts have combined their efforts to create a software driver (the \textit{librtlsdr}, cf.
+\cite{steve-m_librtlsdr}) that allows DVB-T dongles based on the RTL2832 to be converted into a low-cost wideband SDR receiver.
+The cost of traditional SDR's has been in the range of a few hundred to thousands of USD.
+Albeit much less powerful than the typical SDR, an RTL2832U-based DVB-T receiver can be bought with antenna for less than 10 USD.
 An RTL-SDR does not only consist of the RTL2832U.
-The second essential component is the tuner, which mixes down the radio frequency (RF) to a intermediate frequency (IF).
-In the case of the used hardware this tuner component was represented by a \textit{Rafael Micro R820T/2}.
+The second essential component is the tuner, which allows the user to select the received frequency and outputs that signal at an intermediate frequency (IF).
+The IF can then be fed into the demodulator for digitalization.
+In the case of the used hardware this tuner is a \textit{Rafael Micro R820T/2} which offers a tuning range from 42 to 1002 MHz~\cite{rafael_r820t_2011}.
 It subsumes a low-noise amplifier (LNA) and filters.
-The image frequency is rejected automatically with 65 dBc and the noise figure is with 3.5 dB quite low.
-It alows a frequency range from 42 to 1002 MHz (cf.
-\cite{rafael_r820t_2011}).
-The I/Q demodulator, also shown in figure \ref{fig:receiver_arch}, is the most important part of the \textit{RTL2832U} for our work. The received signal can be intepreted as
+It presents a low noise figure of 3.5 dB and features 65 dBc of image rejection.
+Figure \ref{fig:receiver_arch} \cp{We need a much better photo here.}, shows a photograph of the RTL-SDR used in our work.
+
+\subsection{Quadrature Demodulation}
+
+The received signal can be interpreted as
 \begin{equation}
@@
 \end{multline}
-One can see, that the interesting inphase part in this case is represented by a DC value after mixing.
-With a low-pass filter this DC value can be seperated from the undesired rest.
-Similar things are hapening with the quadrature part when mixing with a sine, which is indeed done to get the quadrature part.
+One can see, that the interesting in-phase part (I) in this case is represented by a DC value after mixing.
+With a low-pass filter this DC value can be separated from the undesired rest.
+An analogous procedure is done with the quadrature part (Q) when mixing with a sine.
+\cp{We need to explain this more clearly and possibly more concisely.}
+
 \section{Design}

### 2017-04-06 paper.tex
Some corrections in the design section

@@
 \includegraphics[width=0.8\columnwidth]{./fig/s11v2}
-\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the selfmade backsacttering ground plane antenna in dB.}
+\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
 \label{fig:s11}
 \end{figure}
-A backscattering antenna is the most important part of backscattering sender.
+A backscattering antenna is the most important part of backscatter transmitter.
 Our antenna is optimized for usage with the television signal, which has its center frequency at 626 MHz.
-It is a ground plane antenna, which has been built up on a big PCB.
+It is a ground plane antenna, which has been built up on a large unetched PCB plate.
 This rectangular PCB, with dimensions of 200 mm x 300 mm, is covered with a copper layer on one side and left blank on the other.
-A hole in the midle of the PCB serves for connecting a thicker wire to a 50 \ensuremath{\Omega} microstrip transmission line.
-The wire is the vertical antenna element has a length of 350 mm, which is roughly 0.73 \ensuremath{\lambda}.
-It is soldered to the transmission line.
-The transmission line has been calculated for 626 MHz using FR4 substrate with a thickness of 1.6 mm.
-According to microstrip line theory the length of the line changes the phase and the width of the line changes the impedance (cf.
-chapter 3 in \cite{pozar_microwave_2011}).
+A hole in the middle of the PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip transmission line.
+The wire, which acts as the vertical antenna element and has a length of 350 mm (roughly 0.73 \ensuremath{\lambda}), is soldered to the transmission line.
+The transmission line has been calculated for 626 MHz using an FR4 substrate with a thickness of 1.6 mm.
+According to micro strip line theory the length of the line changes the phase and the width of the line changes the impedance (cf.
+Chapter 3 in \cite{pozar_microwave_2011}).
 The design equations given in \cite{pozar_microwave_2011} are implemented in various programs e.g.
 KiCAD, which has been used for solving the design equations.
-The given parameters lead to a microstrip line width of 3 mm.
-An open-stub, attached directly at the puncture where the vertical antenna element is attached to the transmission line, is used for fine tuning of the input impedance.
+The given parameters lead to a micro strip line width of 3 mm.
+An open-stub, attached directly at the juncture where the vertical antenna element is attached to the transmission line, is used for fine tuning of the input impedance.
 The stub length has been adjusted manually (shortened) while measuring the reflection coefficient with a vector network analyzer (VNA).
-Open-stub matching is discussed in chapter 5 in \cite{pozar_microwave_2011}.
-Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA.
-At some point the \ensuremath{S_{\text{11}}} had the minimum quite close to where we wanted to have it (cf.
-\ref{fig:s11}).
-In the plot one can also see, that the antenna is quite narrowband.
-For backscattering this is not necessarily a disadvantage, since everything, which is within the operating frequency of the antenna is backscattered.
-So with a narrowband antenna the effect, that signals are backscattered, which are not targeted, is reduced.
-This statement has to be qualified as the antenna also has multiple resonance frequencies in multiple bands.
-With our approach we were able to achieve a input resistance at 634 MHz of \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}.
+Open-stub matching is discussed in Chapter 5 in \cite{pozar_microwave_2011}.
+Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA until \ensuremath{S_{\text{11}}} had the minimum close to the desired frequency (Figure~\ref{fig:s11}\cp{Can we replace all these figures by PDFs?}).
+In the plot one can also see, that the antenna is quite narrowband\cp{We should report something like the 3dB bandwidth rather than saying something non-specific like this}.
+For backscattering this is not necessarily a disadvantage, since everything that is within the operating frequency of the antenna is backscattered.
+With our narrowband antenna the undesired backscattering of out-of-band signals is reduced.
+With our approach we were able to achieve an input impedance at 634 MHz of \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}.
 At 626 MHz \ensuremath{S_{\text{11}}} is still approximately -17 dB.
-Besides the antenna the transmitter consists out of a \textit{MSP430}, which acts as subscriber, that sends data to the receiver.
-This transmitter is connected with one I/O pin and ground to a rf switch.
-The switch can change between matched (50 \ensuremath{\Omega}) and open.
-The software now steers this switch with a two 2 MHz rectangular timer signal for a 1 and is just leavs it open for a 0.
-Therefore the result is a shift of the television signal by 2 MHz for a 1 and no shift for a 0.
+
+Besides the antenna, the transmitter consists of an \textit{MSP430} MCU, that sends data to the receiver.
+This transmitter controls an RF switch through an I/O pin.
+The switch can alternate the impedance connected to the antenna between matched (50 \ensuremath{\Omega}) and open.
+The software now steers this switch with a two 2 MHz rectangular timer signal to transmit a data 1 and is just leaves it open to transmit a data 0.
+Therefore the result is a frequency shift\cp{We need to explain this in the background} of the television signal by 2 MHz for a 1 and no shift for a 0.
 Hence a classical binary amplitude shift keying is the implemented modulation technique.
@@
 \includegraphics[width=\columnwidth]{./fig/receiver_arch}
-\caption{Receiver architecture from a system point of view. Sinal flow is from the left to the right.}
+\caption{Receiver architecture from a system point of view. Signal flow is from left to right.}
 \label{fig:receiver_arch}
 \end{figure}
-The receiver was the most work intensive part of the design.
 C++ has been used to implement the receiver.
 We provide the code publicly available under \url{https://github.com/s3xm3x/backscatterBASKReceiver}.
-In figure \ref{fig:receiver_arch} one can see the architecture of the receiver from an abstarct point of view.
+In figure \ref{fig:receiver_arch} we show the architecture of the receiver from an abstract point of view.
 A highly sophisticated bus system has been developed to exchange data between the different components of the system.
@@
 \cite{steve-m_librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program.
-As described under \ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are in SW adjustable).
-Subsequently the data is mixed with sine and cosine, low-pass filtered and finally transfered in the digital world.
-This data is available as two 8 Bit values.
-It is a well-known fact, that this technique, which is commonly known as I/Q demodulation, results in a real and an imagenery part.
-With this two values phase and absolute value of the data can be determined for every sample.
-So every sample consists out of a real and and imaginary part.
+As described in section~\ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are software adjustable).
+Subsequently the data is mixed with sine and cosine, low-pass filtered and finally transferred to the digital domain.
+This data is available as two 8-bit values.
+It is a well-known fact, that this technique, which is commonly known as I/Q demodulation, results in a decomposition of the signal into in-phase and quadrature components which are generally represented as real and imaginary parts respectively.
+With this two values, the phase and absolute value of the data can be determined for every sample.
 And can be seen as:
 \begin{equation}
-	I+j\,Q = abs(I,Q) \cdot e^{ang(I,Q)} 
+	I+j\,Q = abs(I,Q) \cdot e^{j \times ang(I,Q)} 
 \end{equation} 
-So the first block, which is entitled with RTL2832U, provides the interface to the TV dongle.
+So the first block, which is entitled RTL2832U, provides the interface to the TV dongle.
 This block is represented as one object.
@@
 The raw samples are pushed to the bus.
-Demodulator is the block called, which is responsible to convert the sampled signals into something, which looks like ones and zeroes.
-Thereforre the real and imgainary values first have to be deinterleaved out of the sample message.
+The demodulator block is responsible for converting the sampled signals into ones and zeroes.
+Therefore the real and imaginary values first have to be deinterleaved out of the sample message.
 Now they are represented as integer values between 0 and 255 (8 bit).
@@
 Usually we were sampling with 250 kHz.
-This is still much more than we need assuming an additive white gausian noise (AWGN) channel with a \ensuremath{S/N = 10\,dB} (cf.
+This is still much more than we need assuming an Additive White Gaussian Noise (AWGN) channel with a \ensuremath{S/N = 10\,dB} (cf.
 equation \ref{eq:awgn}).
@@
 \ensuremath{f_{a}} is the highest frequency in the signal. 
-A FIR filter naivly implemented with floats has been used to achieve this.
+A FIR filter naively implemented with floats has been used to achieve this.
 10 kHz is the cut-off frequency of the first filter.
@@
 This filter has a cut-off frequency of 3 kHz.
-The last demodulation step is to take the signal, rectify it (cf.
+The last demodulation step is to rectify the signal (cf.
 equation \ref{eq:rectify}) and decide with a software-defined Schmitt trigger whether a sample is a 1 or a 0.
@@
 \end{equation} 
-These samples are broadcasted on the bus.
+These samples are broadcast on the bus.
 A registered listener receives the processed samples.
@@
 After the decoder received the samples it decides when a frame is sent or when the channel is idle.
-On the basis of a defined threshold the decoder determines wether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
-The baudrate and the sampling rate lead to a certain length of one bit.
-For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combinded with a baudrate of 1 kbit/s leads to 25 samples/bit.
+On the basis of a defined threshold the decoder determines whether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
+The baudrate and the sampling rate lead to a certain bit duration.
+For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combined with a baudrate of 1 kbit/s leads to 25 samples/bit.
 The decoder also includes a function to correlate the received bitstream with an expected pattern.
-Hence it is able to determine the bit ro error ratio (BER).
+Hence it is able to determine the bit error ratio (BER).
 Furthermore the code we have written so far subsumes a simulator to simulate ideal data coming out of the demodulator to test the algorithms behind the decoder.
-Also a simulator for the RTL2832U has been written to replay recorded samples, which facialiates debugging as well.
+Also a simulator for the RTL2832U has been written to replay recorded samples, which facilitates debugging as well.
 Another handy utility, which has been written in Octave, is a tool (an oscilloscope) to look at the data, which is written into files by the demodulator.

### 2017-04-07 paper.tex
Wrote a lot about signal strength variation. Lunch now.

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 % A popular package from the American Mathematical Society that provides
@@
 \section{Evaluation}
+The hard results, we collected so far are presented in this section. Of course the \textit{RTL-SDR} is the main tool combined with intelligent signal processing in \textit{Octave}.
 
 \subsection{Signal Strength Variation within the City}
+\begin{figure}[h]
+	\centering
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
+	\end{minipage}
+	\caption{Visualization of the signal processing, which is done.
+Right: Raw TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
+Left: Smoothed signal, normalized to maximum average measured in decibel with average shown as red line.}
+	\label{fig:tv_record} 
+\end{figure}
+
 To observe the signal strength variation within the city the \textit{RTL-SDR} has been used as well.
+\textit{Octave} scripts have been written for this purpose.
+These scripts are available online as well under \url{https://github.com/s3xm3x/RTLSDRspecAn}.
+Basically the scripts perform a frequency sweep through the desired band.
+All samples are written into files.
+With another script these samples, which end up in one directory, can be transfered from the the time domain into the frequency domain.
+This is done using an FFT.
+The basic problem is, that the \textit{RTL-SDR} does not have the huge sampling frequency, which is required to look at the huge band (of e.g.
+20 MHz).
+So the Nyquist-Shannon sampling theorem, which is explained also in \ref{sub:rtl2832}i, strikes again.
+The trick is, that we are no interested in real-time data.
+This allows us to only look at one part of the interesting band and sweep through everything within roughly two minutes.
+The sampling frequency used is 1.8 MHz.
+This results in a sampled band of 900 kHz.
+Hence the resulting 900 kHz bands are glued together (shifted in frequency).
+The negative frequncies of the FFT are not used at all.
+In this case they are not playing any role of improtance.
+The gain is tuned to 40.2 dB.
+To get the signal strength we simply take the average over a 10 MHz band arround the TV signal.
+This approach can be justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
+And the local TV signal provider is to its own statements using DVB-T 2.
+The signal \ensuremath{|H|} is calculated as follows.
+\begin{equation}
+	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
+\end{equation}     
+In the equation above \ensuremath{N} is the number of smaller bands which are accumulated into a big picture.
+The FFT is only carried out over the real values, which are received from the \textit{RTL-SDR}.
+\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one band.
+There are as many samples taken in one band as needed for a FFT size of x.
+In our case we chose x to be 2048.
+No window is used when glueing the data together.
+One can see, the effects of this a little bit, when taking a closer look at the plotted data.
+Rising and falling edges at the beginning of each band are visible.
+To get \ensuremath{|H_{tv,avg}|} one has to normalize everything and convert calculate the power.
+Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
+\begin{equation}
+	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
+\end{equation}   
+\ensuremath{|H_0|} is a defined value for normalization.
+This actually can be everything e.g.
+a signal with a defined power from a signal generator.
+In our case it is just the maximum value measured on a certain point.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{./fig/haversine}
+\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers).
+474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and 594 MHz belongs to a tower in Uppsala, Rickomberga.
+Distances have been calculated with the Haversine formula.}
+\label{fig:haversine}
+\end{figure}
 

### 2017-04-07 introduction.tex
Added introduction

Backscatter enables wireless communication at orders of magnitude lower
energy cost than traditional radios. A backscatter transmitter operates
by either reflecting or absorbing existing wireless signals, which only
incurrs in very low power consumption in the order of
\SI{}{\micro\watt}s~\cite{liu_ambient_2013}. As a consequence backscatter
communication is emerging as the mechanism of choice to network
battery-free devices.  Over the past few years significant progress has
been made to make backscatter communication practical for Internet of
Things~(IoT) devices. Early backscatter systems, like RFID, required a
dedicated device to generate the necessary carrier to be backscattered.
State-of-the-art systems, however, reflect ambient WiFi or television
signals~\cite{liu_ambient_2013} instead, which does away with the need for
dedicated infrastructure.

Recent ambient backscatter systems demonstrate tag-to-tag communication
by leveraging TV signals both as a source of carrier and energy. For
example, Liu et al. presents a proof-of-concept system that reflects
ambient TV signals and enables tag-to-tag communication up to a few
feet. Parks et al. further improve the communication range to tens of
feet by using analog coding~\cite{parks_turbocharging_2014}.  While these
systems can enable many applications, they are also restricted to
operate only in environments where TV signals are fairly strong~(approx
\SI{-30}{dBm}) due to the low sensitivity of the receiver used in these
tags. On the other hand, TV signals are known to vary greatly in
strength~\cite{wang_fm_2017} over space time making existing systems very
restricted in their coverage.

On the other hand, Software Defined Radios~(SDRs) are powerful devices
that have significant processing abilities, and also offer relatively
high receive sensitivity levels. The high sensitivity levels of the SDRs
compared to the typical ambient backscatter receiver tag could
significantly help improve the communication range and coverage when
receiving ambient backscatter transmissions. More importantly, SDRs
offer incomparable flexibility due to their hability to be reprogrammed.
This flexibility offers the opportunity for ample experimentation which
is welcomed in this sort of emerging technology.

In this paper we explore the following questions: Can we create an SDR
based receiver for ambient backscatter transmissions? Can we leverage
the relatively high sensitivity of such a receiver to improve range and
coverage of ambient backscatter systems? Can we provide a flexible and
low-cost experimentation platform for ambient backscatter research?

\fakepar{Contributions} The contributions of this work are twofold:
\begin{itemize}
				\item We perform a survey of the signal strength of TV
								broandcast signals.

				\item We build a low-cost SDR-based receiver capable of
								receiving data encoded in backscatterer TV signals.
\end{itemize}

### 2017-04-07 paper.tex
Added introduction
Some cleanup in the evaluation section

--- 2017-04-06 paper.tex
@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[conference]{IEEEtran}
+\usepackage{booktabs} % For formal tables
+\newenvironment{itemize*}{\begin{itemize}}{\end{itemize}}
+\usepackage[utf8x]{inputenc}
+\usepackage{subfigure}
+\usepackage{balance}
+\usepackage{comment}
+\usepackage{amsmath}
+\usepackage{xspace}
+\usepackage{caption}
+\usepackage{siunitx}
+\usepackage{booktabs}
 % Some Computer Society conferences also require the compsoc mode option,
@@
 \newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
+\newcommand{\fakeparagraph}[1]{\vspace{.5mm}\noindent\textbf{#1.}}
+\newcommand{\fakepar}[1]{\fakeparagraph{#1}}
 
@@
 % Do not put math or special symbols in the title.
-\title{Enabling Ambient Backscatter Using Low-cost Software-defined-radio}
+\title{Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio}
 
@@
 
-\hfill April 7, 2017
+\input{introduction}
 
@@
 \section{Evaluation}
+In this section we present our evaluation results. The \textit{RTL-SDR}, combined with intelligent signal processing in \textit{Octave}, are the main tools we have employed.
 
-\subsection{Signal Strength Variation within the City}
-To observe the signal strength variation within the city the \textit{RTL-SDR} has been used as well.
+\subsection{Signal Strength Variation in Space}
+\begin{figure}[h]
+	\centering
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
+	\end{minipage}
+	\caption{Spectrum of the TV signal.
+Right: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
+Left: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
+	\label{fig:tv_record} 
+\end{figure}
+
+We have used the textit{RTL-SDR} to observe the space variations of the signal strength.
+\textit{Octave} scripts have been written to aid in this purpose.
+These scripts are available online under \url{https://github.com/s3xm3x/RTLSDRspecAn}.
+The scripts perform a frequency sweep through the desired band.
+The obtained samples are then transferred from the time domain to the frequency domain using an FFT.
+
+Because the RTL-SDR has a maximum sampling rate of 1.8 MSPs, we are
+limited to a maximum simultaneous acquisition band of 900 kHz. Because
+of this, we sample the desired frequency band at 900 kHz intervals and
+stitch these bands together to form the overall frequency sweep. This
+approach is valid only because we are not interested in real-time data.
+With this approach we are able to scan a 20 MHz band in roughly two
+minutes.\cp{Please confirm these things.}
+
+During the scanning process	the gain is set to 40.2 dB. To get the
+signal strength we simply take the average over a 10 MHz band around the
+center frequency of the desired TV signal.
+This approach can be justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
+And the local TV signal provider is to its own statements using DVB-T 2.
+The signal \ensuremath{|H|} is calculated as follows.
+\begin{equation}
+	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
+\end{equation}     
+Where \ensuremath{N} is the total number of FFT bins accumulated over
+the entire frequency sweep\cp{Please confirm I got this right}. The FFT
+is only carried out over the real values, which are received from the
+\textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
+time domain of one band. There are as many samples taken in one band as
+needed for an FFT of size 2048. No window is used when stitching
+individual scans together.
+One can see, the effects of this a little bit, when taking a closer look at the plotted data.
+Rising and falling edges at the beginning of each band are visible.
+To get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power.
+Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
+\begin{equation}
+	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
+\end{equation}   
+where \ensuremath{|H_0|} is a normalization value measured at a certain
+point in space. 
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{./fig/haversine}
+\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers).
+474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and 594 MHz belongs to a tower in Uppsala, Rickomberga.
+Distances have been calculated with the Haversine formula.}
+\label{fig:haversine}
+\end{figure}
 
@@
 % references section
+\bibliographystyle{IEEEtran}
+\bibliography{paper}
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliography{paper}
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-%
-%\bibitem{github:librtlsdr}
-%Steve Markgraf et. al. \emph{librtlsdr}. \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
-%
-%\bibitem{rafael:r820t}
-%\emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
-%
-%\bibitem{book:pozar}
-%D. M. Pozar. \emph{Microwave Engineering}. Fourth Edition. John Wiley \& Sons Inc. 111 River Street, Hoboken. 2012
-%
-%\end{thebibliography}
-
-% that's all folks
 \end{document}

### 2017-04-07 paper.tex
Added another picture to show how well the data transmission works.
Finished writing. Fine tuning now.

@@
 \centering
-\includegraphics[width=\columnwidth]{./fig/frame_and_idle}
-\caption{Transmission of 101010.. with 1 kbit/s and idle state. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.}
-\label{fig:frame}
+\includegraphics[width=\columnwidth]{./fig/transmission}
+\caption{Start of a transmission of 101010..
+with 1 kbit/s with idle state before.
+Data from the Octave oscilloscope.
+Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
+This is due to a low signal strength.}
+\label{fig:transmission}
 \end{figure}
@@
 \subsection{Communication Performance}
-
+We tested the communication with different data rates from 1 bit/s up to 1 kbit/s.
+As one can see in figure \ref{fig:transmission} the signal strength is really low after backscattering.
+This leads to the problem, that the signal has a really bad quality.
+This related to the fact, that not all bits of the A/D converter are used.
+Hence a lot of quantization noise is introduced.
+In other words more gain is needed.
+The transmission shown in figure \ref{fig:transmission} was carried out with the maximum gain of 50 dB, which is available with the in \ref{sub:rtl2832u} decribed tuner.
+So we got to the limits of the \textit{RTL-SDR} at that point.
+We were able to achieve a range indoors with 1 bit/s of a couple of meteres, but when we use a higher bitrate we can only communicate over a range of a couple of decimeter and still have a bit error ratio of approximately 40 \%.
+The communiction experiment was also carried out outside closer to the tower.
+We could not figure out any improvements going closer to the tower.
 \section{Discussion}
+We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology.
+These are aspects, where fine tuning is necessary to enhance the performance:
+\begin{itemize}
+	\item Antennas: Different backscattering antennas have to be tried out to find the one with the best results.
+The question is mainly to which frequency the antenna has to be tuned (e.g.
+to the center of the TV signal or more to the edge of it).
+Moreover the standard \textit{RTL-SDR} antenna including the coaxial cable are of quite low quality.
+The antenna is e.g.
+to short.
+	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
+	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
+\end{itemize}   
+To come to a conclusion one can say, that we were able
+\begin{itemize}
+	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
+	\item to provide a heat map of tv signals in a mid-sized swedish city as well as the tools to create it.
+	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
+\end{itemize}
 

### 2017-04-07 paper.tex
Fixed some ?? errors.

@@
 Another handy utility, which has been written in Octave, is a tool (an oscilloscope) to look at the data, which is written into files by the demodulator.
-An example of a recorded data stream with 25 kS/s can be seen in figure \ref{fig:osci}.
+An example of a recorded data stream with 25 kS/s can be seen in figure \ref{fig:transmission}.
 \begin{figure}[h]
@@
 In other words more gain is needed.
-The transmission shown in figure \ref{fig:transmission} was carried out with the maximum gain of 50 dB, which is available with the in \ref{sub:rtl2832u} decribed tuner.
+The transmission shown in figure \ref{fig:transmission} was carried out with the maximum gain of 50 dB, which is available with the in \ref{sub:rtl2832} decribed tuner.
 So we got to the limits of the \textit{RTL-SDR} at that point.

### 2017-04-07 paper.tex
Added heatmap.
Worked on the RTL2832U section.

@@
 
-In this section we examine some background aspects that are essential to our
-work. \cp{Put a simple intro to each section so that they don't look empty
-before the first subsection.}
+In this background section we introduce the necessary knowledge about sofware defined radios and the used hardware, the \textit{RTL-SDR}.
 
@@
 \includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
-				\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner with integrated LNA \textit{Rafael Micro R820T/2} (right IC).}
+\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q demodulator
+\textit{Raeltek RTL2832U} (left IC) and the tuner with integrated LNA
+\textit{Rafael Micro R820T/2} (right IC).}
  \label{fig:receiver_arch}
 \end{figure}
-The Raeltek RTL2832U is a terrestrial digital video broadcast (DVB-T) demodulator IC, which is at the core of a wide range of USB-based DVB-T receiver dongles.
-One prominent feature of this chip is that it allows to retrieve the raw I/Q samples via USB.
-This is originally intended for the chip to work as a simple DAB/FM software receiver.
-Ham radio enthusiasts have combined their efforts to create a software driver (the \textit{librtlsdr}, cf.
-\cite{steve-m_librtlsdr}) that allows DVB-T dongles based on the RTL2832 to be converted into a low-cost wideband SDR receiver.
-The cost of traditional SDR's has been in the range of a few hundred to thousands of USD.
-Albeit much less powerful than the typical SDR, an RTL2832U-based DVB-T receiver can be bought with antenna for less than 10 USD.
-An RTL-SDR does not only consist of the RTL2832U.
-The second essential component is the tuner, which allows the user to select the received frequency and outputs that signal at an intermediate frequency (IF).
-The IF can then be fed into the demodulator for digitalization.
-In the case of the used hardware this tuner is a \textit{Rafael Micro R820T/2} which offers a tuning range from 42 to 1002 MHz~\cite{rafael_r820t_2011}.
-It subsumes a low-noise amplifier (LNA) and filters.
-It presents a low noise figure of 3.5 dB and features 65 dBc of image rejection.
-Figure \ref{fig:receiver_arch} \cp{We need a much better photo here.}, shows a photograph of the RTL-SDR used in our work.
+The \textit{Realtek RTL2832U} is a terrestrial digital video broadcast
+(DVB-T) demodulator IC, which is the main part of a wide range of
+USB-based DVB-T receiver dongles. One prominent feature of this chip is
+that it allows to retrieve the raw I/Q samples via USB. This is
+originally intended for the chip to work as a simple DAB/FM software
+receiver. Ham radio enthusiasts have combined their efforts to create a
+software driver (the \textit{librtlsdr}, cf. \cite{steve-m_librtlsdr})
+that allows DVB-T dongles based on the \textit{RTL2832} to be converted
+into low-cost wideband SDR receivers. The cost of traditional SDRs has
+been in the range of a few hundred to thousands of USD. Albeit much less
+powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
+can be bought with antenna for less than 10 USD.  A RTL-SDR does not
+only consist of the \textit{RTL2832U}, but also a tuner. The tuner is an
+essential component, as it allows the user to select the received
+frequency and outputs that signal at an intermediate frequency (IF). The
+IF can then be fed into the demodulator for digitalization. In the case
+of the used hardware this tuner is a \textit{Rafael Micro R820T/2} which
+offers a tuning range from 42 to 1002 MHz~\cite{rafael_r820t_2011}. It
+subsumes a low-noise amplifier (LNA) and filters. It presents a low
+noise figure of 3.5 dB and features 65 dBc of image rejection. Figure
+\ref{fig:receiver_arch}, shows a photograph of the RTL-SDR used in our
+work.
 
@@
 \end{multline}
-One can see, that the interesting in-phase part (I) in this case is represented by a DC value after mixing.
-With a low-pass filter this DC value can be separated from the undesired rest.
-An analogous procedure is done with the quadrature part (Q) when mixing with a sine.
-\cp{We need to explain this more clearly and possibly more concisely.}
+One can see, that the interesting in-phase part (I) in this case is
+represented by a DC value after mixing. With a low-pass filter this DC
+value can be separated from the undesired rest. An analogous procedure
+is done with the quadrature part (Q) when mixing with a sine.
 
@@
 \centering
-\includegraphics[width=0.8\columnwidth]{./fig/s11v2}
+\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
 \caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
@@
 	\centering
-\includegraphics[width=0.5\columnwidth]{./fig/haversine}
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/haversine}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
+	\end{minipage}
 	\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers).

### 2017-04-07 paper.tex
Worked on the transmitter part.

@@
 \end{figure}
-A backscattering antenna is the most important part of backscatter transmitter.
-Our antenna is optimized for usage with the television signal, which has its center frequency at 626 MHz.
-It is a ground plane antenna, which has been built up on a large unetched PCB plate.
-This rectangular PCB, with dimensions of 200 mm x 300 mm, is covered with a copper layer on one side and left blank on the other.
-A hole in the middle of the PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip transmission line.
-The wire, which acts as the vertical antenna element and has a length of 350 mm (roughly 0.73 \ensuremath{\lambda}), is soldered to the transmission line.
-The transmission line has been calculated for 626 MHz using an FR4 substrate with a thickness of 1.6 mm.
-According to micro strip line theory the length of the line changes the phase and the width of the line changes the impedance (cf.
-Chapter 3 in \cite{pozar_microwave_2011}).
-The design equations given in \cite{pozar_microwave_2011} are implemented in various programs e.g.
-KiCAD, which has been used for solving the design equations.
-The given parameters lead to a micro strip line width of 3 mm.
-An open-stub, attached directly at the juncture where the vertical antenna element is attached to the transmission line, is used for fine tuning of the input impedance.
-The stub length has been adjusted manually (shortened) while measuring the reflection coefficient with a vector network analyzer (VNA).
-Open-stub matching is discussed in Chapter 5 in \cite{pozar_microwave_2011}.
-Furthermore the resonance frequency has been tuned by cutting the vertical antenna element piece by piece while measuring with the VNA until \ensuremath{S_{\text{11}}} had the minimum close to the desired frequency (Figure~\ref{fig:s11}\cp{Can we replace all these figures by PDFs?}).
-In the plot one can also see, that the antenna is quite narrowband\cp{We should report something like the 3dB bandwidth rather than saying something non-specific like this}.
-For backscattering this is not necessarily a disadvantage, since everything that is within the operating frequency of the antenna is backscattered.
-With our narrowband antenna the undesired backscattering of out-of-band signals is reduced.
-With our approach we were able to achieve an input impedance at 634 MHz of \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}.
-At 626 MHz \ensuremath{S_{\text{11}}} is still approximately -17 dB.
 
-Besides the antenna, the transmitter consists of an \textit{MSP430} MCU, that sends data to the receiver.
-This transmitter controls an RF switch through an I/O pin.
-The switch can alternate the impedance connected to the antenna between matched (50 \ensuremath{\Omega}) and open.
-The software now steers this switch with a two 2 MHz rectangular timer signal to transmit a data 1 and is just leaves it open to transmit a data 0.
-Therefore the result is a frequency shift\cp{We need to explain this in the background} of the television signal by 2 MHz for a 1 and no shift for a 0.
-Hence a classical binary amplitude shift keying is the implemented modulation technique.
+A backscattering antenna is the most important part of backscatter
+transmitter. Our antenna is optimized for usage with the television
+signal, which has its center frequency at 626 MHz. It is a ground plane
+antenna, which has been built up on a large unetched PCB plate. A hole
+in the middle of the PCB serves for connecting a wire to a 50
+\ensuremath{\Omega} micro strip transmission line. The wire, which acts
+as a vertical antenna element is soldered to the transmission line. The
+transmission line has been calculated for 626 MHz using an FR4 substrate
+with a thickness of 1.6 mm. The design equations given in
+\cite{pozar_microwave_2011} are implemented in various programs e.g.
+KiCAD, which has been used for solving the design equations. An
+open-stub, attached directly at the juncture where the vertical antenna
+element is attached to the transmission line, is used for fine tuning of
+the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
+Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
+say, that the antenna works in a range from 626 MHz to 645 MHz. For
+backscattering this small bandwidth is not necessarily a disadvantage,
+since everything that is within the operating frequency of the antenna
+is backscattered. With our narrowband antenna the undesired
+backscattering of out-of-band signals is reduced.  With our approach we
+were able to achieve an input impedance at 634 MHz of
+\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
+\ensuremath{S_{\text{11}}} is still approximately -17 dB.
+
+Besides the antenna, the transmitter consists of an \textit{MSP430} MCU,
+that sends data to the receiver. This transmitter controls an RF switch
+through an I/O pin. The switch can alternate the impedance connected to
+the antenna between matched (50 \ensuremath{\Omega}) and open. The
+software now steers this switch with a two 2 MHz rectangular timer
+signal to transmit a data 1 or it just leaves the antenna open to
+transmit a data 0. Therefore the result is a frequency shift of the
+television signal by 2 MHz for a 1 and no shift for a 0. This can be
+easily understood thinking about a simple cosine, which is shifted by
+another cosine: \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)}. Of
+course this example is simplified compared to the actual situation as
+the actual signal, which is used for shifting contains more than one
+frequency as well as the TV signal, which is shifted. Hence a classical
+binary amplitude shift keying is the implemented modulation technique.
 

### 2017-04-07 paper.tex
Worked on the receiver part.

@@
 \begin{equation}
-	s_{IF}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
+	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
 \end{equation}     
@@
 \begin{multline}
-        s_{IF}(t) \cdot s_L(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+        s_{\text{IF}}(t) \cdot s_\text{L}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
 \end{multline}
@@
 \begin{multline}
-        2 \cdot s_{IF}(t) \cdot s_L(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+        2 \cdot s_{\text{IF}}(t) \cdot s_L(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
 \end{multline}
@@
 \end{figure}
-C++ has been used to implement the receiver.
-We provide the code publicly available under \url{https://github.com/s3xm3x/backscatterBASKReceiver}.
-In figure \ref{fig:receiver_arch} we show the architecture of the receiver from an abstract point of view.
-A highly sophisticated bus system has been developed to exchange data between the different components of the system.
-The signal flow is from the left to the right.
-\textit{Librtlsdr} (cf.
-\cite{steve-m_librtlsdr}), which is based on \textit{libusb}, is used to transfer the data from the TV stick into our program.
-As described in section~\ref{sub:rtl2832}, the RTL2832 mixes down the high frequency to a intermediate frequency (both values are software adjustable).
-Subsequently the data is mixed with sine and cosine, low-pass filtered and finally transferred to the digital domain.
-This data is available as two 8-bit values.
-It is a well-known fact, that this technique, which is commonly known as I/Q demodulation, results in a decomposition of the signal into in-phase and quadrature components which are generally represented as real and imaginary parts respectively.
-With this two values, the phase and absolute value of the data can be determined for every sample.
-And can be seen as:
+C++ has been used to implement the receiver. We provide the code
+publicly available under
+\url{https://github.com/s3xm3x/backscatterBASKReceiver}. A highly
+sophisticated bus system has been developed to exchange data between the
+different components of the system. The signal flow in figure
+\ref{fig:receiver_arch} is from the left to the right.
+\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
+\textit{libusb}, is used to transfer the data from the TV stick into our
+program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
+down the high frequency to a intermediate frequency (both values are
+software adjustable). Subsequently the data is mixed with sine and
+cosine, low-pass filtered and finally transferred to the digital domain.
+This data is available as two 8-bit values (real and imaginary). With
+this two values, the phase and absolute value of the data can be
+determined for every sample. It can be intepreted can be seen as:
 \begin{equation}
@@
 \end{equation}
-So the first block, which is entitled RTL2832U, provides the interface to the TV dongle.
-This block is represented as one object.
-It can set the frequency \ensuremath{f_{tuned}}, where the receiver is listening.
-Also it can set the sampling frequency \ensuremath{f_{samp}} and the analog gain of the tuner.
-The raw samples are pushed to the bus.
-The demodulator block is responsible for converting the sampled signals into ones and zeroes.
-Therefore the real and imaginary values first have to be deinterleaved out of the sample message.
-Now they are represented as integer values between 0 and 255 (8 bit).
-So one converts them to floats, whereas 127.5 equals 0.
-After that the values have to be downsampled.
-Usually we were sampling with 250 kHz.
-This is still much more than we need assuming an Additive White Gaussian Noise (AWGN) channel with a \ensuremath{S/N = 10\,dB} (cf.
-equation \ref{eq:awgn}).
+
+So the first block, which is entitled RTL2832U,
+provides the interface to the TV dongle. This block is represented as
+one object. It can set the frequency \ensuremath{f_{tuned}}, where the
+receiver is listening, besides other interesting parameters e.g. the
+analog gain.
+The demodulator block is responsible for converting the sampled signals
+into ones and zeroes. Therefore the real and imaginary values first have
+to be deinterleaved out of the sample message. Now they are represented
+as integer values between 0 and 255 (8 bit). So one converts them to
+floats, whereas 127.5 equals 0. After that the values have to be
+downsampled. Usually we were sampling with 250 kHz. This is still much
+more than we need assuming an Additive White Gaussian Noise (AWGN)
+channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
 So we decided to reduce the sampling frequency to 25 kHz.
@@
 \ensuremath{f_{a}} is the highest frequency in the signal. 
-A FIR filter naively implemented with floats has been used to achieve this.
-10 kHz is the cut-off frequency of the first filter.
-All filters have been constructed with the internal filter design tool of GNU radio.
-After downsampling another filter is used to suppress noise.
-This filter has a cut-off frequency of 3 kHz.
-The last demodulation step is to rectify the signal (cf.
-equation \ref{eq:rectify}) and decide with a software-defined Schmitt trigger whether a sample is a 1 or a 0.
+ A FIR
+filter naively implemented with floats has been used to achieve this. 10
+kHz is the cut-off frequency of the first filter. After downsampling
+another filter is used to suppress noise. This filter has a cut-off
+frequency of 3 kHz. The last demodulation step is to rectify the signal
+(cf. equation \ref{eq:rectify}) and decide with a software-defined
+Schmitt trigger whether a sample is a 1 or a 0.  
 \begin{equation}
@@
 \end{equation} 
-These samples are broadcast on the bus.
-A registered listener receives the processed samples.
-This registered listener is the decoder, which is represented as an object as well.
-After the decoder received the samples it decides when a frame is sent or when the channel is idle.
-On the basis of a defined threshold the decoder determines whether a certain sample pattern of ones and zeros is a 1 bit or a 0 bit.
-The baudrate and the sampling rate lead to a certain bit duration.
-For example a sample rate \ensuremath{f_{sampling}} of 25 kHz combined with a baudrate of 1 kbit/s leads to 25 samples/bit.
-The decoder also includes a function to correlate the received bitstream with an expected pattern.
-Hence it is able to determine the bit error ratio (BER).
-Furthermore the code we have written so far subsumes a simulator to simulate ideal data coming out of the demodulator to test the algorithms behind the decoder.
-Also a simulator for the RTL2832U has been written to replay recorded samples, which facilitates debugging as well.
-Another handy utility, which has been written in Octave, is a tool (an oscilloscope) to look at the data, which is written into files by the demodulator.
-An example of a recorded data stream with 25 kS/s can be seen in figure \ref{fig:transmission}.
-\begin{figure}[h]
+These samples are broadcast on the bus. A registered listener receives
+the processed samples. This registered listener is the decoder, which is
+represented as an object as well. After the decoder received the samples
+it decides when a frame is sent or when the channel is idle.  The
+decoder also includes a function to correlate the received bitstream
+with an expected pattern. Hence it is able to determine the bit error
+ratio (BER). Furthermore the code we have written so far subsumes
+different simulators for e.g. playing back recoreded data. Another handy
+utility, which has been written in Octave, is a tool (an oscilloscope)
+to look at the data, which is written into files by the demodulator. An
+example of a recorded data stream with 25 kS/s can be seen in figure
+\ref{fig:transmission}.  \begin{figure}[h]
 \centering

### 2017-04-07 paper.tex
Cleanup the abstract a little

@@
 \begin{abstract}
-Backscatter communication is attractive for energy-constrained devices due to its very low power requirements.
-Ambient backscatter takes this aspect to the limit by leveraging existing radio frequency signals for the purpose of communication without the need for generating energy expensive carrier signal.
-In this paper we investigate the use of ambient television broadcast signals for communication.
-As opposed to state-of-the-art restricted to operations under conditions of strong signal strength, we demonstrate a low cost software defined radio as receiver
-enables operations even in conditions when ambient signals are  weak in strength.
-We build the system using low-cost off-the-shelf microcontroller, and RTLSDR software-defined radio receiver.
-We also conduct survey of signal strength of TV broadcast in a mid-sized swedish city, and observe that our system can operate in most parts of the city.
+Backscatter communication is attractive for energy-constrained devices
+				due to its very low power requirements. Ambient backscatter
+				takes this point to the limit by leveraging existing radio
+				frequency signals for the purpose of communication, without the
+				need for generating an energy-expensive carrier signal. In this
+				paper we investigate the use of television broadcast signals as
+				a carrier for backscatter communication. As opposed to the
+				state-of-the-art, restricted to operations under conditions of
+				high signal strength, we demonstrate a low-cost software defined
+				radio receiver that operates even in conditions when ambient
+				signals are weak. We build the system using a low-cost
+				off-the-shelf microcontroller and an RTLSDR software-defined
+				radio receiver. We also conduct a survey of the signal strength
+				of TV broadcast in a mid-sized Swedish city and observe that our
+				system can operate in most parts of the city.
 \end{abstract}

### 2017-04-07 paper.tex
Worked on evaluation and discussion.

--- 2017-04-07 paper.tex
@@
 \section{Evaluation}
-In this section we present our evaluation results. The \textit{RTL-SDR}, combined with intelligent signal processing in \textit{Octave}, are the main tools we have employed.
+In this section we present our evaluation results. The \textit{RTL-SDR},
+combined with intelligent signal processing in \textit{Octave}, are the
+main tools we have employed for measurements.
 
@@
 
-We have used the textit{RTL-SDR} to observe the space variations of the signal strength.
-\textit{Octave} scripts have been written to aid in this purpose.
-These scripts are available online under \url{https://github.com/s3xm3x/RTLSDRspecAn}.
-The scripts perform a frequency sweep through the desired band.
-The obtained samples are then transferred from the time domain to the frequency domain using an FFT.
+We have used the textit{RTL-SDR} to observe the space variations of the
+signal strength. \textit{Octave} scripts have been written to aid in
+this purpose. These scripts are available online under
+\url{https://github.com/s3xm3x/RTLSDRspecAn}. The scripts perform a
+frequency sweep through the desired band. The obtained samples are then
+transferred from the time domain to the frequency domain using a fast
+fourier transformation (FFT).
 
-Because the RTL-SDR has a maximum sampling rate of 1.8 MSPs, we are
-limited to a maximum simultaneous acquisition band of 900 kHz. Because
-of this, we sample the desired frequency band at 900 kHz intervals and
-stitch these bands together to form the overall frequency sweep. This
-approach is valid only because we are not interested in real-time data.
-With this approach we are able to scan a 20 MHz band in roughly two
-minutes.\cp{Please confirm these things.}
+Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MSPs, we
+are limited to a maximum simultaneous acquisition band of 1.2 MHz. We
+are keep a safety margin to the maximum capabilities and sample the
+desired frequency band at 900 kHz intervals and stitch these bands
+together to form the overall frequency sweep. This approach is valid
+only because we are not interested in real-time data.  With this
+approach we are able to scan a 20 MHz band in roughly two minutes.
 
@@
 signal strength we simply take the average over a 10 MHz band around the
-center frequency of the desired TV signal.
-This approach can be justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
+center frequency of the desired TV signal. This approach can be
+justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
 And the local TV signal provider is to its own statements using DVB-T 2.
@@
 \end{equation}     
-Where \ensuremath{N} is the total number of FFT bins accumulated over
-the entire frequency sweep\cp{Please confirm I got this right}. The FFT
+where \ensuremath{N} is the total number of intervals accumulated over
+the entire frequency sweep. The FFT
 is only carried out over the real values, which are received from the
 \textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
-time domain of one band. There are as many samples taken in one band as
-needed for an FFT of size 2048. No window is used when stitching
-individual scans together.
-One can see, the effects of this a little bit, when taking a closer look at the plotted data.
-Rising and falling edges at the beginning of each band are visible.
-To get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power.
+time domain of one interval (one subband). There are as many samples taken in one band as
+needed for an FFT of size 2048.
+To finally get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power.
 Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
@@
 	\end{minipage}
-	\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers).
-474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and 594 MHz belongs to a tower in Uppsala, Rickomberga.
-Distances have been calculated with the Haversine formula.}
+	\caption{Left: Signal strength fading against distance relative
+to the spot where the maximum signal strength has been captured (TV
+towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
+594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
+calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
+signal in Uppsala.}
+
 \label{fig:haversine}
@@
 
-\subsection{Signal Strength Variation over Time}
-
 \subsection{Communication Performance}
-We tested the communication with different data rates from 1 bit/s up to 1 kbit/s.
-As one can see in figure \ref{fig:transmission} the signal strength is really low after backscattering.
-This leads to the problem, that the signal has a really bad quality.
-This related to the fact, that not all bits of the A/D converter are used.
-Hence a lot of quantization noise is introduced.
-In other words more gain is needed.
-The transmission shown in figure \ref{fig:transmission} was carried out with the maximum gain of 50 dB, which is available with the in \ref{sub:rtl2832} decribed tuner.
-So we got to the limits of the \textit{RTL-SDR} at that point.
-We were able to achieve a range indoors with 1 bit/s of a couple of meteres, but when we use a higher bitrate we can only communicate over a range of a couple of decimeter and still have a bit error ratio of approximately 40 \%.
-The communiction experiment was also carried out outside closer to the tower.
-We could not figure out any improvements going closer to the tower.
+We tested the communication with different data rates from 1 bit/s up to
+1 kbit/s. As one can see in figure \ref{fig:transmission} the signal
+strength is quite low after backscattering. So the problem appears, that
+not all bits of the A/D converter are used. Hence a lot of quantization
+noise is introduced. In other words more gain is needed. The
+transmission shown in figure \ref{fig:transmission} was carried out with
+the maximum gain of 50 dB, which is available with the in
+\ref{sub:rtl2832} decribed tuner. So we got to the limits of the
+\textit{RTL-SDR} at that point. We were able to achieve a range indoors
+with 1 bit/s of a couple of meteres, but when we use a higher bitrate we
+can only communicate over a range of a couple of decimeter and still
+have a bit error ratio of approximately 40 \%. The communiction
+experiment was also carried out outside closer to the tower. We could
+not figure out any improvements going closer to the tower.  
 \section{Discussion}
@@
 \begin{itemize}
-	\item Antennas: Different backscattering antennas have to be tried out to find the one with the best results.
-The question is mainly to which frequency the antenna has to be tuned (e.g.
-to the center of the TV signal or more to the edge of it).
-Moreover the standard \textit{RTL-SDR} antenna including the coaxial cable are of quite low quality.
-The antenna is e.g.
-to short.
+	\item Antennas: Different backscattering antennas have to be
+tried out to find the one with the best results. The question is mainly
+to which frequency the antenna has to be tuned (e.g. to the center of
+the TV signal or more to the edge of it). Moreover the standard
+\textit{RTL-SDR} antenna including the coaxial cable are of quite low
+quality. The antenna is e.g. too short. 
 	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
@@
 	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
-	\item to provide a heat map of tv signals in a mid-sized swedish city as well as the tools to create it.
+	\item to provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
 	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
+%\section*{Acknowledgment}
 
-The authors would like to thank...
+%The authors would like to thank...
 

### 2017-04-07 introduction.tex

@@
 incurrs in very low power consumption in the order of
-\SI{}{\micro\watt}s~\cite{liu_ambient_2013}. As a consequence backscatter
+micro-Watts~\cite{liu_ambient_2013}. As a consequence backscatter
 communication is emerging as the mechanism of choice to network
@@
 Recent ambient backscatter systems demonstrate tag-to-tag communication
-by leveraging TV signals both as a source of carrier and energy. For
-example, Liu et al. presents a proof-of-concept system that reflects
-ambient TV signals and enables tag-to-tag communication up to a few
-feet. Parks et al. further improve the communication range to tens of
-feet by using analog coding~\cite{parks_turbocharging_2014}.  While these
-systems can enable many applications, they are also restricted to
+by leveraging TV signals as a source of both carrier and energy. For
+example, Liu et al. present a proof-of-concept system that reflects
+ambient TV signals and enables tag-to-tag communication up to almost a
+meter. Parks et al. further improve the communication range to several
+meters by using analog coding~\cite{parks_turbocharging_2014}. While
+these systems can enable many applications, they are also restricted to
 operate only in environments where TV signals are fairly strong~(approx
@@
 tags. On the other hand, TV signals are known to vary greatly in
-strength~\cite{wang_fm_2017} over space time making existing systems very
-restricted in their coverage.
+strength~\cite{wang_fm_2017} over space and time, making existing
+systems very restricted in their coverage.
 
 On the other hand, Software Defined Radios~(SDRs) are powerful devices
-that have significant processing abilities, and also offer relatively
-high receive sensitivity levels. The high sensitivity levels of the SDRs
-compared to the typical ambient backscatter receiver tag could
+that have significant processing abilities and offer relatively high
+receive sensitivity levels. The high sensitivity levels of the SDRs,
+compared to the typical ambient backscatter receiver tag, could
 significantly help improve the communication range and coverage when
 receiving ambient backscatter transmissions. More importantly, SDRs
-offer incomparable flexibility due to their hability to be reprogrammed.
-This flexibility offers the opportunity for ample experimentation which
+offer incomparable flexibility due to their ability to be reprogrammed.
+This flexibility offers the opportunity for ample experimentation, which
 is welcomed in this sort of emerging technology.
 
-In this paper we explore the following questions: Can we create an SDR
-based receiver for ambient backscatter transmissions? Can we leverage
-the relatively high sensitivity of such a receiver to improve range and
-coverage of ambient backscatter systems? Can we provide a flexible and
-low-cost experimentation platform for ambient backscatter research?
+In this paper we explore the following questions: Can we create an
+SDR-based receiver for ambient backscatter transmissions? Can we
+leverage the relatively high sensitivity of such a receiver to improve
+range and coverage of ambient backscatter systems? Can we provide a
+flexible and low-cost experimentation platform for ambient backscatter
+research?
 

### 2017-04-07 backscatter.tex
Added backscatter bacground and cleaned up little

% vim: textwidth=72 spell spelllang=en

Consider an unmodulated carrier wave impinging on a backscatter tag's
antenna. The signal observed by the receiver is:
\begin{equation}
				S_r(t) = S_{rc}(t) + \sigma B(t)S_{bc}(t)
				\label{eq:signals}
\end{equation}
where $S_{rc}(t)$ is the signal coming directly from the carrier
generator and $S_{bc}(t)$ is the signal from the carrier generator that
reaches the backscatter tag. $B(t)$ is either zero or one and represents
the instantaneous state of the backscatter tag: absorbing or reflecting,
respectively.

Equation~(\ref{eq:signals}) reveals an important issue for backscatter
communication systems: self-interference. The carrier signal $S_{rc}(t)$
interferes at the receiver with the data-carrying signal from the tag,
$\sigma B(t)S_{bc}(t)$.

Recent work on generating backscatter transmissions has avoided
self-interference through \textit{frequency-shifted}
backscatter~\cite{kellogg_passive_2016,wang_fm_2017}.
The tags modulate their antenna in such a way that their transmissions
occur at a certain frequency offset from the carrier signal thus
allowing the receiver to avoid interference from the carrier by tunning
to the offset frequency where the tag is transmitting.

Consider the case when $B(t)$ periodically
alternates between its two states at a frequency $\Delta f$ while an
unmodulated carrier of frequency $f_c$ reaches the backscatter tag:

\begin{equation}
    2\sin(f_c t) \sin(\Delta f t) = \cos[(f_c + \Delta f)t] - \cos[(f_c - \Delta f) t]
    \label{eq:mixing}
\end{equation}

If we focus on only one of the two generated images, or employ single
sideband backscatter~\cite{interscatter}, the data transmission can now
be received at frequency $f_d = f_c + \Delta f$.

### 2017-04-07 paper.tex
Changed title.
Added backscatter bacground and cleaned up little

--- 2017-04-07 paper.tex
@@
 
-In this background section we introduce the necessary knowledge about sofware defined radios and the used hardware, the \textit{RTL-SDR}.
+In this section we introduce some necessary background related to our
+work.
 
-\subsection{Software Defined Radio}
-Traditionally basic radio components like mixers or filters have been realized
-in hardware. Nowadays as soon as one works in the baseband frequency the
-computational resources of a personal computer are sufficient to realize these
-components with signal processing done in software. \cp{We need to expand on
-this introduction to SDR, assuming we need it and not just assume everyone
-knows what that is.}
+%\subsection{Software Defined Radio}
+%Traditionally basic radio components like mixers or filters have been realized
+%in hardware. Nowadays as soon as one works in the baseband frequency the
+%computational resources of a personal computer are sufficient to realize these
+%components with signal processing done in software. \cp{We need to expand on
+%this introduction to SDR, assuming we need it and not just assume everyone
+%knows what that is.}
+
+\subsection{Backscatter Transmissions}
+\input{backscatter}
 
@@
 \label{sub:rtl2832}
+
 \begin{figure}[h]
@@
 				\includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
-\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q demodulator
-\textit{Raeltek RTL2832U} (left IC) and the tuner with integrated LNA
-\textit{Rafael Micro R820T/2} (right IC).}
+				\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q
+				demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner
+				with integrated LNA \textit{Rafael Micro R820T/2} (right IC).} 
 				\label{fig:receiver_arch}
@@
 powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
-can be bought with antenna for less than 10 USD.  A RTL-SDR does not
-only consist of the \textit{RTL2832U}, but also a tuner. The tuner is an
-essential component, as it allows the user to select the received
-frequency and outputs that signal at an intermediate frequency (IF). The
-IF can then be fed into the demodulator for digitalization. In the case
-of the used hardware this tuner is a \textit{Rafael Micro R820T/2} which
-offers a tuning range from 42 to 1002 MHz~\cite{rafael_r820t_2011}. It
-subsumes a low-noise amplifier (LNA) and filters. It presents a low
-noise figure of 3.5 dB and features 65 dBc of image rejection. Figure
-\ref{fig:receiver_arch}, shows a photograph of the RTL-SDR used in our
-work.
+can be bought with antenna for less than 10 USD.
+
+An RTL-SDR also contains a tuner that allows the user to select the
+received frequency.  In our case the tuner is a \textit{Rafael Micro
+R820T/2} which offers a tuning range from 42 to 1002
+MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:receiver_arch}, shows a
+photograph of the RTL-SDR used in our work.
 
@@
 \end{equation}     
-This is multiplied with a cosine (and a sine as well) from a local oscillator.
+This is multiplied with a cosine of the carrier frequency from a local oscillator.
 \begin{multline}

### 2017-04-07 paper.tex

--- 2017-04-07 paper.tex
@@
 % Do not put math or special symbols in the title.
-\title{Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio}
+\title{Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio}
 
@@
 \begin{equation}
-	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
+	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
 \end{equation}     

### 2017-04-07 paper.tex
Compressed the receiver part once more.

--- 2017-04-07 paper.tex
@@
 
-A backscattering antenna is the most important part of backscatter
-transmitter. Our antenna is optimized for usage with the television
-signal, which has its center frequency at 626 MHz. It is a ground plane
-antenna, which has been built up on a large unetched PCB plate. A hole
-in the middle of the PCB serves for connecting a wire to a 50
-\ensuremath{\Omega} micro strip transmission line. The wire, which acts
-as a vertical antenna element is soldered to the transmission line. The
-transmission line has been calculated for 626 MHz using an FR4 substrate
-with a thickness of 1.6 mm. The design equations given in
-\cite{pozar_microwave_2011} are implemented in various programs e.g.
-KiCAD, which has been used for solving the design equations. An
-open-stub, attached directly at the juncture where the vertical antenna
-element is attached to the transmission line, is used for fine tuning of
-the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
-Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
-say, that the antenna works in a range from 626 MHz to 645 MHz. For
-backscattering this small bandwidth is not necessarily a disadvantage,
-since everything that is within the operating frequency of the antenna
-is backscattered. With our narrowband antenna the undesired
-backscattering of out-of-band signals is reduced.  With our approach we
-were able to achieve an input impedance at 634 MHz of
+Our backscattering antenna is optimized for usage with the television signal,
+which has its center frequency at 626 MHz. It is a ground plane antenna, which
+has been built up on a large unetched PCB plate. A hole in the middle of the
+PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip
+transmission line. The wire, which acts as a vertical antenna element is
+soldered to the transmission line. The microstrip transmission line design
+equations given in \cite{pozar_microwave_2011} are implemented in various
+programs e.g.  KiCAD, which has been used for designing them. An open-stub,
+attached directly at the juncture where the vertical antenna element is
+attached to the transmission line, is used for fine tuning of the input
+impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).  Accepting -15 dB as
+sufficient value for \ensuremath{S_{11}}, one can say, that the antenna works
+within a range from 626 MHz to 645 MHz. With our narrowband antenna the
+undesired backscattering of out-of-band signals is reduced. With our approach
+we were able to achieve an input impedance at 634 MHz of
 \ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
@@
 \end{figure}
-C++ has been used to implement the receiver. We provide the code
-publicly available under
-\url{https://github.com/s3xm3x/backscatterBASKReceiver}. A highly
-sophisticated bus system has been developed to exchange data between the
-different components of the system. The signal flow in figure
-\ref{fig:receiver_arch} is from the left to the right.
-\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
-\textit{libusb}, is used to transfer the data from the TV stick into our
-program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
-down the high frequency to a intermediate frequency (both values are
-software adjustable). Subsequently the data is mixed with sine and
-cosine, low-pass filtered and finally transferred to the digital domain.
-This data is available as two 8-bit values (real and imaginary). With
-this two values, the phase and absolute value of the data can be
-determined for every sample. It can be intepreted can be seen as:
+We provide our C++ receiver code publicly available under
+\url{https://github.com/s3xm3x/backscatterBASKReceiver}. A highly sophisticated
+bus system has been developed to exchange data between the different components
+of the system. The signal flow in figure \ref{fig:receiver_arch} is from the
+left to the right.  \textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is
+based on \textit{libusb}, is used to transfer the data from the TV stick into
+our program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes down
+the high frequency to a intermediate frequency (both values are software
+adjustable). The sampled data is available as two 8-bit values (real and
+imaginary). With this two values, the phase and absolute value of the data can
+be determined for every sample. Every sample can be intepreted as:
 \begin{equation}
@@
 
-So the first block, which is entitled RTL2832U,
-provides the interface to the TV dongle. This block is represented as
-one object. It can set the frequency \ensuremath{f_{tuned}}, where the
-receiver is listening, besides other interesting parameters e.g. the
-analog gain.
-The demodulator block is responsible for converting the sampled signals
-into ones and zeroes. Therefore the real and imaginary values first have
-to be deinterleaved out of the sample message. Now they are represented
-as integer values between 0 and 255 (8 bit). So one converts them to
-floats, whereas 127.5 equals 0. After that the values have to be
-downsampled. Usually we were sampling with 250 kHz. This is still much
-more than we need assuming an Additive White Gaussian Noise (AWGN)
-channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
-So we decided to reduce the sampling frequency to 25 kHz.
+So the first block, which is entitled RTL2832U, provides the interface to the
+TV dongle. This block controls the frequency \ensuremath{f_{tuned}}, where the
+receiver is listening, besides other interesting parameters e.g. the analog
+gain.
+
+The demodulator block is responsible for converting the sampled signals into
+ones and zeroes. Therefore the real and imaginary values first have to be
+deinterleaved out of the sample message. These 8 bit integers are converted
+into floats, whereas 127.5 equals 0. After that the values have to be
+downsampled. Usually we were sampling with 250 kHz. This is still much more
+than we need assuming an Additive White Gaussian Noise (AWGN) channel with a
+\ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).  So we decided to
+reduce the sampling frequency to 25 kHz.
+
 \begin{equation}
@@
 \ensuremath{f_{a}} is the highest frequency in the signal. 
- A FIR
-filter naively implemented with floats has been used to achieve this. 10
-kHz is the cut-off frequency of the first filter. After downsampling
-another filter is used to suppress noise. This filter has a cut-off
-frequency of 3 kHz. The last demodulation step is to rectify the signal
-(cf. equation \ref{eq:rectify}) and decide with a software-defined
-Schmitt trigger whether a sample is a 1 or a 0.  
+ A FIR filter
+naively implemented with floats has been used to achieve this.  After
+downsampling another filter is used to suppress noise.  The last demodulation
+step is to rectify the signal (cf. equation \ref{eq:rectify}) and decide with a
+software-defined Schmitt trigger whether a sample is a 1 or a 0.  
 \begin{equation}
@@
 \end{equation} 
-These samples are broadcast on the bus. A registered listener receives
-the processed samples. This registered listener is the decoder, which is
-represented as an object as well. After the decoder received the samples
-it decides when a frame is sent or when the channel is idle.  The
-decoder also includes a function to correlate the received bitstream
-with an expected pattern. Hence it is able to determine the bit error
-ratio (BER). Furthermore the code we have written so far subsumes
-different simulators for e.g. playing back recoreded data. Another handy
-utility, which has been written in Octave, is a tool (an oscilloscope)
-to look at the data, which is written into files by the demodulator. An
-example of a recorded data stream with 25 kS/s can be seen in figure
-\ref{fig:transmission}.  \begin{figure}[h]
+A registered listener receives the processed samples, which are broadcasted by
+the demodulator. This registered listener is the decoder. After the decoder
+received the samples it decides when a frame is sent or when the channel is
+idle.  The decoder also includes a function to correlate the received bitstream
+with an expected pattern. Hence it is able to determine the bit error ratio
+(BER). Furthermore the code we have written so far subsumes different
+simulators for e.g. playing back recoreded data. Another handy utility, which
+has been written in Octave, is a tool (an oscilloscope) to look at the data,
+which is written into files by the demodulator (cf. figure \ref{fig:transmission}).
+\begin{figure}[h]
+
 \centering
@@
 Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
-This is due to a low signal strength.}
+This is due to a low signal strength.
+Sampling frequency is 25 kHz.}
 \label{fig:transmission}
@@
 
-We have used the textit{RTL-SDR} to observe the space variations of the
+We have used the \textit{RTL-SDR} to observe the space variations of the
 signal strength. \textit{Octave} scripts have been written to aid in

### 2017-04-07 paper.tex
Changed some subscripts to non-italic. Always use \text{}.

@@
 \begin{multline}
-        s_{\text{IF}}(t) \cdot s_\text{L}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+        s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
 \end{multline}
@@
 \begin{multline}
-        2 \cdot s_{\text{IF}}(t) \cdot s_L(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+        2 \cdot s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
 \end{multline}
@@
 \includegraphics[width=0.6\columnwidth]{./fig/s11v2}
-\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
+\caption{Input reflection coefficient \ensuremath{\Gamma_{\text{in}}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
 \label{fig:s11}
@@
 we were able to achieve an input impedance at 634 MHz of
-\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
+\ensuremath{Z_{\text{in}}=(52.2 + j 4.3) \Omega}. At 626 MHz
 \ensuremath{S_{\text{11}}} is still approximately -17 dB.
@@
 	\label{eq:nyquist}
-	f_{samp} \geq 2 \cdot f_{a}
+	f_{\text{samp}} \geq 2 \cdot f_{\text{a}}
 \end{equation}
-\ensuremath{f_{a}} is the highest frequency in the signal. 
+\ensuremath{f_{\text{a}}} is the highest frequency in the signal. 
  A FIR filter

### 2017-04-07 paper.tex
Some minor changes.

@@
 
-Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MSPs, we
-are limited to a maximum simultaneous acquisition band of 1.2 MHz. We
-are keep a safety margin to the maximum capabilities and sample the
-desired frequency band at 900 kHz intervals and stitch these bands
-together to form the overall frequency sweep. This approach is valid
-only because we are not interested in real-time data.  With this
-approach we are able to scan a 20 MHz band in roughly two minutes.
+Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MS/s, we are
+limited to a maximum simultaneous acquisition band of 1.2 MHz. We keep a safety
+margin to the maximum capabilities and sample the desired frequency band at 900
+kHz intervals and stitch these bands together to form the overall frequency
+sweep. This approach is valid only because we are not interested in real-time
+data.  With this approach we are able to scan a 20 MHz band in roughly 30
+seconds.
 
@@
 	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
-\end{equation}     
-where \ensuremath{N} is the total number of intervals accumulated over
-the entire frequency sweep. The FFT
-is only carried out over the real values, which are received from the
-\textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
-time domain of one interval (one subband). There are as many samples taken in one band as
-needed for an FFT of size 2048.
-To finally get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power.
-Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
+\end{equation}     where \ensuremath{N} is the total number of intervals
+accumulated over the entire frequency sweep. The FFT is only carried out over
+the real values, which are received from the \textit{RTL-SDR}.
+\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one
+interval (one subband). There are as many samples taken in one band as needed
+for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
+normalize everything and calculate the power. Before doing that a simple
+smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
 \begin{equation}
@@
 \end{equation}   
-where \ensuremath{|H_0|} is a normalization value measured at a certain
-point in space. 
+where \ensuremath{|H_0|} is a normalization value measured at a certain point
+in space. The result of combining measured signal strength values with the
+haversine function to calculate the distance to the point where the maximum
+signal has been captured can be observed in figure \ref{fig:haversine}.
 \begin{figure}[h]

### 2017-04-07 paper.tex
Aggressive reduction of design section

--- 2017-04-07 paper.tex
@@
 
+In this section we describe the most important aspects of our design.
+
 \subsection{Transmitter}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
-\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
-\label{fig:s11}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
+%\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
+%\label{fig:s11}
+%\end{figure}
 
-Our backscattering antenna is optimized for usage with the television signal,
-which has its center frequency at 626 MHz. It is a ground plane antenna, which
-has been built up on a large unetched PCB plate. A hole in the middle of the
-PCB serves for connecting a wire to a 50 \ensuremath{\Omega} micro strip
-transmission line. The wire, which acts as a vertical antenna element is
-soldered to the transmission line. The microstrip transmission line design
-equations given in \cite{pozar_microwave_2011} are implemented in various
-programs e.g.  KiCAD, which has been used for designing them. An open-stub,
-attached directly at the juncture where the vertical antenna element is
-attached to the transmission line, is used for fine tuning of the input
-impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).  Accepting -15 dB as
-sufficient value for \ensuremath{S_{11}}, one can say, that the antenna works
-within a range from 626 MHz to 645 MHz. With our narrowband antenna the
-undesired backscattering of out-of-band signals is reduced. With our approach
-we were able to achieve an input impedance at 634 MHz of
-\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
-\ensuremath{S_{\text{11}}} is still approximately -17 dB.
+%A backscattering antenna is the most important part of our backscatter
+%transmitter. 
+Our backscatter transmitter is designed to be used with the television
+signal, which has its center frequency at 626 MHz. To that end, we build
+an
+antenna using an unetched PCB plate as the ground plane and a wire that acts
+as the vertical antenna element.% is soldered to the transmission line. The
+%transmission line has been calculated for 626 MHz using an FR4 substrate
+%with a thickness of 1.6 mm. The design equations given in
+%\cite{pozar_microwave_2011} are implemented in various programs e.g.
+%KiCAD, which has been used for solving the design equations. An
+%open-stub, attached directly at the juncture where the vertical antenna
+%element is attached to the transmission line, is used for fine tuning of
+%the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
+%Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
+%say, that the antenna works in a range from 626 MHz to 645 MHz. For
+%backscattering this small bandwidth is not necessarily a disadvantage,
+%since everything that is within the operating frequency of the antenna
+%is backscattered. With our narrowband antenna the undesired
+%backscattering of out-of-band signals is reduced.  With our approach we
+%were able to achieve an input impedance at 634 MHz of
+%\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
+%\ensuremath{S_{\text{11}}} is still approximately -17 dB.
+This antenna works well in the range from 626 to 245 MHz.
 
@@
 transmit a data 0. Therefore the result is a frequency shift of the
-television signal by 2 MHz for a 1 and no shift for a 0. This can be
-easily understood thinking about a simple cosine, which is shifted by
-another cosine: \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)}. Of
-course this example is simplified compared to the actual situation as
-the actual signal, which is used for shifting contains more than one
-frequency as well as the TV signal, which is shifted. Hence a classical
+television signal by 2 MHz for a 1 and no shift for a 0. Hence a classical
 binary amplitude shift keying is the implemented modulation technique.
@@
 \end{figure}
-We provide our C++ receiver code publicly available under
-\url{https://github.com/s3xm3x/backscatterBASKReceiver}. A highly sophisticated
-bus system has been developed to exchange data between the different components
-of the system. The signal flow in figure \ref{fig:receiver_arch} is from the
-left to the right.  \textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is
-based on \textit{libusb}, is used to transfer the data from the TV stick into
-our program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes down
-the high frequency to a intermediate frequency (both values are software
-adjustable). The sampled data is available as two 8-bit values (real and
-imaginary). With this two values, the phase and absolute value of the data can
-be determined for every sample. Every sample can be intepreted as:
+C++ has been used to implement the receiver. We provide the code
+publicly available under
+\url{https://github.com/s3xm3x/backscatterBASKReceiver}. 
+%A highly
+%sophisticated bus system has been developed to exchange data between the
+%different components of the system. 
+Figure
+\ref{fig:receiver_arch} shows the architecture of our receiver, the
+signal flows from the left to the right.
+%\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
+%\textit{libusb}, is used to transfer the data from the TV stick into our
+%program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
+%down the high frequency to a intermediate frequency (both values are
+%software adjustable). Subsequently the data is mixed with sine and
+%cosine, low-pass filtered and finally transferred to the digital domain.
+The raw I/Q data is available as two 8-bit values (real and imaginary)
+from the RTL-SDR. With
+these two values, the phase and magnitude of the data can be
+determined for every sample. The received signal can be represented as:
 \begin{equation}
@@
 
-So the first block, which is entitled RTL2832U, provides the interface to the
-TV dongle. This block controls the frequency \ensuremath{f_{tuned}}, where the
-receiver is listening, besides other interesting parameters e.g. the analog
-gain.
+So the first block, entitled RTL2832U, provides the interface to the TV
+dongle. It can set the frequency \ensuremath{f_{tuned}}, where the
+receiver is listening, and the analog gain.
 
-The demodulator block is responsible for converting the sampled signals into
-ones and zeroes. Therefore the real and imaginary values first have to be
-deinterleaved out of the sample message. These 8 bit integers are converted
-into floats, whereas 127.5 equals 0. After that the values have to be
-downsampled. Usually we were sampling with 250 kHz. This is still much more
-than we need assuming an Additive White Gaussian Noise (AWGN) channel with a
-\ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).  So we decided to
-reduce the sampling frequency to 25 kHz.
-
-\begin{equation}
-	\label{eq:awgn}
-	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
-\end{equation}     
-To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
-\begin{equation}
-	\label{eq:nyquist}
-	f_{samp} \geq 2 \cdot f_{a}
-\end{equation}
-\ensuremath{f_{a}} is the highest frequency in the signal. 
- A FIR filter
-naively implemented with floats has been used to achieve this.  After
-downsampling another filter is used to suppress noise.  The last demodulation
-step is to rectify the signal (cf. equation \ref{eq:rectify}) and decide with a
+The demodulator block is responsible for converting the sampled signals
+into ones and zeroes. 
+%Therefore the real and imaginary values first have
+%to be deinterleaved out of the sample message. Now they are represented
+%as integer values between 0 and 255 (8 bit). So one converts them to
+%floats, whereas 127.5 equals 0. After that the values have to be
+%downsampled. Usually we were sampling with 250 kHz. This is still much
+%more than we need assuming an Additive White Gaussian Noise (AWGN)
+%channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
+%So we decided to reduce the sampling frequency to 25 kHz.
+%\begin{equation}
+%	\label{eq:awgn}
+%	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
+%\end{equation}     
+%To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
+%\begin{equation}
+%	\label{eq:nyquist}
+%	f_{samp} \geq 2 \cdot f_{a}
+%\end{equation}
+%\ensuremath{f_{a}} is the highest frequency in the signal. 
+ A FIR
+%filter naively implemented with floats has been used to achieve this. 10
+%kHz is the cut-off frequency of the first filter. After downsampling
+%another filter is used to suppress noise. This filter has a cut-off
+%frequency of 3 kHz. 
+As part of the demodulation process the signal is down-sampled and
+low-pass-filtered two times. The last demodulation step is to rectify
+the signal (cf. equation \ref{eq:rectify}) and decide with a
 software-defined Schmitt trigger whether a sample is a 1 or a 0.  
@@
 \end{equation}
-A registered listener receives the processed samples, which are broadcasted by
-the demodulator. This registered listener is the decoder. After the decoder
-received the samples it decides when a frame is sent or when the channel is
-idle.  The decoder also includes a function to correlate the received bitstream
-with an expected pattern. Hence it is able to determine the bit error ratio
-(BER). Furthermore the code we have written so far subsumes different
-simulators for e.g. playing back recoreded data. Another handy utility, which
-has been written in Octave, is a tool (an oscilloscope) to look at the data,
-which is written into files by the demodulator (cf. figure \ref{fig:transmission}).
-\begin{figure}[h]
 
+The resulting values are sent through the bus to the decoder. After the
+decoder received the samples it decides when a frame is sent or when the
+channel is idle.  
+The decoder also includes a function to correlate the
+received bitstream with an expected pattern. Hence it is able to
+determine the bit error ratio (BER). 
+Furthermore the code we have
+written so far subsumes different simulators for e.g. playing back
+recoreded data. Another handy utility, which has been written in Octave,
+is a tool (an oscilloscope) to look at the data, which is written into
+files by the demodulator. An example of a recorded data stream with 25
+kS/s can be seen in figure
+\ref{fig:transmission}.  \begin{figure}[h]
 \centering
@@
 Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
-This is due to a low signal strength.
-Sampling frequency is 25 kHz.}
+This is due to a low signal strength.}
 \label{fig:transmission}
@@
 
-We have used the \textit{RTL-SDR} to observe the space variations of the
+We have used the textit{RTL-SDR} to observe the space variations of the
 signal strength. \textit{Octave} scripts have been written to aid in

### 2017-04-07 paper.tex
Removed Master Programme ...

--- 2017-04-07 paper.tex
@@
 \author{\IEEEauthorblockN{Maximilian Stiefel}
-\IEEEauthorblockA{Master Programme Embdedded Systems\\
-Uppsala University\\
+\IEEEauthorblockA{Uppsala University\\
 maximilian.stiefel.8233@student.uu.se}
@@
 \IEEEauthorblockN{Elmar van Rijnswou}
-\IEEEauthorblockA{Master Programme Embdedded Systems\\
-Uppsala University\\
+\IEEEauthorblockA{Uppsala University\\
 elmar.vanrijnswou.9818@student.uu.se}
@@
 \IEEEauthorblockA{Uppsala University\\
-carlos.penichet@it.uu.se}
+carlos.penichet@it.uu.se                              }
 \and

### 2017-04-07 paper.tex

--- 2017-04-07 paper.tex
@@
 \end{figure}
-C++ has been used to implement the receiver. We provide the code
-publicly available under
-\url{https://github.com/s3xm3x/backscatterBASKReceiver}. 
+We provide our C++ receiver code publicly available under
+\url{https://github.com/s3xm3x/backscatterBASKReceiver}
 %A highly
@@
 kS/s can be seen in figure
-\ref{fig:transmission}.  \begin{figure}[h]
+\ref{fig:transmission}.  
+\begin{figure}[h]
 \centering

### 2017-04-07 paper.tex
Found another italic subscript problem.
Minor changes of the design section.

@@
 %\ensuremath{S_{\text{11}}} is still approximately -17 dB.
-This antenna works well in the range from 626 to 245 MHz.
+This antenna works well in the range from 626 to 245 MHz, which has been
+measured with a vector network analyzer (VNA).
 
@@
 %different components of the system. 
-Figure
-\ref{fig:receiver_arch} shows the architecture of our receiver, the
+Figure \ref{fig:receiver_arch} shows the architecture of our receiver, the
 signal flows from the left to the right.
@@
 %cosine, low-pass filtered and finally transferred to the digital domain.
-The raw I/Q data is available as two 8-bit values (real and imaginary)
-from the RTL-SDR. With
-these two values, the phase and magnitude of the data can be
+The raw I/Q data is available as two 8-bit values (real and imaginary) from the
+RTL-SDR. With these two values, the phase and magnitude of the data can be
 determined for every sample. The received signal can be represented as:
@@
 So the first block, entitled RTL2832U, provides the interface to the TV
-dongle. It can set the frequency \ensuremath{f_{tuned}}, where the
-receiver is listening, and the analog gain.
+dongle. It controls the frequency \ensuremath{f_{tuned}}, where the
+receiver is listening as well as other interesting parameters e.g. the analog gain.
 
@@
 
-The resulting values are sent through the bus to the decoder. After the
-decoder received the samples it decides when a frame is sent or when the
-channel is idle.  
-The decoder also includes a function to correlate the
-received bitstream with an expected pattern. Hence it is able to
-determine the bit error ratio (BER). 
-Furthermore the code we have
-written so far subsumes different simulators for e.g. playing back
-recoreded data. Another handy utility, which has been written in Octave,
-is a tool (an oscilloscope) to look at the data, which is written into
-files by the demodulator. An example of a recorded data stream with 25
-kS/s can be seen in figure
-\ref{fig:transmission}.  
+The resulting values are sent through the bus to the decoder. After the decoder
+received the samples it decides when a frame is sent or when the channel is
+idle. The decoder also includes a function to correlate the received bitstream
+with an expected pattern. Hence it is able to determine the bit error ratio
+(BER). Furthermore the code we have written so far subsumes different
+simulators for e.g. playing back recoreded data. Another handy utility, which
+has been written in Octave, is a tool (an oscilloscope) to look at the data,
+which is written into files by the demodulator. An example of a recorded data
+stream with 25 kS/s can be seen in figure \ref{fig:transmission}.  
 \begin{figure}[h]
@@
 \begin{equation}
-	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
+	|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
 \end{equation}   

### 2017-04-07 paper.tex
Fixed discussion.

@@
 We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology.
-These are aspects, where fine tuning is necessary to enhance the performance:
-\begin{itemize}
-	\item Antennas: Different backscattering antennas have to be
-tried out to find the one with the best results. The question is mainly
-to which frequency the antenna has to be tuned (e.g. to the center of
-the TV signal or more to the edge of it). Moreover the standard
-\textit{RTL-SDR} antenna including the coaxial cable are of quite low
-quality. The antenna is e.g. too short. 
-	\item Receiver: Some fine tuning can be made when it comes to the filter coefficients of the different filters.
-	\item Frame: The frame structure can be optimized to be able to correct errors (e.g. Hamming code).  
-\end{itemize}   
-To come to a conclusion one can say, that we were able
-\begin{itemize}
-	\item to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
-	\item to provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
-	\item to show for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
-\end{itemize}
+Different backscattering antennas have to be tried out to find the one with the best results.
+The question is mainly to which frequency the antenna has to be tuned (e.g.
+to the center of the TV signal or more to the edge of it).
+Moreover the standard \textit{RTL-SDR} antenna including the coaxial cable are of quite low quality.
+The antenna is e.g.
+too short.
+Hence a custom antenna would be benificial at this point as well.
+Moreover some fine tuning can be made when it comes to the filter coefficients of the different filters.
+Finally the frame structure can be optimized to be able to correct errors (e.g.
+Hamming code).
+
+To come to a conclusion one can say, that we were able to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
+Therupon we could provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
+And last but not least we have shown for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
 

### 2017-04-07 paper.tex
Fixed figure 4.
A little cleanup in the results section

--- 2017-04-07 paper.tex
@@
 justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
-And the local TV signal provider is to its own statements using DVB-T 2.
+And the local TV provider advertises to be transmitting DVB-T 2.
 The signal \ensuremath{|H|} is calculated as follows.  
@@
 for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
-normalize everything and calculate the power. Before doing that a simple
+normalize everything and calculate the power. Before doing that, a simple
 smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
@@
 We tested the communication with different data rates from 1 bit/s up to
-1 kbit/s. As one can see in figure \ref{fig:transmission} the signal
-strength is quite low after backscattering. So the problem appears, that
-not all bits of the A/D converter are used. Hence a lot of quantization
-noise is introduced. In other words more gain is needed. The
+1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
+strength from the backscatter transmitter is rather low. This leads to a
+significant amount of quantization noise to appear. This situation can
+be improved by increasing the analog gain of the receiver. The
 transmission shown in figure \ref{fig:transmission} was carried out with
-the maximum gain of 50 dB, which is available with the in
-\ref{sub:rtl2832} decribed tuner. So we got to the limits of the
-\textit{RTL-SDR} at that point. We were able to achieve a range indoors
-with 1 bit/s of a couple of meteres, but when we use a higher bitrate we
-can only communicate over a range of a couple of decimeter and still
-have a bit error ratio of approximately 40 \%. The communiction
-experiment was also carried out outside closer to the tower. We could
-not figure out any improvements going closer to the tower.  
+the maximum gain (50 dB) available to our receiver.
+
+With a data rate of 1 bit/s, we were able to achieve a range indoors
+of a couple of meters. With higher bitrates we
+can only communicate over a range of a few decimeters and the 
+bit error rate is still approximately 40 \%. The communication
+experiment was also carried out outside, close to the TV tower. We could
+not observe any improvements going closer to the tower.
+
 \section{Discussion}

### 2017-04-07 paper.tex

@@
 	\caption{Spectrum of the TV signal.
-Right: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
+Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
 Left: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}

### 2017-04-07 paper.tex
Fixed links to Github.

--- 2017-04-07 paper.tex
@@
 We provide our C++ receiver code publicly available under
-\url{https://github.com/s3xm3x/backscatterBASKReceiver}
+\cite{s3xm3x_backscatterBASKReceiver}
 %A highly
@@
 	\caption{Spectrum of the TV signal.
-Right: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
-Left: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
+Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
+Right: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
 	\label{fig:tv_record} 
@@
 signal strength. \textit{Octave} scripts have been written to aid in
-this purpose. These scripts are available online under
-\url{https://github.com/s3xm3x/RTLSDRspecAn}. The scripts perform a
+this purpose. These scripts are available under \cite{s3xm3x_RTLSDRSpecAn}. The scripts perform a
 frequency sweep through the desired band. The obtained samples are then

### 2017-04-07 paper.tex

--- 2017-04-07 paper.tex
@@
 Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
-Left: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
+Right: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
 	\label{fig:tv_record} 

### 2017-04-07 paper.tex
A little of cleanup in the discussion

--- 2017-04-07 paper.tex
@@
 \section{Discussion}
-We are aware of the fact, that our communication system has to be improved until it is able to be used in allday technology.
-Different backscattering antennas have to be tried out to find the one with the best results.
-The question is mainly to which frequency the antenna has to be tuned (e.g.
-to the center of the TV signal or more to the edge of it).
-Moreover the standard \textit{RTL-SDR} antenna including the coaxial cable are of quite low quality.
-The antenna is e.g.
-too short.
-Hence a custom antenna would be benificial at this point as well.
-Moreover some fine tuning can be made when it comes to the filter coefficients of the different filters.
-Finally the frame structure can be optimized to be able to correct errors (e.g.
-Hamming code).
 
-To come to a conclusion one can say, that we were able to show, that spectrum scanning over a huge band with sweeping is possible using the \textit{RTL-SDR}.
-Therupon we could provide a heat map of TV signals in a mid-sized swedish city as well as the tools to create it.
-And last but not least we have shown for the first time, that backscatter communication is possible with the \textit{RTL-SDR}.
+Our system still has a lot of room for improvement.  Different
+backscattering antennas could be tried to find the one with the best
+combination of gain, tuning frequency and bandwidth. Moreover the
+standard \textit{RTL-SDR} antenna and its coaxial cable are of low
+quality.  Hence a custom antenna would be beneficial. Furthermore some
+fine tuning can be made when it comes to the filter coefficients of the
+different filters in the demodulator chain. Finally  errors can be
+further reduced by using one of many available forward error correcting
+codes.
+
+To come to a conclusion one can say, that we were able to show, that
+spectrum scanning over a huge band with sweeping is possible using the
+\textit{RTL-SDR}.  This allowed us to perform a survey of the TV signal
+strength in the city. And finally we have shown for the first time, that
+backscatter communication is possible with the \textit{RTL-SDR}.
 

### 2017-04-07 paper.tex
Balance columns in the last page
Fixed affiliation of Thiemo.
Improved 1st paragraph of the receiver

@@
 \end{figure}
-We provide our C++ receiver code publicly available under
-\cite{s3xm3x_backscatterBASKReceiver}
+The receiver consists of three main modules interconnected by a bus. The
+RTL-SDR block is in charge of collecting samples from the SDR-device. The
+demodulator turns the sampled signal into a streams of ones and zeroes.
+Finally, the Decoder block is in charge of detecting frames. We provide our C++
+receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.
+
 %A highly
@@
 
+\balance
+
 \subsection{Communication Performance}

### 2017-04-07 paper.tex
Capitalized figure references

@@
 \IEEEauthorblockA{Uppsala University\\
-and SICS Swedish ICT\\
+and RISE SICS\\
 thiemo@sics.se}}
@@
 which is written into files by the demodulator. An example of a recorded data
-stream with 25 kS/s can be seen in figure \ref{fig:transmission}.  
+stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.  
 \begin{figure}[h]
@@
 haversine function to calculate the distance to the point where the maximum
-signal has been captured can be observed in figure \ref{fig:haversine}.
+signal has been captured can be observed in Figure \ref{fig:haversine}.
 \begin{figure}[h]
@@
 be improved by increasing the analog gain of the receiver. The
-transmission shown in figure \ref{fig:transmission} was carried out with
+transmission shown in Figure \ref{fig:transmission} was carried out with
 the maximum gain (50 dB) available to our receiver.

### 2017-04-07 paper.tex
Seems like its done.

--- 2017-04-07 paper.tex
@@
 bit error rate is still approximately 40 \%. The communication
-experiment was also carried out outside, close to the TV tower. We could
-not observe any improvements going closer to the tower.
+experiment was also carried out outside, where we got similar results.
 

### 2017-05-14 abstract.tex
updated abstract

Backscatter communication enables ultra-low
power wireless transmissions, and is attractive
for devices operating on harvested energy. Ambient backscatter
takes the concept further forward, by leveraging
ambient wireless signals like
television signals as both source of power
and carrier signal. Existing state-of-the-art
ambient backscatter systems demonstrate ability to 
achieve tag-to-tag communication under conditions when
signal strength of ambient signal is sufficiently
strong.  In this paper, we demonstrate that it is possible
to backscatter even weak signals, when received using
a low cost software defined radio receiver. Our initial results 
demonstrate that the strength of ambient television signals is good
enough in most parts of a mid-sized swedish city for the system
to operate, and also communicate.

### 2017-05-14 paper.tex
Removed reference to tower proximity
Fixed some typos.
updated abstract

--- 2017-04-07 paper.tex
@@
 \usepackage{booktabs}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
 
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
 \usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
 
@@
 \begin{abstract}
-Backscatter communication is attractive for energy-constrained devices
-				due to its very low power requirements. Ambient backscatter
-				takes this point to the limit by leveraging existing radio
-				frequency signals for the purpose of communication, without the
-				need for generating an energy-expensive carrier signal. In this
-				paper we investigate the use of television broadcast signals as
-				a carrier for backscatter communication. As opposed to the
-				state-of-the-art, restricted to operations under conditions of
-				high signal strength, we demonstrate a low-cost software defined
-				radio receiver that operates even in conditions when ambient
-				signals are weak. We build the system using a low-cost
-				off-the-shelf microcontroller and an RTLSDR software-defined
-				radio receiver. We also conduct a survey of the signal strength
-				of TV broadcast in a mid-sized Swedish city and observe that our
-				system can operate in most parts of the city.
+\input{abstract}
+
 \end{abstract}
@@
 %\end{equation}
-%\ensuremath{f_{a}} is the highest frequency in the signal. 
- A FIR
+%\ensuremath{f_{a}} is the highest frequency in the signal. A FIR
 %filter naively implemented with floats has been used to achieve this. 10
@@
 (BER). Furthermore the code we have written so far subsumes different
-simulators for e.g. playing back recoreded data. Another handy utility, which
+simulators for e.g. playing back recorded data. Another handy utility, which
 has been written in Octave, is a tool (an oscilloscope) to look at the data,

### 2017-05-14 abstract.tex
updated abstract

@@
 power wireless transmissions, and is attractive
-for devices operating on harvested energy. Ambient backscatter
-takes the concept further forward, by leveraging
+for networking devices which operate 
+on harvested energy. Ambient backscatter
+takes the concept further, by leveraging
 ambient wireless signals like
-television signals as both source of power
-and carrier signal. Existing state-of-the-art
-ambient backscatter systems demonstrate ability to 
-achieve tag-to-tag communication under conditions when
-signal strength of ambient signal is sufficiently
-strong.  In this paper, we demonstrate that it is possible
-to backscatter even weak signals, when received using
-a low cost software defined radio receiver. Our initial results 
-demonstrate that the strength of ambient television signals is good
-enough in most parts of a mid-sized swedish city for the system
-to operate, and also communicate.
+television signals, as both the source of power
+and carrier signal. Existing state-of-the-art ambient backscatter
+systems demonstrate ability to achieve tag-to-tag communication 
+under conditions when tags are close to television tower, i.e., with the
+strength of TV signals  sufficiently high (-30 dBm).  In this paper, we present our preliminary work
+which demonstrates the possibility to backscatter and communicate using signficiantly weaker 
+ambient television signals. The key to using weaker broadcast television signals, is to leverage a low-cost
+software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
+demonstrate that the strength of television signals is strong 
+enough in most parts of a mid-sized swedish city for our system
+to operate, and communicate, a signficiant improvement over the
+state-of-the-art.

### 2017-05-14 abstract.tex
updated abstract

@@
 systems demonstrate ability to achieve tag-to-tag communication 
-under conditions when tags are close to television tower, i.e., with the
+under conditions when the
 strength of TV signals  sufficiently high (-30 dBm).  In this paper, we present our preliminary work
-which demonstrates the possibility to backscatter and communicate using signficiantly weaker 
-ambient television signals. The key to using weaker broadcast television signals, is to leverage a low-cost
+which demonstrates the possibility to backscatter and communicate even when the ambient television signals
+are significantly weaker. The key to achieving this is to leverage a low-cost
 software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
@@
 to operate, and communicate, a signficiant improvement over the
-state-of-the-art.
+state-of-the-art restricted to operate only when the tags are close to the television towers.

### 2017-05-14 introduction.tex
updated abstract

@@
-Backscatter enables wireless communication at orders of magnitude lower
-energy cost than traditional radios. A backscatter transmitter operates
-by either reflecting or absorbing existing wireless signals, which only
-incurrs in very low power consumption in the order of
-micro-Watts~\cite{liu_ambient_2013}. As a consequence backscatter
+Backscatter communication enables wireless transmissions
+at energy consumption which is several orders of magnitude
+lower than traditional radios. Backscatter achieves ultra-low
+power wireless transmissions by reflecting or absorbing ambient
+wireless signals, which as an operation consumes only 
+\SI{}{\micro\watt}s of power consumption~\cite{liu_ambient_2013}. 
+As a consequence backscatter
 communication is emerging as the mechanism of choice to network
-battery-free devices.  Over the past few years significant progress has
-been made to make backscatter communication practical for Internet of
-Things~(IoT) devices. Early backscatter systems, like RFID, required a
-dedicated device to generate the necessary carrier to be backscattered.
-State-of-the-art systems, however, reflect ambient WiFi or television
-signals~\cite{liu_ambient_2013} instead, which does away with the need for
-dedicated infrastructure.
+devices operating on harvested energy.  
+Over the past few years there has been significant progress
+to make backscatter a viable mechanism to network Internet of
+Things~(IoT) devices. Early backscatter systems, like RFID readers, required a
+dedicated device to generate the necessary carrier signal, which were later
+reflected by the tags. On the other hand, recent state-of-the art systems 
+does away the need for a dedicated device to generate carrier signals. 
+Recent systems can leverage either already deployed infrastructure of
+devices to generate carrier signal, or leverage ambient WiFi or TV
+signals~\cite{liu_ambient_2013}.
 

### 2017-05-14 eval.tex
more changes

--- 2017-05-14 paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\usepackage{booktabs} % For formal tables
-\newenvironment{itemize*}{\begin{itemize}}{\end{itemize}}
-\usepackage[utf8x]{inputenc}
-\usepackage{subfigure}
-\usepackage{balance}
-\usepackage{comment}
-\usepackage{amsmath}
-\usepackage{xspace}
-\usepackage{caption}
-\usepackage{siunitx}
-\usepackage{booktabs}
+In this section, we present the results of evaluation of our system. As experimental 
+setup, we use RTLSDR together with signal processing algorithm implemented
+in Octave.
 
-\usepackage{cite}
+\subsection{Spatial variation of ambient television signals}
 
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
- \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
- \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
- \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
- \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-\usepackage{amsmath}
-\usepackage{amssymb}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% For special characters like ä,ö,...
-\usepackage[utf8x]{inputenc}
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
-\newcommand{\fakeparagraph}[1]{\vspace{.5mm}\noindent\textbf{#1.}}
-\newcommand{\fakepar}[1]{\fakeparagraph{#1}}
-
-\begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Maximilian Stiefel}
-\IEEEauthorblockA{Uppsala University\\
-maximilian.stiefel.8233@student.uu.se}
-\and
-\IEEEauthorblockN{Elmar van Rijnswou}
-\IEEEauthorblockA{Uppsala University\\
-elmar.vanrijnswou.9818@student.uu.se}
-\and
-\IEEEauthorblockN{Carlos Pérez-Penichet}
-\IEEEauthorblockA{Uppsala University\\
-carlos.penichet@it.uu.se                              }
-\and
-\IEEEauthorblockN{Ambuj Varshney}
-\IEEEauthorblockA{Uppsala University\\
-ambuj.varshney@it.uu.se}
-\and
-\IEEEauthorblockN{Christian Rohner}
-\IEEEauthorblockA{Uppsala University\\
-christian.rohner@it.uu.se}
-\and
-\IEEEauthorblockN{Thiemo Voigt}
-\IEEEauthorblockA{Uppsala University\\
-and RISE SICS\\
-thiemo@sics.se}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-\input{abstract}
-
-\end{abstract}
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-% no \IEEEPARstart
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-
-\input{introduction}
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-\section{Background}
-
-In this section we introduce some necessary background related to our
-work.
-
-%\subsection{Software Defined Radio}
-%Traditionally basic radio components like mixers or filters have been realized
-%in hardware. Nowadays as soon as one works in the baseband frequency the
-%computational resources of a personal computer are sufficient to realize these
-%components with signal processing done in software. \cp{We need to expand on
-%this introduction to SDR, assuming we need it and not just assume everyone
-%knows what that is.}
-
-\subsection{Backscatter Transmissions}
-\input{backscatter}
-
-\subsection{RTL2832U}
-\label{sub:rtl2832}
-
-\begin{figure}[h]
-				\centering
-				\includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
-				\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q
-				demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner
-				with integrated LNA \textit{Rafael Micro R820T/2} (right IC).} 
-				\label{fig:receiver_arch}
-\end{figure}
-The \textit{Realtek RTL2832U} is a terrestrial digital video broadcast
-(DVB-T) demodulator IC, which is the main part of a wide range of
-USB-based DVB-T receiver dongles. One prominent feature of this chip is
-that it allows to retrieve the raw I/Q samples via USB. This is
-originally intended for the chip to work as a simple DAB/FM software
-receiver. Ham radio enthusiasts have combined their efforts to create a
-software driver (the \textit{librtlsdr}, cf. \cite{steve-m_librtlsdr})
-that allows DVB-T dongles based on the \textit{RTL2832} to be converted
-into low-cost wideband SDR receivers. The cost of traditional SDRs has
-been in the range of a few hundred to thousands of USD. Albeit much less
-powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
-can be bought with antenna for less than 10 USD.
-
-An RTL-SDR also contains a tuner that allows the user to select the
-received frequency.  In our case the tuner is a \textit{Rafael Micro
-R820T/2} which offers a tuning range from 42 to 1002
-MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:receiver_arch}, shows a
-photograph of the RTL-SDR used in our work.
-
-\subsection{Quadrature Demodulation}
-
-The received signal can be interpreted as
-\begin{equation}
-	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
-\end{equation}     
-This is multiplied with a cosine of the carrier frequency from a local oscillator.
-\begin{multline}
-        s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
-\end{multline}
-With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
-\begin{multline}
-        2 \cdot s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
-\end{multline}
-One can see, that the interesting in-phase part (I) in this case is
-represented by a DC value after mixing. With a low-pass filter this DC
-value can be separated from the undesired rest. An analogous procedure
-is done with the quadrature part (Q) when mixing with a sine.
-
-\section{Design}
-
-In this section we describe the most important aspects of our design.
-
-\subsection{Transmitter}
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
-%\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
-%\label{fig:s11}
-%\end{figure}
-
-%A backscattering antenna is the most important part of our backscatter
-%transmitter. 
-Our backscatter transmitter is designed to be used with the television
-signal, which has its center frequency at 626 MHz. To that end, we build
-an
-antenna using an unetched PCB plate as the ground plane and a wire that acts
-as the vertical antenna element.% is soldered to the transmission line. The
-%transmission line has been calculated for 626 MHz using an FR4 substrate
-%with a thickness of 1.6 mm. The design equations given in
-%\cite{pozar_microwave_2011} are implemented in various programs e.g.
-%KiCAD, which has been used for solving the design equations. An
-%open-stub, attached directly at the juncture where the vertical antenna
-%element is attached to the transmission line, is used for fine tuning of
-%the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
-%Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
-%say, that the antenna works in a range from 626 MHz to 645 MHz. For
-%backscattering this small bandwidth is not necessarily a disadvantage,
-%since everything that is within the operating frequency of the antenna
-%is backscattered. With our narrowband antenna the undesired
-%backscattering of out-of-band signals is reduced.  With our approach we
-%were able to achieve an input impedance at 634 MHz of
-%\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
-%\ensuremath{S_{\text{11}}} is still approximately -17 dB.
-This antenna works well in the range from 626 to 245 MHz, which has been
-measured with a vector network analyzer (VNA).
-
-Besides the antenna, the transmitter consists of an \textit{MSP430} MCU,
-that sends data to the receiver. This transmitter controls an RF switch
-through an I/O pin. The switch can alternate the impedance connected to
-the antenna between matched (50 \ensuremath{\Omega}) and open. The
-software now steers this switch with a two 2 MHz rectangular timer
-signal to transmit a data 1 or it just leaves the antenna open to
-transmit a data 0. Therefore the result is a frequency shift of the
-television signal by 2 MHz for a 1 and no shift for a 0. Hence a classical
-binary amplitude shift keying is the implemented modulation technique.
-
-\subsection{Receiver}
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{./fig/receiver_arch}
-\caption{Receiver architecture from a system point of view. Signal flow is from left to right.}
-\label{fig:receiver_arch}
-\end{figure}
-The receiver consists of three main modules interconnected by a bus. The
-RTL-SDR block is in charge of collecting samples from the SDR-device. The
-demodulator turns the sampled signal into a streams of ones and zeroes.
-Finally, the Decoder block is in charge of detecting frames. We provide our C++
-receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.
-
-%A highly
-%sophisticated bus system has been developed to exchange data between the
-%different components of the system. 
-Figure \ref{fig:receiver_arch} shows the architecture of our receiver, the
-signal flows from the left to the right.
-%\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
-%\textit{libusb}, is used to transfer the data from the TV stick into our
-%program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
-%down the high frequency to a intermediate frequency (both values are
-%software adjustable). Subsequently the data is mixed with sine and
-%cosine, low-pass filtered and finally transferred to the digital domain.
-The raw I/Q data is available as two 8-bit values (real and imaginary) from the
-RTL-SDR. With these two values, the phase and magnitude of the data can be
-determined for every sample. The received signal can be represented as:
-\begin{equation}
-	I+j\,Q = abs(I,Q) \cdot e^{j \times ang(I,Q)} 
-\end{equation}
-
-So the first block, entitled RTL2832U, provides the interface to the TV
-dongle. It controls the frequency \ensuremath{f_{tuned}}, where the
-receiver is listening as well as other interesting parameters e.g. the analog gain.
-
-The demodulator block is responsible for converting the sampled signals
-into ones and zeroes. 
-%Therefore the real and imaginary values first have
-%to be deinterleaved out of the sample message. Now they are represented
-%as integer values between 0 and 255 (8 bit). So one converts them to
-%floats, whereas 127.5 equals 0. After that the values have to be
-%downsampled. Usually we were sampling with 250 kHz. This is still much
-%more than we need assuming an Additive White Gaussian Noise (AWGN)
-%channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
-%So we decided to reduce the sampling frequency to 25 kHz.
-%\begin{equation}
-%	\label{eq:awgn}
-%	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
-%\end{equation}     
-%To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
-%\begin{equation}
-%	\label{eq:nyquist}
-%	f_{samp} \geq 2 \cdot f_{a}
-%\end{equation}
-%\ensuremath{f_{a}} is the highest frequency in the signal. A FIR
-%filter naively implemented with floats has been used to achieve this. 10
-%kHz is the cut-off frequency of the first filter. After downsampling
-%another filter is used to suppress noise. This filter has a cut-off
-%frequency of 3 kHz. 
-As part of the demodulation process the signal is down-sampled and
-low-pass-filtered two times. The last demodulation step is to rectify
-the signal (cf. equation \ref{eq:rectify}) and decide with a
-software-defined Schmitt trigger whether a sample is a 1 or a 0.  
-\begin{equation}
-	\label{eq:rectify}
-	abs(I,Q) = \sqrt{I^2 +  Q^2}
-\end{equation}
-
-The resulting values are sent through the bus to the decoder. After the decoder
-received the samples it decides when a frame is sent or when the channel is
-idle. The decoder also includes a function to correlate the received bitstream
-with an expected pattern. Hence it is able to determine the bit error ratio
-(BER). Furthermore the code we have written so far subsumes different
-simulators for e.g. playing back recorded data. Another handy utility, which
-has been written in Octave, is a tool (an oscilloscope) to look at the data,
-which is written into files by the demodulator. An example of a recorded data
-stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.  
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{./fig/transmission}
-\caption{Start of a transmission of 101010..
-with 1 kbit/s with idle state before.
-Data from the Octave oscilloscope.
-Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
-This is due to a low signal strength.
-Sampling frequency is 25 kHz.}
-\label{fig:transmission}
-\end{figure}
-
-\section{Evaluation}
-In this section we present our evaluation results. The \textit{RTL-SDR},
-combined with intelligent signal processing in \textit{Octave}, are the
-main tools we have employed for measurements.
-
-\subsection{Signal Strength Variation in Space}
 \begin{figure}[h]
@@
 	\end{minipage}
-	\caption{Spectrum of the TV signal.
-Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
-Right: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
+	\vspace{-6mm}
+	\caption{\emph{Observed spectrum of television signal}.
+The left hand image shows rawspectrum of the TV signal with the centre frequency of \SI{626}{\mega\hertz}.
+The right hand image shows the smoothed spectrum, which is  normalized to maximum average.
+The average is  shown by the horizontal red line.}
 	\label{fig:tv_record} 
+		\vspace{-6mm}
 \end{figure}
@@
 experiment was also carried out outside, where we got similar results.
-
-\section{Discussion}
-
-Our system still has a lot of room for improvement.  Different
-backscattering antennas could be tried to find the one with the best
-combination of gain, tuning frequency and bandwidth. Moreover the
-standard \textit{RTL-SDR} antenna and its coaxial cable are of low
-quality.  Hence a custom antenna would be beneficial. Furthermore some
-fine tuning can be made when it comes to the filter coefficients of the
-different filters in the demodulator chain. Finally  errors can be
-further reduced by using one of many available forward error correcting
-codes.
-
-To come to a conclusion one can say, that we were able to show, that
-spectrum scanning over a huge band with sweeping is possible using the
-\textit{RTL-SDR}.  This allowed us to perform a survey of the TV signal
-strength in the city. And finally we have shown for the first time, that
-backscatter communication is possible with the \textit{RTL-SDR}.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-
-%The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-\bibliographystyle{IEEEtran}
-\bibliography{paper}
-
-\end{document}

### 2017-05-14 introduction.tex
more changes

@@
 to make backscatter a viable mechanism to network Internet of
-Things~(IoT) devices. Early backscatter systems, like RFID readers, required a
+Things~(IoT) devices. Traditional backscatter systems, like RFID readers, required a
 dedicated device to generate the necessary carrier signal, which were later
-reflected by the tags. On the other hand, recent state-of-the art systems 
+reflected by the tags. On the other hand, state-of-the art systems 
 does away the need for a dedicated device to generate carrier signals. 
-Recent systems can leverage either already deployed infrastructure of
-devices to generate carrier signal, or leverage ambient WiFi or TV
-signals~\cite{liu_ambient_2013}.
+Recent backscatter systems leverage already deployed infrastructure of
+devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{zhang2016hitchhike,kellogg2015wi} or TV
+signals~\cite{liu_ambient_2013,parks_turbocharging_2014}.
 

### 2017-05-14 paper.tex
updated abstract
more changes

@@
 
-% *** GRAPHICS RELATED PACKAGES ***
-%
 \ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
  \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
 \usepackage{amsmath}
 \usepackage{amssymb}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
 
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
+\usepackage{url}
 
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% For special characters like ä,ö,...
 \usepackage[utf8x]{inputenc}
 
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+
 \title{Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
 \author{\IEEEauthorblockN{Maximilian Stiefel}
@@
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
@@
 
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \section{Introduction}
-% no \IEEEPARstart
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
 
 \input{introduction}
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
 \section{Background}
@@
 work.
-
-%\subsection{Software Defined Radio}
-%Traditionally basic radio components like mixers or filters have been realized
-%in hardware. Nowadays as soon as one works in the baseband frequency the
-%computational resources of a personal computer are sufficient to realize these
-%components with signal processing done in software. \cp{We need to expand on
-%this introduction to SDR, assuming we need it and not just assume everyone
-%knows what that is.}
 
@@
 \subsection{Transmitter}
-%\begin{figure}[h]
-%\centering
-%\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
-%\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
-%\label{fig:s11}
-%\end{figure}
 
-%A backscattering antenna is the most important part of our backscatter
-%transmitter. 
 Our backscatter transmitter is designed to be used with the television
@@
 antenna using an unetched PCB plate as the ground plane and a wire that acts
-as the vertical antenna element.% is soldered to the transmission line. The
-%transmission line has been calculated for 626 MHz using an FR4 substrate
-%with a thickness of 1.6 mm. The design equations given in
-%\cite{pozar_microwave_2011} are implemented in various programs e.g.
-%KiCAD, which has been used for solving the design equations. An
-%open-stub, attached directly at the juncture where the vertical antenna
-%element is attached to the transmission line, is used for fine tuning of
-%the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
-%Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
-%say, that the antenna works in a range from 626 MHz to 645 MHz. For
-%backscattering this small bandwidth is not necessarily a disadvantage,
-%since everything that is within the operating frequency of the antenna
-%is backscattered. With our narrowband antenna the undesired
-%backscattering of out-of-band signals is reduced.  With our approach we
-%were able to achieve an input impedance at 634 MHz of
-%\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
-%\ensuremath{S_{\text{11}}} is still approximately -17 dB.
+as the vertical antenna element.
 This antenna works well in the range from 626 to 245 MHz, which has been
@@
 \includegraphics[width=\columnwidth]{./fig/receiver_arch}
-\caption{Receiver architecture from a system point of view. Signal flow is from left to right.}
+\caption{\emph{Design of the ambient backscatter receiver.} The flow of operation of the
+receiver is from the left to right.}
 \label{fig:receiver_arch}
@@
 into ones and zeroes. 
-%Therefore the real and imaginary values first have
-%to be deinterleaved out of the sample message. Now they are represented
-%as integer values between 0 and 255 (8 bit). So one converts them to
-%floats, whereas 127.5 equals 0. After that the values have to be
-%downsampled. Usually we were sampling with 250 kHz. This is still much
-%more than we need assuming an Additive White Gaussian Noise (AWGN)
-%channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
-%So we decided to reduce the sampling frequency to 25 kHz.
-%\begin{equation}
-%	\label{eq:awgn}
-%	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
-%\end{equation}     
-%To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
-%\begin{equation}
-%	\label{eq:nyquist}
-%	f_{samp} \geq 2 \cdot f_{a}
-%\end{equation}
-%\ensuremath{f_{a}} is the highest frequency in the signal. A FIR
-%filter naively implemented with floats has been used to achieve this. 10
-%kHz is the cut-off frequency of the first filter. After downsampling
-%another filter is used to suppress noise. This filter has a cut-off
-%frequency of 3 kHz. 
 As part of the demodulation process the signal is down-sampled and
@@
 \section{Evaluation}
-In this section we present our evaluation results. The \textit{RTL-SDR},
-combined with intelligent signal processing in \textit{Octave}, are the
-main tools we have employed for measurements.
-
-\subsection{Signal Strength Variation in Space}
-\begin{figure}[h]
-	\centering
-	\begin{minipage}{0.49\columnwidth}
-	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.49\columnwidth}
-	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
-	\end{minipage}
-	\caption{Spectrum of the TV signal.
-Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}.
-Right: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
-	\label{fig:tv_record} 
-\end{figure}
-
-We have used the \textit{RTL-SDR} to observe the space variations of the
-signal strength. \textit{Octave} scripts have been written to aid in
-this purpose. These scripts are available under \cite{s3xm3x_RTLSDRSpecAn}. The scripts perform a
-frequency sweep through the desired band. The obtained samples are then
-transferred from the time domain to the frequency domain using a fast
-fourier transformation (FFT).
-
-Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MS/s, we are
-limited to a maximum simultaneous acquisition band of 1.2 MHz. We keep a safety
-margin to the maximum capabilities and sample the desired frequency band at 900
-kHz intervals and stitch these bands together to form the overall frequency
-sweep. This approach is valid only because we are not interested in real-time
-data.  With this approach we are able to scan a 20 MHz band in roughly 30
-seconds.
-
-During the scanning process the gain is set to 40.2 dB. To get the
-signal strength we simply take the average over a 10 MHz band around the
-center frequency of the desired TV signal. This approach can be
-justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
-And the local TV provider advertises to be transmitting DVB-T 2.
-The signal \ensuremath{|H|} is calculated as follows.  
-\begin{equation}
-	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
-\end{equation}     where \ensuremath{N} is the total number of intervals
-accumulated over the entire frequency sweep. The FFT is only carried out over
-the real values, which are received from the \textit{RTL-SDR}.
-\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one
-interval (one subband). There are as many samples taken in one band as needed
-for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
-normalize everything and calculate the power. Before doing that, a simple
-smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
-\begin{equation}
-	|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
-\end{equation}   
-where \ensuremath{|H_0|} is a normalization value measured at a certain point
-in space. The result of combining measured signal strength values with the
-haversine function to calculate the distance to the point where the maximum
-signal has been captured can be observed in Figure \ref{fig:haversine}.
-\begin{figure}[h]
-	\centering
-	\begin{minipage}{0.49\columnwidth}
-	\includegraphics[width=\columnwidth]{./fig/haversine}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.49\columnwidth}
-	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
-	\end{minipage}
-	\caption{Left: Signal strength fading against distance relative
-to the spot where the maximum signal strength has been captured (TV
-towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
-594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
-calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
-signal in Uppsala.}
-
-\label{fig:haversine}
-\end{figure}
-
-\balance
-
-\subsection{Communication Performance}
-We tested the communication with different data rates from 1 bit/s up to
-1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
-strength from the backscatter transmitter is rather low. This leads to a
-significant amount of quantization noise to appear. This situation can
-be improved by increasing the analog gain of the receiver. The
-transmission shown in Figure \ref{fig:transmission} was carried out with
-the maximum gain (50 dB) available to our receiver.
-
-With a data rate of 1 bit/s, we were able to achieve a range indoors
-of a couple of meters. With higher bitrates we
-can only communicate over a range of a few decimeters and the 
-bit error rate is still approximately 40 \%. The communication
-experiment was also carried out outside, where we got similar results.
+\input{eval}
 
@@
 
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-%\section*{Acknowledgment}
-
-%The authors would like to thank...
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
 \bibliographystyle{IEEEtran}

### 2017-05-14 eval.tex
space fixes

@@
 	\caption{\emph{Observed spectrum of television signal}.
-The left hand image shows rawspectrum of the TV signal with the centre frequency of \SI{626}{\mega\hertz}.
-The right hand image shows the smoothed spectrum, which is  normalized to maximum average.
+The left hand image shows the raw spectrum of the TV signal, with centre frequency of \SI{626}{\mega\hertz}.
+The right hand image shows the smoothed spectrum normalized to maximum average.
 The average is  shown by the horizontal red line.}

### 2017-05-14 introduction.tex
space fixes

@@
 Things~(IoT) devices. Traditional backscatter systems, like RFID readers, required a
-dedicated device to generate the necessary carrier signal, which were later
+dedicated device to generate the necessary carrier signal
 reflected by the tags. On the other hand, state-of-the art systems 

### 2017-05-14 paper.tex
space fixes

@@
 \section{Design}
-
-In this section we describe the most important aspects of our design.
+In this section, we describe the design 
+of our system. We first describe the
+design of the transmitter, next we describe 
+the design of the receiver.
 
@@
 
-Our backscatter transmitter is designed to be used with the television
-signal, which has its center frequency at 626 MHz. To that end, we build
-an
-antenna using an unetched PCB plate as the ground plane and a wire that acts
-as the vertical antenna element.
-This antenna works well in the range from 626 to 245 MHz, which has been
-measured with a vector network analyzer (VNA).
+We design our backscatter transmitter to be used with ambient
+television signals.  We design the transmitter for ambient
+television signal present in the band with  
+center frequency of \SI{626}{\mega\hertz}. Crucial to the design
+of the transmitter is the implementation of an antenna, as
+the antennas are known to be  frequency selective. To this end, we design an
+antenna on an unetched board made of FR4 substrate commonly
+used to design PCBs. The board acts like the ground plane, at the center
+of the board we have a wire acting as a monopole antenna element. 
+We have verified using a vector network analyzer (VNA), that the antenna
+can operate from frequency range most commonly
+used for television broadcast, 
+from \SI{245}{\mega\hertz} to \SI{626}{\mega\hertz}.
 
@@
 \includegraphics[width=\columnwidth]{./fig/receiver_arch}
+\vspace{-6mm}
 \caption{\emph{Design of the ambient backscatter receiver.} The flow of operation of the
@@
 \label{fig:receiver_arch}
+\vspace{-6mm}
 \end{figure}
@@
 stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.
+
 \begin{figure}[h]
 \centering
+\vspace{-30mm}
+
 \includegraphics[width=\columnwidth]{./fig/transmission}
+\vspace{-30mm}
 \caption{Start of a transmission of 101010..
@@
 \label{fig:transmission}
+\vspace{-6mm}
+
 \end{figure}
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{abbrv}
 \bibliography{paper}

### 2017-05-14 paper.tex
update transmitter section

@@
 
-Besides the antenna, the transmitter consists of an \textit{MSP430} MCU,
-that sends data to the receiver. This transmitter controls an RF switch
-through an I/O pin. The switch can alternate the impedance connected to
-the antenna between matched (50 \ensuremath{\Omega}) and open. The
-software now steers this switch with a two 2 MHz rectangular timer
-signal to transmit a data 1 or it just leaves the antenna open to
-transmit a data 0. Therefore the result is a frequency shift of the
-television signal by 2 MHz for a 1 and no shift for a 0. Hence a classical
+Backscatter operates by reflecting or absorbing impinging
+radio signals on the antenna. To purposefully toggle the antenna
+to achieve these states, we  terminate the antenna to a 
+RF switch. The RF switch enables us to alternate the impedance connected to
+the antenna between matched (50 \ensuremath{\Omega}) and open state.
+We control the RF switch through an I/O pin on a low-powered MCU, Texas
+instruments MSP430. As discussed earlier, to reduce the effects of
+self-interference, we frequency shift the backscatter transmission
+away from the TV signal. Hence, we operate the RF switch using a 
+\SI{2}{\mega\hertz} intermediate frequency signal. Therefore the result 
+is a frequency shift of the television signal by 2 MHz, when
+transmitting a bit 1, and no shift for a 0.  Hence,  we 
+leverage amplitude modulation at receiver in our design.
 binary amplitude shift keying is the implemented modulation technique.

### 2017-05-15 introduction.tex
updates

@@
 Recent backscatter systems leverage already deployed infrastructure of
-devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{zhang2016hitchhike,kellogg2015wi} or TV
+devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{hitchhike,kellogg2015wi} or TV
 signals~\cite{liu_ambient_2013,parks_turbocharging_2014}.
 
-Recent ambient backscatter systems demonstrate tag-to-tag communication
-by leveraging TV signals as a source of both carrier and energy. For
+Recent backscatter systems demonstrate ability to leverage
+existing signals like TV signals as a both source of carrier and energy. For
 example, Liu et al. present a proof-of-concept system that reflects
@@
 meters by using analog coding~\cite{parks_turbocharging_2014}. While
-these systems can enable many applications, they are also restricted to
-operate only in environments where TV signals are fairly strong~(approx
-\SI{-30}{dBm}) due to the low sensitivity of the receiver used in these
-tags. On the other hand, TV signals are known to vary greatly in
-strength~\cite{wang_fm_2017} over space and time, making existing
-systems very restricted in their coverage.
+these systems can enable many applications, these systems are severely restricted
+to operate in the vicinity of television towers where ambient signal
+are sufficiently strong~(approx
+\SI{-30}{dBm}). This is primarily due to poor sensitivity levels of receivers
+employed on these devices, which together with weak backscatter reflections 
+severely limits the operating range from the tower.
+On the other hand, TV signals are known to vary greatly in
+strength~\cite{wang_fm_2017} both over space, and in time,  which further aggravates the
+problem of limited range of ambient backscatter systems.
 
-On the other hand, Software Defined Radios~(SDRs) are powerful devices
-that have significant processing abilities and offer relatively high
-receive sensitivity levels. The high sensitivity levels of the SDRs,
-compared to the typical ambient backscatter receiver tag, could
-significantly help improve the communication range and coverage when
-receiving ambient backscatter transmissions. More importantly, SDRs
-offer incomparable flexibility due to their ability to be reprogrammed.
-This flexibility offers the opportunity for ample experimentation, which
-is welcomed in this sort of emerging technology.
+On the other hand, Software Defined Radios~(SDRs) are powerful devices, and also
+have significant processing abilities. These devices offer high
+sensitivity levels, as compared to the receivers employed on typical 
+ambient backscatter tags. Thus, SDRs might help to
+significantly improve the communication range, and also coverage 
+area to receive ambient backscatter transmissions.
 
-In this paper we explore the following questions: Can we create an
-SDR-based receiver for ambient backscatter transmissions? Can we
-leverage the relatively high sensitivity of such a receiver to improve
-range and coverage of ambient backscatter systems? Can we provide a
-flexible and low-cost experimentation platform for ambient backscatter
-research?
+In this paper we explore the following questions: Can we leverage an
+SDR-based receiver to receive ambient backscatter transmissions ?, and, Does the  
+relatively high sensitivity of SDR receivers improve
+range and coverage of ambient backscatter systems? A positive answer to the above question would provide a
+flexible and low-cost experimentation platform to the wider research community to explore ambient backscatter
+systems.
 
-\fakepar{Contributions} The contributions of this work are twofold:
+\fakepar{Contributions}  In this paper, we make the following novel contributions:
+
 \begin{itemize}
-				\item We perform a survey of the signal strength of TV
-								broandcast signals.
 
-				\item We build a low-cost SDR-based receiver capable of
-								receiving data encoded in backscatterer TV signals.
+				\item We design the first system which leverages a  low-cost 
+					  SDR to receive ambient-backscatter transmissions. 
+				\item  Using the system designed, we demonstrate ambient backscatter using TV signals
+					   to be feasible in wide parts of a city. The range represents a significant improvement
+					   over the state-of-the-art.
+
 \end{itemize}

### 2017-05-15 abstract.tex

@@
 software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
-demonstrate that the strength of television signals is strong 
+demonstrate that the  television signals are strong 
 enough in most parts of a mid-sized swedish city for our system
-to operate, and communicate, a signficiant improvement over the
-state-of-the-art restricted to operate only when the tags are close to the television towers.
+to operate, and communicate, a significant improvement over the
+state-of-the-art which are
+restricted to operate only when the tags are close to the TV towers.

### 2017-05-15 backscatter.tex

@@
-% vim: textwidth=72 spell spelllang=en
-
 Consider an unmodulated carrier wave impinging on a backscatter tag's
@@
 self-interference through \textit{frequency-shifted}
-backscatter~\cite{kellogg_passive_2016,wang_fm_2017}.
+backscatter~\cite{kellogg_passive_2016,wang_fm_2017,varshney2016lorea}.
 The tags modulate their antenna in such a way that their transmissions
 occur at a certain frequency offset from the carrier signal thus
-allowing the receiver to avoid interference from the carrier by tunning
+allowing the receiver to avoid interference from the carrier by tuning
 to the offset frequency where the tag is transmitting.

### 2017-05-15 eval.tex

@@
 	\end{minipage}
-	\caption{Left: Signal strength fading against distance relative
-to the spot where the maximum signal strength has been captured (TV
-towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
-594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
-calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
-signal in Uppsala.}
+	\vspace{-6mm}
+	\caption{\emph{Spatial variation of TV signals in Uppsala city}. 
+	The left image demonstrates the signal strength of TV signal fading	
+	as opposed to observed signal strength near the TV tower. 
+	%474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
+%594 MHz belongs to a tower in Uppsala, Rickomberga. 
+	The right image visualises the signal strength as heatmap for the signal
+	present in the \SI{626}{\mega\hertz} band. Distances have been
+calculated with the Haversine formula. }
+ 	\vspace{-6mm}
 

### 2017-05-15 backscatter.tex
Fixed missing reference

@@
 If we focus on only one of the two generated images, or employ single
-sideband backscatter~\cite{interscatter}, the data transmission can now
+sideband backscatter~\cite{iyer2016inter}, the data transmission can now
 be received at frequency $f_d = f_c + \Delta f$.

### 2017-05-15 paper.tex
Smaller changes. References to github look better now.

@@
 \begin{equation}
-	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
+	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
 \end{equation}     
@@
 \begin{multline}
-        s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
 \end{multline}
@@
 \begin{multline}
-        2 \cdot s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
 \end{multline}
@@
 \begin{equation}
-	I+j\,Q = abs(I,Q) \cdot e^{j \times ang(I,Q)} 
+	I+j\,Q = abs(I,Q) \cdot e^{j \cdot ang(I,Q)} 
 \end{equation}

### 2017-05-15 paper.tex
Alternative author block for improved formatting

@@
 
-\author{\IEEEauthorblockN{Maximilian Stiefel}
-\IEEEauthorblockA{Uppsala University\\
-maximilian.stiefel.8233@student.uu.se}
-\and
-\IEEEauthorblockN{Elmar van Rijnswou}
-\IEEEauthorblockA{Uppsala University\\
-elmar.vanrijnswou.9818@student.uu.se}
-\and
-\IEEEauthorblockN{Carlos Pérez-Penichet}
-\IEEEauthorblockA{Uppsala University\\
-carlos.penichet@it.uu.se                              }
-\and
-\IEEEauthorblockN{Ambuj Varshney}
-\IEEEauthorblockA{Uppsala University\\
-ambuj.varshney@it.uu.se}
-\and
-\IEEEauthorblockN{Christian Rohner}
-\IEEEauthorblockA{Uppsala University\\
-christian.rohner@it.uu.se}
-\and
-\IEEEauthorblockN{Thiemo Voigt}
-\IEEEauthorblockA{Uppsala University\\
-and RISE SICS\\
-thiemo@sics.se}}
+\author{\IEEEauthorblockN{Maximilian Stiefel\IEEEauthorrefmark{1}, Elmar van Rijnswou\IEEEauthorrefmark{1}\\ 
+Carlos Pérez-Penichet\IEEEauthorrefmark{1} Ambuj Varshney\IEEEauthorrefmark{1}, 
+Christian Rohner\IEEEauthorrefmark{1} and Thiemo Voigt\IEEEauthorrefmark{2}}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}Uppsala University}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala Univerisyt and RISE SICS}}
 

### 2017-05-15 abstract.tex
Smaller changes. References to github look better now.
Several small corrections

@@
 and carrier signal. Existing state-of-the-art ambient backscatter
-systems demonstrate ability to achieve tag-to-tag communication 
+systems demonstrate the ability to achieve tag-to-tag communication 
 under conditions when the
-strength of TV signals  sufficiently high (-30 dBm).  In this paper, we present our preliminary work
+strength of TV signals is sufficiently high (-30 dBm).  In this paper, we present our preliminary work
 which demonstrates the possibility to backscatter and communicate even when the ambient television signals
 are significantly weaker. The key to achieving this is to leverage a low-cost
-software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
+software defined radio receiver (RTL-SDR) to receive backscattered transmissions. Our results 
 demonstrate that the  television signals are strong 

### 2017-05-15 backscatter.tex
Several small corrections

@@
 Consider an unmodulated carrier wave impinging on a backscatter tag's
-antenna. The signal observed by the receiver is:
-\begin{equation}
+antenna. In the presense of an umodulated carrier. The signal observed
+by the receiver is: \begin{equation}
 				S_r(t) = S_{rc}(t) + \sigma B(t)S_{bc}(t)
@@
 generator and $S_{bc}(t)$ is the signal from the carrier generator that
-reaches the backscatter tag. $B(t)$ is either zero or one and represents
-the instantaneous state of the backscatter tag: absorbing or reflecting,
-respectively.
+reaches the backscatter tag. $\sigma$ is the tag's radar cross section
+and $B(t)$ is either zero or one and represents the instantaneous state
+of the backscatter tag: absorbing or reflecting, respectively.
 
@@
 self-interference through \textit{frequency-shifted}
-backscatter~\cite{kellogg_passive_2016,wang_fm_2017,varshney2016lorea}.
+backscatter~\cite{kellogg_passive_2016,wang_fm_2017,varshney2016lorea,
+iyer2016inter}.
 The tags modulate their antenna in such a way that their transmissions

### 2017-05-15 eval.tex
Smaller changes. References to github look better now.
Several small corrections

@@
 In this section, we present the results of evaluation of our system. As experimental 
-setup, we use RTLSDR together with signal processing algorithm implemented
+setup, we use RTLSDR together with a signal processing algorithm implemented
 in Octave.
@@
 center frequency of the desired TV signal. This approach can be
-justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
+justified by the fact, that DVB-T 2 is specified up to 10 MHz bandwidth.
 And the local TV provider advertises to be transmitting DVB-T 2.
@@
 1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
-strength from the backscatter transmitter is rather low. This leads to a
-significant amount of quantization noise to appear. This situation can
+strength from the backscatter transmitter is rather low. This leads to the appearance of a
+significant amount of quantization noise. This situation can
 be improved by increasing the analog gain of the receiver. The

### 2017-05-15 introduction.tex
Several small corrections

@@
 Backscatter communication enables wireless transmissions
-at energy consumption which is several orders of magnitude
+at an energy consumption several orders of magnitude
 lower than traditional radios. Backscatter achieves ultra-low
 power wireless transmissions by reflecting or absorbing ambient
-wireless signals, which as an operation consumes only 
-\SI{}{\micro\watt}s of power consumption~\cite{liu_ambient_2013}. 
+wireless signals, which consumes only 
+\SI{}{\micro\watt}s of power~\cite{liu_ambient_2013}. 
 As a consequence backscatter
@@
 dedicated device to generate the necessary carrier signal
-reflected by the tags. On the other hand, state-of-the art systems 
-does away the need for a dedicated device to generate carrier signals. 
-Recent backscatter systems leverage already deployed infrastructure of
-devices to generate carrier signal~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi~\cite{hitchhike,kellogg2015wi} or TV
-signals~\cite{liu_ambient_2013,parks_turbocharging_2014}.
+reflected by the tags. On the other hand, state-of-the art systems does away
+with the need for a dedicated device to generate carrier signals.  Recent
+backscatter systems leverage the already deployed infrastructure such as
+WiFi~\cite{hitchhike,kellogg2015wi} or TV
+towers~\cite{liu_ambient_2013,parks_turbocharging_2014} to generate the carrier
+signal. %~\cite{varshney2016lorea,iyer2016inter}, or ambient WiFi.
 
-Recent backscatter systems demonstrate ability to leverage
-existing signals like TV signals as a both source of carrier and energy. For
+Recent backscatter systems demonstrate the ability to leverage
+existing signals, such as TV, as the source of both carrier and energy. For
 example, Liu et al. present a proof-of-concept system that reflects
@@
 meters by using analog coding~\cite{parks_turbocharging_2014}. While
-these systems can enable many applications, these systems are severely restricted
+these systems can enable many applications, they are severely restricted
 to operate in the vicinity of television towers where ambient signal
 are sufficiently strong~(approx
-\SI{-30}{dBm}). This is primarily due to poor sensitivity levels of receivers
+\SI{-30}{dBm}). This is primarily due to the poor sensitivity of the receivers
 employed on these devices, which together with weak backscatter reflections 
 severely limits the operating range from the tower.
-On the other hand, TV signals are known to vary greatly in
+Furthermore, TV signals are known to vary greatly in
 strength~\cite{wang_fm_2017} both over space, and in time,  which further aggravates the
-problem of limited range of ambient backscatter systems.
+problem of the limited range of ambient backscatter systems.
 
@@
 In this paper we explore the following questions: Can we leverage an
-SDR-based receiver to receive ambient backscatter transmissions ?, and, Does the  
-relatively high sensitivity of SDR receivers improve
-range and coverage of ambient backscatter systems? A positive answer to the above question would provide a
+SDR-based receiver to receive ambient backscatter transmissions?, and, Does the  
+relatively high sensitivity of SDR receivers improve the
+range and coverage of ambient backscatter systems? A positive answer to the above questions would provide a
 flexible and low-cost experimentation platform to the wider research community to explore ambient backscatter
@@
 
-				\item We design the first system which leverages a  low-cost 
+				\item We design the first system that leverages a  low-cost 
 					  SDR to receive ambient-backscatter transmissions. 
-				\item  Using the system designed, we demonstrate ambient backscatter using TV signals
+				\item  We demonstrate ambient backscatter using TV signals
 					   to be feasible in wide parts of a city. The range represents a significant improvement
@@
 \end{itemize}
+
+The paper proceeds as follows. First, we discuss relevant background information of backscatter communication,
+SDR and reception process. Next, we present the design of the transmitter and the receiver employed in this
+paper. We then present initial results which evaluates our system. Finally, we present concluding remarks.

### 2017-05-15 paper.tex
Fixed wrong figure reference a reviewer pointed to
Several small corrections

@@
 				with integrated LNA \textit{Rafael Micro R820T/2} (right IC).} 
-				\label{fig:receiver_arch}
+				\label{fig:rtl_hw}
 \end{figure}
 The \textit{Realtek RTL2832U} is a terrestrial digital video broadcast
-(DVB-T) demodulator IC, which is the main part of a wide range of
+(DVB-T) demodulator IC, which is the main component in a wide range of
 USB-based DVB-T receiver dongles. One prominent feature of this chip is
@@
 receiver. Ham radio enthusiasts have combined their efforts to create a
-software driver (the \textit{librtlsdr}, cf. \cite{steve-m_librtlsdr})
+software driver (the \textit{librtlsdr}~\cite{steve-m_librtlsdr})
 that allows DVB-T dongles based on the \textit{RTL2832} to be converted
-into low-cost wideband SDR receivers. The cost of traditional SDRs has
-been in the range of a few hundred to thousands of USD. Albeit much less
+into low-cost wideband SDR receivers. The cost of traditional SDRs is
+in the range of a few hundred to thousands of USD. Albeit much less
 powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
@@
 R820T/2} which offers a tuning range from 42 to 1002
-MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:receiver_arch}, shows a
+MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:rtl_hw}, shows a
 photograph of the RTL-SDR used in our work.
@@
 instruments MSP430. As discussed earlier, to reduce the effects of
-self-interference, we frequency shift the backscatter transmission
+self-interference, we frequency-shift the backscatter transmission
 away from the TV signal. Hence, we operate the RF switch using a 
@@
 with an expected pattern. Hence it is able to determine the bit error ratio
-(BER). Furthermore the code we have written so far subsumes different
-simulators for e.g. playing back recorded data. Another handy utility, which
-has been written in Octave, is a tool (an oscilloscope) to look at the data,
-which is written into files by the demodulator. An example of a recorded data
-stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.
+(BER). Furthermore we have implemented a series of aditional tools that include
+different simulators for e.g. playing back recorded data and a tool (an
+oscilloscope) to plot the data, which is written into files by the demodulator.
+An example of a recorded data stream with 25 kS/s can be seen in Figure
+\ref{fig:transmission}.
 
@@
 
-To come to a conclusion one can say, that we were able to show, that
-spectrum scanning over a huge band with sweeping is possible using the
+%To come to a conclusion one can say, that 
+We have shown, that
+spectrum scanning over a wide band with sweeping is possible using the
 \textit{RTL-SDR}.  This allowed us to perform a survey of the TV signal
-strength in the city. And finally we have shown for the first time, that
-backscatter communication is possible with the \textit{RTL-SDR}.
+strength in the city. We found that the signal strength around most of the city
+is adequate for backscatter communication using our receiver. And finally we
+have shown for the first time, that backscatter communication is possible with
+the \textit{RTL-SDR}.
 

### 2017-05-15 introduction.tex
Added sth. to the intro and fixed a typo.

@@
 area to receive ambient backscatter transmissions.
+Moreover the I/Q demodulator ICs, which represent the heart of an SDR system are very cheap as they are used a lot in digital communication.
+We are confident to be able to accelerate the research in the backscattering field, since we introduce a backscattering equipement, which is easily affordable by any student.
 

### 2017-05-15 paper.tex
Added sth. to the intro and fixed a typo.
some changes
Saved Thiemos changes.

@@
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Uppsala University}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala Univerisyt and RISE SICS}}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala University and RISE SICS}}
 
@@
 
-The received signal can be interpreted as
+Quadrature demodulation is the major feature of the \textit{RTL2832U},  which makes it possible to decode a digitally amplitude- and phase-shifted signal.
+The received signal (behind the tuner at the input of the \textit{RTL2832U}) can be interpreted as
 \begin{equation}
@@
 of the board we have a wire acting as a monopole antenna element. 
-We have verified using a vector network analyzer (VNA), that the antenna
-can operate from frequency range most commonly
-used for television broadcast, 
-from \SI{245}{\mega\hertz} to \SI{626}{\mega\hertz}.
+The antenna has a quite good reflection coefficient \ensuremath{S_{11}} 
+of -17 dB at 626 MHz, which is the frequency aimed for. This means, that at 
+this frequency the antenna absorber more energy form the electromagnetic field compared to another band. A vector network analyzer (VNA) has been used to analyze antenna performance.
 

### 2017-05-15 abstract.tex

--- 2017-05-15 abstract.tex
@@
-Backscatter communication enables ultra-low
-power wireless transmissions, and is attractive
-for networking devices which operate 
-on harvested energy. Ambient backscatter
-takes the concept further, by leveraging
-ambient wireless signals like
-television signals, as both the source of power
-and carrier signal. Existing state-of-the-art ambient backscatter
-systems demonstrate ability to achieve tag-to-tag communication 
-under conditions when the
-strength of TV signals  sufficiently high (-30 dBm).  In this paper, we present our preliminary work
-which demonstrates the possibility to backscatter and communicate even when the ambient television signals
-are significantly weaker. The key to achieving this is to leverage a low-cost
-software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
-demonstrate that the  television signals are strong 
-enough in most parts of a mid-sized swedish city for our system
-to operate, and communicate, a significant improvement over the
-state-of-the-art which are
-restricted to operate only when the tags are close to the TV towers.
+Ambient backscatter communication enables ultra-low
+power wireless transmissions and is hence attractive
+for networking devices which operate on harvested energy. 
+The data to be sent is modulated on an ambient signal, 
+such as TV signals. Tag-to-tag backscatter communication 
+requires rather strong ambient signals to achieve acceptable
+error rates and range.
+In this paper, we present a design and implementation of a
+backscatter receiver on a software defined radio plattform 
+(RTL-SDR), achieving higher sensitivity than state-of-the-art
+backscatter tags and thus enabling wide area ambient backscatter 
+communication from low-power tags. 
+Our measurements show that the TV signal is strong 
+enough in most parts of a mid-sized Swedish city for our
+system to operate and communicate.

### 2017-05-15 abstract.tex
Samll changes in abstract and fig. 5.

@@
 In this paper, we present a design and implementation of a
-backscatter receiver on a software defined radio plattform 
+backscatter receiver on a extremely low-cost software defined radio plattform 
 (RTL-SDR), achieving higher sensitivity than state-of-the-art

### 2017-05-15 paper.tex
Samll changes in abstract and fig. 5.

@@
 leverage amplitude modulation at receiver in our design.
-binary amplitude shift keying is the implemented modulation technique.
 
@@
 The resulting values are sent through the bus to the decoder. After the decoder
-received the samples it decides when a frame is sent or when the channel is
+received the samples it decides if a frame is received or if the channel is
 idle. The decoder also includes a function to correlate the received bitstream
@@
 \vspace{-30mm}
-\caption{Start of a transmission of 101010..
+\caption{Data from the Octave oscilloscope showing the start of a transmission of 101010..
 with 1 kbit/s with idle state before.
-Data from the Octave oscilloscope.
-Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
+The amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
 This is due to a low signal strength.
-Sampling frequency is 25 kHz.}
+Sampling frequency is 25 kHz.
+At the top one can see the raw signal from the ADC, in the middle the filtered signal and at the bottom the signal after the Schmitt trigger.}
 \label{fig:transmission}

### 2017-05-15 eval.tex

@@
 
-We have used the \textit{RTL-SDR} to observe the space variations of the
-signal strength. \textit{Octave} scripts have been written to aid in
-this purpose. These scripts are available under \cite{s3xm3x_RTLSDRSpecAn}. The scripts perform a
-frequency sweep through the desired band. The obtained samples are then
-transferred from the time domain to the frequency domain using a fast
-fourier transformation (FFT).
+In this experiment, we investigate the variation of the television signal
+in space, i.e., how does the strength of the signal varies over 
+the area of the city. We use the RTLSDR reader to obsevre the signal strength.
+We perform the frequency sweep of the desired television band. Next, we
+obtain the sample in the time domain, and convert them to corresponding
+samples in frequency domain by performing fast-fourier transformation~(FFT).
 
-Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MS/s, we are
-limited to a maximum simultaneous acquisition band of 1.2 MHz. We keep a safety
-margin to the maximum capabilities and sample the desired frequency band at 900
-kHz intervals and stitch these bands together to form the overall frequency
-sweep. This approach is valid only because we are not interested in real-time
-data.  With this approach we are able to scan a 20 MHz band in roughly 30
-seconds.
+As we are not interested in realtime visualisation of the television
+signals, we acquire the signals offline and visualise the samples.
+RTLSDR dongle has a maximum sampling rate of 2.4 MS/s, which limits us to
+scanning a band of \SI{1.2}{\mega\hertz} band in a single run. We  keep some safety-margin, and
+sample the desired band at \SI{900}{\kilo\hertz} intervals. We sample the band 
+\SI{20}{\mega\hertz} wide, by stitching individual bands. In our experiments, it takes roughly 30
+seconds for us to perform the scanning operation.
+
+%These scripts are available under \cite{s3xm3x_RTLSDRSpecAn}
 
@@
 where \ensuremath{|H_0|} is a normalization value measured at a certain point
-in space. The result of combining measured signal strength values with the
-haversine function to calculate the distance to the point where the maximum
-signal has been captured can be observed in Figure \ref{fig:haversine}.
+in space.
+
+Figure~\ref{fig:haversine} demonstrates the result of combing the measured
+signal strength values, together with the harvesine function enables us to
+calculate the distance from the point where maximum signal has been
+capture.
+
 \begin{figure}[h]
@@
 	as opposed to observed signal strength near the TV tower. 
-	%474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
-%594 MHz belongs to a tower in Uppsala, Rickomberga. 
 	The right image visualises the signal strength as heatmap for the signal
@@
 \subsection{Communication Performance}
-We tested the communication with different data rates from 1 bit/s up to
-1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
-strength from the backscatter transmitter is rather low. This leads to the appearance of a
-significant amount of quantization noise. This situation can
-be improved by increasing the analog gain of the receiver. The
-transmission shown in Figure \ref{fig:transmission} was carried out with
-the maximum gain (50 dB) available to our receiver.
+In this experiment, we present some preliminary results to evaluate the communication
+performance of our system.  In the experiment, we varied the bitrate of the transmitter
+from 1 bit/s up to 1 kbit/s. As we had seen earlier in the Figure~\ref{fig:transmission}, the strength
+of the backscattered signal is rather low leading to the appearance of significant quantisation noise.
+We improve the situation by increasing the analog gain of the receiver, and perform
+the experiment with the maximum gain \SI{50}{\decibel} available at out receiver.
 
-With a data rate of 1 bit/s, we were able to achieve a range indoors
-of a couple of meters. With higher bitrates we
-can only communicate over a range of a few decimeters and the 
-bit error rate is still approximately 40 \%. The communication
-experiment was also carried out outside, where we got similar results.
+In an indoor environment, where the ambient signal strength is low, we are
+able to achieve a data rate of 1 b/s with distances of
+couple of meters. At higher bitrates, we observe a range of few 
+few decimeters and the  bit error rate is still approximately 40 \%. 
+We observed similar results outdoors. We acknowledge the high bit errors
+observed due to primarily limited gain available on the low cost dongles.
+In future, we will work to improve this. We note, existing ambient backscatter
+systems do not operate well indoors at large distances, and our results
+provide preliminary results of  improvements possible by leveraging low-cost
+SDRs.

### 2017-05-15 eval.tex
Changed some things.

@@
 Figure~\ref{fig:haversine} demonstrates the result of combing the measured
-signal strength values, together with the harvesine function enables us to
-calculate the distance from the point where maximum signal has been
-capture.
+signal strength values, together with the haversine function enables us to
+calculate the distance from the point where the maximum signal has been
+captured. The fast fading of the signal has been proven to be highly fluctuating, but it is still hight enough to enable backscattering.
 
@@
 of the backscattered signal is rather low leading to the appearance of significant quantisation noise.
-We improve the situation by increasing the analog gain of the receiver, and perform
-the experiment with the maximum gain \SI{50}{\decibel} available at out receiver.
+The experiments have been carried out with the maximum gain available from the tuner of \SI{50}{\decibel}.
 
@@
 able to achieve a data rate of 1 b/s with distances of
-couple of meters. At higher bitrates, we observe a range of few 
-few decimeters and the  bit error rate is still approximately 40 \%. 
+couple of meters. At higher bitrates, we observe a range of a few decimeters and the  bit error rate is still approximately 40 \%. 
 We observed similar results outdoors. We acknowledge the high bit errors

### 2017-05-15 paper.tex
Changed some things.

@@
 
+\subsection{Quadrature Demodulation}
+
+Quadrature demodulation is the major feature of the \textit{RTL2832U},  which makes it possible to decode a digitally amplitude- and phase-shifted signal.
+The received signal (behind the tuner at the input of the \textit{RTL2832U}) can be interpreted as
+\begin{equation}
+	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
+\end{equation}     
+This is multiplied with a cosine of the carrier frequency from a local oscillator.
+\begin{multline}
+        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+\end{multline}
+With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
+\begin{multline}
+        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+\end{multline}
+One can see, that the interesting in-phase part (I) in this case is
+represented by a DC value after mixing. With a low-pass filter this DC
+value can be separated from the undesired rest. An analogous procedure
+is done with the quadrature part (Q) when mixing with a sine.
+
 \subsection{Backscatter Transmissions}
@@
 photograph of the RTL-SDR used in our work.
-
-\subsection{Quadrature Demodulation}
-
-Quadrature demodulation is the major feature of the \textit{RTL2832U},  which makes it possible to decode a digitally amplitude- and phase-shifted signal.
-The received signal (behind the tuner at the input of the \textit{RTL2832U}) can be interpreted as
-\begin{equation}
-	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
-\end{equation}     
-This is multiplied with a cosine of the carrier frequency from a local oscillator.
-\begin{multline}
-        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
-\end{multline}
-With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
-\begin{multline}
-        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
-\end{multline}
-One can see, that the interesting in-phase part (I) in this case is
-represented by a DC value after mixing. With a low-pass filter this DC
-value can be separated from the undesired rest. An analogous procedure
-is done with the quadrature part (Q) when mixing with a sine.
 
@@
 \vspace{-30mm}
-\caption{Data from the Octave oscilloscope showing the start of a transmission of 101010..
-with 1 kbit/s with idle state before.
-The amplitude is quite low with an average of 12 \% of the maximum, which is 127.5.
-This is due to a low signal strength.
+	\caption{\emph{Plots from Octave oscilloscope showing the start of a transmission of 101010..} In the top plot one can see the raw analog signal, in the middle plot the filtered signal and finally at the bottom the signal after processing though the Schmitt trigger.
+The amplitude is quite low with an average of 12 \% of the maximum, which is 127.5 due to a low signal strength.
 Sampling frequency is 25 kHz.
-At the top one can see the raw signal from the ADC, in the middle the filtered signal and at the bottom the signal after the Schmitt trigger.}
+}
 \label{fig:transmission}

### 2017-05-15 eval.tex
Fixed typos.

@@
 In this section, we present the results of evaluation of our system. As experimental 
-setup, we use RTLSDR together with a signal processing algorithm implemented
+setup, we use RTL-SDR together with a signal processing algorithm implemented
 in Octave.
@@
 in space, i.e., how does the strength of the signal varies over 
-the area of the city. We use the RTLSDR reader to obsevre the signal strength.
+the area of the city. We use the RTL-SDR reader to obsevre the signal strength.
 We perform the frequency sweep of the desired television band. Next, we
@@
 signals, we acquire the signals offline and visualise the samples.
-RTLSDR dongle has a maximum sampling rate of 2.4 MS/s, which limits us to
+RTL-SDR has a maximum sampling rate of 2.4 MS/s, which limits us to
 scanning a band of \SI{1.2}{\mega\hertz} band in a single run. We  keep some safety-margin, and

### 2017-05-21 presentation/presentation.tex
Started with the presentation.

\documentclass[apectratio=169]{beamer}
\usetheme{metropolis}           % Use metropolis theme
% For PDFs
\newcommand{\myast}{\ensuremath{^{\ast}}}
\newcommand{\mydagger}{\ensuremath{^{\dagger}}}
\usepackage{pdfpages}
\usepackage{minted}
\usepackage{siunitx}
\title{Enabling Ambient Backscatter\\
Using a Low-Cost Software Defined Radio}
\subtitle{Saving Energy/Low Power Communications}
\date{\vspace{1em}
\today}
\author{Maximilian Stiefel\myast, Elmar van Rijnswou\myast\\Carlos Pérez-Penichet\myast, Ambuj Varshney\myast, Christian Rohner\myast and Thiemo Voigt \mydagger
\\\myast Uppsala University
\\\mydagger Uppsala University and RISE SICS
}
\institute{13th Swedish National Computer Networking Workshop}
\begin{document}
  \maketitle

\begin{frame}{Table Of Contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Introduction}
\begin{frame}{What is Backscattering? (1)}
	\begin{figure}[H]	
		\centering
		\only<1>{\includegraphics[width=0.8\textwidth]{./fig/backscatter1}}
		\only<2>{\includegraphics[width=0.8\textwidth]{./fig/backscatter2}}
		\caption{Simplest form of backscattering.}
	\end{figure}
\end{frame}

\begin{frame}{What is Backscattering? (2)}
	\only<1->{\begin{figure}[H]	
		\centering
		\includegraphics[width=0.6\textwidth]{./fig/mixer}
		\caption{Classical communications engineering element: The mixer.}
	\end{figure}}
	\only<2->{\begin{equation}
	2\sin(f_c t) \sin(\Delta f t) = \cos[(f_c + \Delta f)t] - \cos[(f_c - \Delta f) t]
	\label{eq:mixing}
	\end{equation}}

\end{frame}

\begin{frame}{What is Backscattering? (3)}
	\only<1->{\begin{figure}[H]	
		\centering
		\includegraphics[height=0.5\textheight]{./fig/rfswitch}
		\caption{\SI{50}{\ohm} connected to a RF switch do the trick.}
	\end{figure}}
	\only<2->{
		\begin{equation}
			\Gamma = \frac{Z_L - Z_A}{Z_L + Z_A}
		\end{equation}
	}
\end{frame}

\begin{frame}{Why Backscattering?}
	\begin{itemize}
		\item<1-> Ultra-low power wireless transmissions by reflecting/absorbing EM waves (in orders of \SI{}{\micro\watt}) \cite{liu_ambient_2013}
		\item<2-> Leverage existing signals such as coming from WiFi \cite{hitchhike,kellogg2015wi} or TV towers \cite{liu_ambient_2013,parks_turbocharging_2014}
		\item<3-> Mechanism of choice to network devices operating on harvested energy
		\item<4-> Communication frontends are much simpler, smaller and cheaper, than traditional RF frontends
	\end{itemize}
\end{frame}

\section{Background}
\begin{frame}{What is Backscattering?}
	\begin{minipage}{0.47\textwidth}
	\begin{itemize}
		\item Different backscattering research has already been done (e.g. \cite{Vincent})
		\item Also a lot of things about the RTL2832U can be found (e.g. \cite{Ana}) 
	\end{itemize}
	\end{minipage}
	\hfill
	\begin{minipage}{0.47\textwidth}
	\begin{figure}[H]	
		\centering
		\includegraphics[width=1\textwidth]{./fig/rtlsdr}
		\caption{The IQ demodulator RTL2832U is available for less than 10 \$. Source: Ebay}
	\end{figure}
	\end{minipage}
\end{frame}

\begin{frame}{Background And Literature (2)}
	\begin{figure}[H]	
		\centering
		\includegraphics[height=0.75\textheight]{./fig/monopole_smith}
		\caption{Smitchart sweep of selfmade ground-plane antenna for 634 MHz (checkout \cite{Pozar}).}
	\end{figure}
\end{frame}

\begin{frame}{Background And Literature (3)}
	\begin{figure}[H]	
		\centering
		\includegraphics[height=0.75\textheight]{./fig/monopole_S11dB}
		\caption{S11 out of the S parameter file of the VNA plotted with Qucs.}
	\end{figure}
\end{frame}

\section{Design}
\begin{frame}{State Of The Project (1)}
	\begin{table}[]
	\centering
	\begin{tabular}{ll}
	\hline
	\textbf{Backscatter Tag}               &    \\\hline
	Manchester encoding                    &   \checkmark \\
	Weak error protection                  &   \checkmark\\
	Shifting by 2 MHz                      &   \checkmark\\
	Advanced frame design                  &   \checkmark\\
	\hline
	\textbf{Antenna}                       &    \\\hline
	Roughly tuned and matched              &    \checkmark\\\
	Mechanical stability                   &    \\
	Lumped elements tuning                 &    \\\hline
	\end{tabular}
	\caption{Backscatter tag and antenna}
	\end{table}
\end{frame}
\begin{frame}{State Of The Project (2)}
	\begin{table}[]
	\centering
	\begin{tabular}{ll}
	\hline
	\textbf{Receiver}                      &    \\\hline
	Signal processing                      &    \checkmark\\\
	Reception with certain bit error rate  &    \checkmark\\\
	Efficient infrastructure               &    \checkmark\\\
	Fast				       &    \\\hline
	\textbf{Measurements}                  &    \\\hline
	Processing and Measurement environment &    \checkmark\\\
	Massive amount of samples              &    \\\hline
	\end{tabular}
	\caption{Receiver and measurements}
	\end{table}
\end{frame}

\begin{frame}{State of the project (3)}
\begin{figure}[H]	
		\centering
		\includegraphics[height=0.75\textheight]{./fig/haversine}
		\caption{Signal strength depending on the distance of different frequencies.}
\end{figure}
\end{frame}

\begin{frame}[fragile]{State of the project (3)}
\begin{minted}[baselinestretch=1, fontsize=\tiny, linenos,frame=single,framesep=5pt]{Octave}
function d = haversine (lat1,lat2,lon1,lon2)
        R = 6371e3;                                                     % Earth radius
        rad = pi/180;

        phi1 = lat1*rad;
        phi2 = lat2*rad;
        dphi = (lat2-lat1)*rad;
        dlambda = (lon2-lon1)*rad;

        a = sin(dphi/2)*sin(dphi/2) + cos(phi1)*cos(phi2)*sin(dlambda/2)*sin(dlambda/2);
        c = 2*atan2(sqrt(a), sqrt(1-a));
        d = R*c;
endfunction;
\end{minted}
\end{frame}

\begin{frame}{State of the project (3)}
\begin{figure}[H]	
		\centering
		\includegraphics[height=0.75\textheight]{./fig/heatmap}
		\caption{Signal strength heatmap of 626 MHz signal.}
\end{figure}
\end{frame}

\begin{frame}{State Of The Project (4)}
	\only<1>{\begin{figure}[H]	
		\centering
		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch_blank}
		\caption{System architecture of the receiver.}
	\end{figure}}
	\only<2>{\begin{figure}[H]	
		\centering
		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch}
		\caption{System architecture of the receiver.}
	\end{figure}}
\end{frame}

\begin{frame}{State Of The Project (5)}
	\begin{figure}[H]	
		\centering
		\includegraphics[height=0.5\textheight]{./fig/osci}
		\caption{Octave oscilloscope written for debugging purposes.}
	\end{figure}
\end{frame}

\section{Evaluation}
\begin{frame}{Conclusion}
\begin{itemize}
	\item Working receiver \checkmark
	\item Spectrum anlayzer based on the RTL2832U \checkmark 
	\item Working backscatter tag \checkmark
	\item Lot of help from the communication group \checkmark
\end{itemize}
\end{frame}

\begin{frame}{Outlook}
\begin{itemize}
	\item Reliable communication (e.g. hamming code)
	\item More advanced signal processing (e.g. oversampling) 
	\item Better (selfmade) antenna
	\item Other receiver hardware? 
\end{itemize}
\end{frame}

   	\begin{frame}{Contact Information}
        	\begin{center}
                \begin{table}[]
                        \begin{tabular}{ll}
                                E: & Elmar.Vanrijnswou.9818@student.uu.se \\
                             	E: & Maximilian.Stiefel.8233@student.uu.se\\
                        \end{tabular}
                \end{table}
                \includegraphics[width=0.05\textwidth]{./fig/github} \hspace{0.1cm} \url{https://github.com/m3x1m0m}\\
        	\end{center}
        	\begin{figure}
                	\includegraphics[width=0.3\textwidth]{./fig/logo}
        	\end{figure}
  	\end{frame}

\begin{frame}[standout]
	Happy Coding :) 
\end{frame}

\begin{frame}[allowframebreaks]{References}

  \bibliography{presentation}
  \bibliographystyle{abbrv}

\end{frame}

\end{document}

### 2017-05-21 presentation/presentation.tex
Presentation finished.

@@
 
+\begin{frame}{Contributions}
+	\begin{itemize}
+		\item<1-> We designed the first system using the RTL-SDR to receive ambient-backscatter.
+		\item<2-> We demonstrate ambient backscatter using TV signals to be feasible in wider parts of the city, which is a significant improvement since the state-of-the-art is restricted to a TV towers proximity.
+	\end{itemize}
+\end{frame}
+
 \section{Background}
-\begin{frame}{What is Backscattering?}
-	\begin{minipage}{0.47\textwidth}
-	\begin{itemize}
-		\item Different backscattering research has already been done (e.g. \cite{Vincent})
-		\item Also a lot of things about the RTL2832U can be found (e.g. \cite{Ana}) 
-	\end{itemize}
-	\end{minipage}
-	\hfill
-	\begin{minipage}{0.47\textwidth}
+\begin{frame}{RTL-SDR Hardware and I/Q Demodulation (1)}
 	\begin{figure}[H]	
 		\centering
-		\includegraphics[width=1\textwidth]{./fig/rtlsdr}
-		\caption{The IQ demodulator RTL2832U is available for less than 10 \$. Source: Ebay}
-	\end{figure}
-	\end{minipage}
-\end{frame}
-
-\begin{frame}{Background And Literature (2)}
-	\begin{figure}[H]	
-		\centering
-		\includegraphics[height=0.75\textheight]{./fig/monopole_smith}
-		\caption{Smitchart sweep of selfmade ground-plane antenna for 634 MHz (checkout \cite{Pozar}).}
+		\only<1>{\includegraphics[width=0.7\textwidth]{./fig/rtlsdr}}
+		\only<2>{\includegraphics[width=0.7\textwidth]{./fig/rtlsdr1}}
+		\only<3>{\includegraphics[width=0.7\textwidth]{./fig/rtlsdr2}}
+		\caption{RTL-SDR hardware with the DVB-T I/Q demodulator
+		Raeltek RTL2832U (left IC) and the tuner with integrated LNA
+		Rafael Micro R820T/2 (right IC).}
 	\end{figure}
@@
 
-\begin{frame}{Background And Literature (3)}
+\begin{frame}{RTL-SDR Hardware and I/Q Demodulation (2)}
+	\only<1->{Impinging signal: 
+	\begin{equation}
+		s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
+	\end{equation}}
+	\only<2->{Mixing with local RF frequency:
+	\begin{multline}
+		s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
+	\end{multline}}
+	\only<3->{With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
+	\begin{multline}
+		        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
+	\end{multline}}
+
+\end{frame}
+
+\section{Design}
+\begin{frame}{Transmitter}
+	\only<1->{\begin{figure}[H]	
+		\centering
+		\includegraphics[width=0.7\textwidth]{./fig/transmitter}
+		\caption{Transmitter architecture.
+The ground plane antenna is roughly tuned to the center frequency of the TV signal, which is \SI{626}{\mega\Hz}.
+A microcontroller steers a RF switch with a rectangular signal to shift the TV wave in another band.}
+	\end{figure}}
+\end{frame}
+
+\begin{frame}{Receiver (1)}
 	\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.75\textheight]{./fig/monopole_S11dB}
-		\caption{S11 out of the S parameter file of the VNA plotted with Qucs.}
+		\only<1>{\includegraphics[width=0.9\textwidth]{./fig/receiver1}}
+		\only<2>{\includegraphics[width=0.9\textwidth]{./fig/receiver2}}
+		\caption{Design of the ambient backscatter receiver. The flow
+		of operation of the receiver is from the left to right. Code is publicly available under \cite{s3xm3x_backscatterBASKReceiver}.}
 	\end{figure}
@@
 
-\section{Design}
-\begin{frame}{State Of The Project (1)}
-	\begin{table}[]
-	\centering
-	\begin{tabular}{ll}
-	\hline
-	\textbf{Backscatter Tag}               &    \\\hline
-	Manchester encoding                    &   \checkmark \\
-	Weak error protection                  &   \checkmark\\
-	Shifting by 2 MHz                      &   \checkmark\\
-	Advanced frame design                  &   \checkmark\\
-	\hline
-	\textbf{Antenna}                       &    \\\hline
-	Roughly tuned and matched              &    \checkmark\\\
-	Mechanical stability                   &    \\
-	Lumped elements tuning                 &    \\\hline
-	\end{tabular}
-	\caption{Backscatter tag and antenna}
-	\end{table}
-\end{frame}
-\begin{frame}{State Of The Project (2)}
-	\begin{table}[]
-	\centering
-	\begin{tabular}{ll}
-	\hline
-	\textbf{Receiver}                      &    \\\hline
-	Signal processing                      &    \checkmark\\\
-	Reception with certain bit error rate  &    \checkmark\\\
-	Efficient infrastructure               &    \checkmark\\\
-	Fast				       &    \\\hline
-	\textbf{Measurements}                  &    \\\hline
-	Processing and Measurement environment &    \checkmark\\\
-	Massive amount of samples              &    \\\hline
-	\end{tabular}
-	\caption{Receiver and measurements}
-	\end{table}
+\begin{frame}{Receiver (2)}
+	\only<1->{The received signal can be interpreted as:
+	\begin{equation}
+		        I+j\,Q = abs(I,Q) \cdot e^{j \cdot ang(I,Q)} 
+	\end{equation}}
+	\only<2->{
+	Rectification:
+	\begin{equation}
+		abs(I,Q) = \sqrt{I^2 +  Q^2}
+	\end{equation}}
 \end{frame}
 
-\begin{frame}{State of the project (3)}
-\begin{figure}[H]	
+\section{Evaluation}
+\begin{frame}{Spatial Variation of Ambient TV Signals (1)}
+\begin{figure}[h]
 	\centering
-		\includegraphics[height=0.75\textheight]{./fig/haversine}
-		\caption{Signal strength depending on the distance of different frequencies.}
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
+	\end{minipage}
+	\caption{\emph{Observed spectrum of television signal}.
+The left hand image shows the raw spectrum of the TV signal, with centre frequency of \SI{626}{\mega\hertz}.
+The right hand image shows the smoothed spectrum normalized to maximum average.
+The average is  shown by the horizontal red line.}
+	\label{fig:tv_record} 
 	\end{figure}
@@
 
-\begin{frame}[fragile]{State of the project (3)}
-\begin{minted}[baselinestretch=1, fontsize=\tiny, linenos,frame=single,framesep=5pt]{Octave}
-function d = haversine (lat1,lat2,lon1,lon2)
-        R = 6371e3;                                                     % Earth radius
-        rad = pi/180;
-
-        phi1 = lat1*rad;
-        phi2 = lat2*rad;
-        dphi = (lat2-lat1)*rad;
-        dlambda = (lon2-lon1)*rad;
-
-        a = sin(dphi/2)*sin(dphi/2) + cos(phi1)*cos(phi2)*sin(dlambda/2)*sin(dlambda/2);
-        c = 2*atan2(sqrt(a), sqrt(1-a));
-        d = R*c;
-endfunction;
-\end{minted}
+\begin{frame}{Spatial Variation of Ambient TV Signals (2)}
+	Perform FFT and stick bands together.
+	\begin{equation}
+		|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
+	\end{equation}
+	Smoothing and signal strength calculation. 
+	\begin{equation}
+		|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
+	\end{equation}
+	This has been done with Octave \cite{s3xm3x_RTLSDRSpecAn}.
 \end{frame}
 
-\begin{frame}{State of the project (3)}
-\begin{figure}[H]	
+\begin{frame}{Spatial Variation of Ambient TV Signals (3)}
+\begin{figure}[h]
 	\centering
-		\includegraphics[height=0.75\textheight]{./fig/heatmap}
-		\caption{Signal strength heatmap of 626 MHz signal.}
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/haversine}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
+	\end{minipage}
+	\caption{\emph{Spatial variation of TV signals in Uppsala city}.
+The left image demonstrates the signal strength of TV signal fading as opposed to observed signal strength near the TV tower.
+The right image visualises the signal strength as heatmap for the signal present in the \SI{626}{\mega\hertz} band.
+Distances have been calculated with the Haversine formula.
+}
+	\label{fig:haversine}
 \end{figure}
@@
 
-\begin{frame}{State Of The Project (4)}
-	\only<1>{\begin{figure}[H]	
+\begin{frame}{Communication Performance (1)}
+	\begin{figure}[h]
 	\centering
-		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch_blank}
-		\caption{System architecture of the receiver.}
-	\end{figure}}
-	\only<2>{\begin{figure}[H]	
-		\centering
-		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch}
-		\caption{System architecture of the receiver.}
-	\end{figure}}
-\end{frame}
-
-\begin{frame}{State Of The Project (5)}
-	\begin{figure}[H]	
-		\centering
-		\includegraphics[height=0.5\textheight]{./fig/osci}
-		\caption{Octave oscilloscope written for debugging purposes.}
+	\includegraphics[width=0.5\textwidth]{./fig/transmission}
+	\caption{\emph{Plots from Octave oscilloscope showing the start of a transmission of 101010..} In the top plot one can see the raw analog signal, in the middle plot the filtered signal and finally at the bottom the signal after processing though the Schmitt trigger.
+The amplitude is quite low with an average of 12 \% of the maximum, which is 127.5 due to a low signal strength.
+Sampling frequency is 25 kHz.
+}
+	\label{fig:transmission}	
 \end{figure}
@@
 
-\section{Evaluation}
-\begin{frame}{Conclusion}
+\begin{frame}{Communication Performance (2)}
 	\begin{itemize}
-	\item Working receiver \checkmark
-	\item Spectrum anlayzer based on the RTL2832U \checkmark 
-	\item Working backscatter tag \checkmark
-	\item Lot of help from the communication group \checkmark
+		\item<1-> We varied the bitrate between 1 bit/s and 1 kbit/s 
+		\item<2-> Strength of the backscattered signal is quite low (high quantisation noise)
+		\item<3-> Higher bitrates (1 kbit/s) lead to a range of a couple of decimeter before the bitrate goes up rapidly (40 \%)
+		\item<4-> We acknowledge the high bit errors observed due to primarily limited gain available on the low cost dongles
 	\end{itemize}
@@
 
-\begin{frame}{Outlook}
+\section{Discussion}
+
+\begin{frame}{Discussion}
 	\begin{itemize}
-	\item Reliable communication (e.g. hamming code)
-	\item More advanced signal processing (e.g. oversampling) 
-	\item Better (selfmade) antenna
-	\item Other receiver hardware? 
+		\item<1-> Our system has a lot of space for improvement
+		\item<2-> Other antennas for reception and transmission have to be tried out
+		\item<3-> The standard cable and the antenna of the RTL-SDR (included in the 10 USD budget) are unsuiteable for our application as we realized
+		\item<4-> The filters involved are setscrews as well as the coding used by the communication system (currently none)
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Results}
+	\begin{itemize}
+		\item<1-> We showed thath spectrum scanning over a huge band is possible using the RTL-SDR.
+		\item<2-> We found by carrying out a TV signal strength survey that the city is adequate for backscatter communication using our receiver.
+		\item<3-> We have shown for the first time, that backscatter communication is possible with the RTL-SDR.
 	\end{itemize}
@@
                              	E: & Maximilian.Stiefel.8233@student.uu.se\\
+            				& \\
+			\includegraphics[width=0.05\textwidth]{./fig/github} & \url{https://github.com/elamre}\\ 
+	& \url{https://github.com/m3x1m0m}
+
 		\end{tabular}
                 \end{table}
-                \includegraphics[width=0.05\textwidth]{./fig/github} \hspace{0.1cm} \url{https://github.com/m3x1m0m}\\
+
         	\end{center}
@@
   \bibliography{presentation}
-  \bibliographystyle{abbrv}
+  \bibliographystyle{ieeetr}
 

### 2017-05-27 presentation/presentation.tex
Finished presentation.

--- 2017-05-21 presentation/presentation.tex
@@
 \newcommand{\mydagger}{\ensuremath{^{\dagger}}}
+\usepackage[utf8x]{inputenc}
 \usepackage{pdfpages}
@@
 \author{Maximilian Stiefel\myast, Elmar van Rijnswou\myast\\Carlos Pérez-Penichet\myast, Ambuj Varshney\myast, Christian Rohner\myast and Thiemo Voigt \mydagger
+\vspace{1em}
 \\\myast Uppsala University
@@
 
-\begin{frame}{Table Of Contents}
-  \setbeamertemplate{section in toc}[sections numbered]
-  \tableofcontents[hideallsubsections]
+\begin{frame}{Contributions}
+	\begin{itemize}
+		\item<1-> First system using the RTL-SDR to receive ambient backscatter
+		\item<2-> Showing that ambient backscatter using TV signals to be feasible in wider parts of the city \only<3->{\ensuremath{\rightarrow} Significant improvement of the state-of-the-art, which is restricted to a TV towers proximity}
+	\end{itemize}
 \end{frame}
@@
 \section{Introduction}
+
 \begin{frame}{What is Backscattering? (1)}
@@
 		\centering
-		\only<1>{\includegraphics[width=0.8\textwidth]{./fig/backscatter1}}
-		\only<2>{\includegraphics[width=0.8\textwidth]{./fig/backscatter2}}
-		\caption{Simplest form of backscattering.}
+		\only<1>{\includegraphics[width=0.7\textwidth]{./fig/backscatter_simpler}}
+		\only<2->{\includegraphics[width=0.7\textwidth]{./fig/backscatter_simple}}
+
+		\caption{Simplest form of backscattering. Simplex with two subscribers.}
 	\end{figure}
+	\metroset{block=fill}
+	\only<3>{\begin{block}{Backscattering}
+		Communication technique similar to passive RFID, but the transmitter has to maintain its signal. 
+	\end{block}}
 \end{frame}
@@
 \begin{frame}{What is Backscattering? (2)}
+\begin{minipage}{0.45\textwidth}
+	\begin{figure}[H]	
+		\centering
+		\includegraphics[height=0.5\textheight]{./fig/rfswitch}
+		\caption{\SI{50}{\ohm} connected to a RF switch do the trick.}
+	\end{figure}
+\end{minipage}
+\hfill
+\begin{minipage}{0.45\textwidth}
+	\only<2->{\begin{equation}
+		P_R = P_A \cdot (1-\Gamma^2)
+	\end{equation}}
+	\only<3->{\begin{equation}
+		\Gamma = \frac{Z_L - Z_A}{Z_L + Z_A}
+	\end{equation}}
+	\only<4->{Total reflection: 
+	\begin{equation*}
+		\lim_{Z_L \to \infty} \Gamma = \frac{Z_L - Z_A}{Z_L + Z_A} = 1
+	\end{equation*}}
+	\only<5->{Total absorption: 
+	\begin{equation*}
+		\lim_{Z_L \to \SI{50}{\ohm}} \Gamma = \frac{Z_L - Z_A}{Z_L + Z_A} = 0
+	\end{equation*}}
+\end{minipage}
+\end{frame}
+
+\begin{frame}{Frequency Shift Keying With Backscatter Tags (1)}
+	\begin{figure}[H]	
+		\centering
+		\only<1>{\includegraphics[width=0.8\textwidth]{./fig/backscatter}}
+		\caption{Frequency shift keying modulation techniques are enabled by switching with a higher frequency.}
+	\end{figure}
+	\metroset{block=fill}
+\end{frame}
+
+\begin{frame}{Frequency Shift Keying With Backscatter Tags (2)}
 	\only<1->{\begin{figure}[H]	
 		\centering
-		\includegraphics[width=0.6\textwidth]{./fig/mixer}
+		\includegraphics[width=0.5\textwidth]{./fig/mixer}
 		\caption{Classical communications engineering element: The mixer.}
@@
 	\only<2->{\begin{equation}
-	2\sin(f_c t) \sin(\Delta f t) = \cos[(f_c + \Delta f)t] - \cos[(f_c - \Delta f) t]
+	2\sin(\omega_0 t) \sin(\omega_B) = \cos[(\omega_0 + \omega_B)t] - \cos[(\omega_0 - \omega_B) t]
 	\label{eq:mixing}
@@
 
-\end{frame}
-
-\begin{frame}{What is Backscattering? (3)}
-	\only<1->{\begin{figure}[H]	
-		\centering
-		\includegraphics[height=0.5\textheight]{./fig/rfswitch}
-		\caption{\SI{50}{\ohm} connected to a RF switch do the trick.}
-	\end{figure}}
-	\only<2->{
-		\begin{equation}
-			\Gamma = \frac{Z_L - Z_A}{Z_L + Z_A}
-		\end{equation}
-	}
 \end{frame}
@@
 		\item<3-> Mechanism of choice to network devices operating on harvested energy
-		\item<4-> Communication frontends are much simpler, smaller and cheaper, than traditional RF frontends
+		\item<4-> Communication frontends much simpler, smaller and cheaper, in comp. with traditional RF frontends
 	\end{itemize}
@@
 
-\section{Background}
-\begin{frame}{What is Backscattering?}
-	\begin{minipage}{0.47\textwidth}
-	\begin{itemize}
-		\item Different backscattering research has already been done (e.g. \cite{Vincent})
-		\item Also a lot of things about the RTL2832U can be found (e.g. \cite{Ana}) 
-	\end{itemize}
+\begin{frame}{TV Signal Backscattering Application Example}
+\begin{minipage}{0.45\textwidth}
+	\begin{figure}[H]	
+		\centering
+		\includegraphics[height=0.5\textheight]{./fig/elmars_home}
+		\caption{Houses in Amsterdam. Source: reddit.com}
+	\end{figure}
 \end{minipage}
 \hfill
-	\begin{minipage}{0.47\textwidth}
-	\begin{figure}[H]	
-		\centering
-		\includegraphics[width=1\textwidth]{./fig/rtlsdr}
-		\caption{The IQ demodulator RTL2832U is available for less than 10 \$. Source: Ebay}
-	\end{figure}
+\begin{minipage}{0.45\textwidth}
+	\begin{itemize}
+		\item<1-> Houses in the Netherlands are constantly sinking
+		\item<2-> Backscattering tags on the roofs next to the TV antenna
+		\item<3-> Sensor network with node which are highly-energy constrained
+	\end{itemize}
 \end{minipage}
@@
 
-\begin{frame}{Background And Literature (2)}
-	\begin{figure}[H]	
+\section{Looking for TV Signals in Uppsala}
+
+\begin{frame}{Spatial Variation of Ambient TV Signals (1)}
+\begin{figure}[h]
 	\centering
-		\includegraphics[height=0.75\textheight]{./fig/monopole_smith}
-		\caption{Smitchart sweep of selfmade ground-plane antenna for 634 MHz (checkout \cite{Pozar}).}
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
+	\end{minipage}
+	\caption{Single measurement at one point in space. Left: Raw spectrum. Right: Spectrum where average power had been determined.}
+	\label{fig:tv_record} 
 	\end{figure}
@@
 
-\begin{frame}{Background And Literature (3)}
-	\begin{figure}[H]	
+\begin{frame}{Spatial Variation of Ambient TV Signals (2)}
+\begin{figure}[h]
 	\centering
-		\includegraphics[height=0.75\textheight]{./fig/monopole_S11dB}
-		\caption{S11 out of the S parameter file of the VNA plotted with Qucs.}
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/haversine}
+	\end{minipage}
+	\hfill
+	\begin{minipage}{0.49\columnwidth}
+		\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
+	\end{minipage}
+	\caption{Multiple measuremnets in space. Left: Signal attenuation vs. distance from TV tower. Right: Interpolated heatmap for the \SI{626}{\mega\Hz}.}
+	\label{fig:haversine}
 \end{figure}
@@
 
-\section{Design}
-\begin{frame}{State Of The Project (1)}
-	\begin{table}[]
-	\centering
-	\begin{tabular}{ll}
-	\hline
-	\textbf{Backscatter Tag}               &    \\\hline
-	Manchester encoding                    &   \checkmark \\
-	Weak error protection                  &   \checkmark\\
-	Shifting by 2 MHz                      &   \checkmark\\
-	Advanced frame design                  &   \checkmark\\
-	\hline
-	\textbf{Antenna}                       &    \\\hline
-	Roughly tuned and matched              &    \checkmark\\\
-	Mechanical stability                   &    \\
-	Lumped elements tuning                 &    \\\hline
-	\end{tabular}
-	\caption{Backscatter tag and antenna}
-	\end{table}
-\end{frame}
-\begin{frame}{State Of The Project (2)}
-	\begin{table}[]
-	\centering
-	\begin{tabular}{ll}
-	\hline
-	\textbf{Receiver}                      &    \\\hline
-	Signal processing                      &    \checkmark\\\
-	Reception with certain bit error rate  &    \checkmark\\\
-	Efficient infrastructure               &    \checkmark\\\
-	Fast				       &    \\\hline
-	\textbf{Measurements}                  &    \\\hline
-	Processing and Measurement environment &    \checkmark\\\
-	Massive amount of samples              &    \\\hline
-	\end{tabular}
-	\caption{Receiver and measurements}
-	\end{table}
-\end{frame}
-
-\begin{frame}{State of the project (3)}
+\begin{frame}{Backscattering a Local Signal}
 	\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.75\textheight]{./fig/haversine}
-		\caption{Signal strength depending on the distance of different frequencies.}
+		\only<1>{\includegraphics[width=0.7\textwidth]{./fig/626mhz_raw.png}}
+			\only<2>{\includegraphics[width=0.7\textwidth]{./fig/626mhz_raw_arrow.png}}	
+		\caption{Spectrum of a local TV signal. Center frequency is \SI{626}{\mega\Hz}.}
 	\end{figure}	
@@
 
-\begin{frame}[fragile]{State of the project (3)}
-\begin{minted}[baselinestretch=1, fontsize=\tiny, linenos,frame=single,framesep=5pt]{Octave}
-function d = haversine (lat1,lat2,lon1,lon2)
-        R = 6371e3;                                                     % Earth radius
-        rad = pi/180;
+\section{Communication System Design}
 
-        phi1 = lat1*rad;
-        phi2 = lat2*rad;
-        dphi = (lat2-lat1)*rad;
-        dlambda = (lon2-lon1)*rad;
+\begin{frame}{Transmitter (1)}
+	\only<1>{\begin{figure}[H]	
+		\centering
+		\includegraphics[width=0.7\textwidth]{./fig/backscatter}
+	\end{figure}}
 
-        a = sin(dphi/2)*sin(dphi/2) + cos(phi1)*cos(phi2)*sin(dlambda/2)*sin(dlambda/2);
-        c = 2*atan2(sqrt(a), sqrt(1-a));
-        d = R*c;
-endfunction;
-\end{minted}
+	\only<2>{\begin{figure}[H]	
+		\centering
+		\includegraphics[width=0.7\textwidth]{./fig/backscatter_simple_bob}
+	\end{figure}}
 \end{frame}
 
-\begin{frame}{State of the project (3)}
+\begin{frame}{Transmitter (2)}
 	\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.75\textheight]{./fig/heatmap}
-		\caption{Signal strength heatmap of 626 MHz signal.}
+		\includegraphics[width=0.7\textwidth]{./fig/transmitter}
+		\caption{Transmitter architecture.
+Ground plane antenna roughly tuned to TV signal center freq.
+(\SI{626}{\mega\Hz}).
+Microcontroller steers a RF switch with a rectangular signal shifting the TV wave in another band.}
 	\end{figure}
@@
 
-\begin{frame}{State Of The Project (4)}
+\begin{frame}{Receiver (1)}	
 	\only<1>{\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch_blank}
-		\caption{System architecture of the receiver.}
+		\includegraphics[width=0.7\textwidth]{./fig/backscatter}
 	\end{figure}}
+
 	\only<2>{\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.7\textheight]{./fig/backscatter_arch}
-		\caption{System architecture of the receiver.}
+		\includegraphics[width=0.7\textwidth]{./fig/backscatter_simple_alice}
 	\end{figure}}
@@
 
-\begin{frame}{State Of The Project (5)}
+\begin{frame}{Receiver (2)}
+
+\begin{minipage}{0.45\textwidth}
 	\begin{figure}[H]	
 		\centering
-		\includegraphics[height=0.5\textheight]{./fig/osci}
-		\caption{Octave oscilloscope written for debugging purposes.}
+		\only<1>{\includegraphics[width=0.8\textwidth]{./fig/rtlsdr}}
+		\only<2,3>{\includegraphics[width=0.8\textwidth]{./fig/rtlsdr1}}
+		\only<4>{\includegraphics[width=0.8\textwidth]{./fig/rtlsdr2}}
+		\caption{RTL-SDR hardware with the DVB-T I/Q demodulator
+		Raeltek RTL2832U (left IC) and the tuner with integrated LNA
+		Rafael Micro R820T/2 (right IC).}
+	\end{figure}
+\end{minipage}
+\hfill
+\begin{minipage}{0.45\textwidth}
+	\begin{itemize}
+		\item<2-> RTL2832U is cheap  
+		\item<3-> Access to I/Q demodulator (DAB and FM radio) can be hacked		\item<4-> Tuner (LNA, filters etc.) included
+	\end{itemize}
+\end{minipage}
+\end{frame}
+
+\begin{frame}{Receiver (3)}
+	\begin{figure}[H]	
+		\centering
+		\only<1>{\includegraphics[width=0.9\textwidth]{./fig/receiver1}}
+		\only<2>{\includegraphics[width=0.9\textwidth]{./fig/receiver2}}
+		\caption{Architecture of backscatter receiver.}
 	\end{figure}
@@
 
-\section{Evaluation}
-\begin{frame}{Conclusion}
+\section{Results and Outlook}
+
+\begin{frame}{Communication Performance (1)}
+	\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\textwidth]{./fig/transmission}
+	\caption{Start of a transmission. Different signal processing steps are visible. First: Raw downsampled signal amplitude. Second: Signal after low-pass filter. Third: Signal after Schmitt trigger.}
+	\label{fig:transmission}	
+\end{figure}
+\end{frame}
+
+\begin{frame}{Communication Performance (2)}
 	\begin{itemize}
-	\item Working receiver \checkmark
-	\item Spectrum anlayzer based on the RTL2832U \checkmark 
-	\item Working backscatter tag \checkmark
-	\item Lot of help from the communication group \checkmark
+		\item<1-> We varied the bitrate between 1 bit/s and 1 kbit/s 
+		\item<2-> Strength of the backscattered signal is quite low (high quantisation noise)
+		\item<3-> Higher bitrates (1 kbit/s) lead to a range of a couple of decimeter before the error rate goes up rapidly
+		\item<4-> High bit error rate is due to the not yet customized HW
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Results}
+	\begin{itemize}
+		\item<1-> Spectrum scanning over a huge band is possible using the RTL-SDR
+		\item<2-> Uppsala is adequate for backscatter communication using our receiver
+		\item<3-> Backscatter communication is possible with the RTL-SDR
 	\end{itemize}
@@
 	\begin{itemize}
-	\item Reliable communication (e.g. hamming code)
-	\item More advanced signal processing (e.g. oversampling) 
-	\item Better (selfmade) antenna
-	\item Other receiver hardware? 
+		\item<1-> Other antennas for reception and transmission have to be tried out
+		\item<2-> Standard cable and antenna of the RTL-SDR (included in the 10 USD budget) are unsuiteable for our application as we realized
+		\item<3-> The filters involved are setscrews as well as the coding used by the communication system
 	\end{itemize}
@@
 
-   	\begin{frame}{Contact Information}
-        	\begin{center}
-                \begin{table}[]
-                        \begin{tabular}{ll}
-                                E: & Elmar.Vanrijnswou.9818@student.uu.se \\
-                             	E: & Maximilian.Stiefel.8233@student.uu.se\\
-                        \end{tabular}
-                \end{table}
-                \includegraphics[width=0.05\textwidth]{./fig/github} \hspace{0.1cm} \url{https://github.com/m3x1m0m}\\
-        	\end{center}
-        	\begin{figure}
-                	\includegraphics[width=0.3\textwidth]{./fig/logo}
-        	\end{figure}
-  	\end{frame}
-
 \begin{frame}[standout]
-	Happy Coding :) 
+	Thank you for your attention. 
 \end{frame}
@@
   \bibliography{presentation}
-  \bibliographystyle{abbrv}
+  \bibliographystyle{apalike}
 
